---
title: "Black Sea Bass Range Expansion"
author: 
  - first_name: "Adam"
    last_name: "Kemberling"
    url: https://github.com/adamkemberling
    affiliation: Gulf of Maine Research Institute
    affiliation_url: https://www.gmri.org
description: | 
  Mapping Black Sea Bass With Ocean Temperature Changes
date: "Updated on: `r Sys.Date()`"
editor_options: 
  chunk_output_type: console
output: 
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

####  Packages  ####
library(targets)
library(tidyverse)
library(gmRi)
library(ggforce)
library(sf)
library(rnaturalearth)
library(terra)


####  Mapping  ####

oisst_path <- cs_path("res", "OISST/oisst_mainstays")

# What area is the report using?
group_name <- "gmri_sst_focal_areas"
region_name <- "apershing_gulf_of_maine"

# File paths for various extents based on "apershing_gulf_of_maine"
region_paths <- get_timeseries_paths(region_group = group_name, 
                                     mac_os = "mojave")

# Timeseries Path
timeseries_path <- region_paths[[region_name]][["timeseries_path"]]
poly_path <- region_paths[[region_name]][["shape_path"]]

# Coastline Shapes
new_england <- ne_states("united states of america", returnclass = "sf")
canada      <- ne_states("canada", returnclass = "sf")

# Load the timeseries for just GOM temps
region_timeseries <- read_csv(timeseries_path, col_types = cols(), guess_max = 1e6)
```

# Black Sea Bass EDA

Look into the habitat use by BSB, see if there is a convex hull or some other way to show trends over time.




####  Biological Data  ####

Obvious resources include both the federal and state trawl survey datasets. Things to consider could be the range extent, center of mass, trailing edge, some sort of latitude summary.

```{r}
# Load the area stratified data
withr::with_dir(rprojroot::find_root('_targets.R'), 
                tar_load(nefsc_stratified))

# Load the black seabass data only, and for Fall
bsb <- nefsc_stratified %>% 
  filter(
    comname == "black sea bass"#,
    #season == "Fall"
    ) %>% 
  drop_na(surftemp, bottemp, numlen_adj) 
```


```{r}
plotrix::weighted.hist(x = bsb$surftemp, w = bsb$numlen_adj, main = "Black Sea Bass Catch in NE Groundfish", xlab = "Surface Temp")
```


```{r}
plotrix::weighted.hist(x = bsb$bottemp, w = bsb$numlen_adj, main = "Black Sea Bass Catch in NE Groundfish", xlab = "Bottom Temp")
hist(bsb$surftemp)
```

###  Federal Survey Abundances

```{r}
decade_summs <- bsb %>% 
  mutate(lat_tenth = round(decdeg_beglat, 1), 
         lon_tenth = round(decdeg_beglon, 1),
         decade = floor_decade(est_year)) %>% 
  group_by(decade, lat_tenth, lon_tenth) %>% 
  summarise(strat_bio = sum(strat_total_biom_s),
            strat_abund = sum(strat_total_abund_s),
            .groups = "drop") %>% 
  st_as_sf(coords = c("lon_tenth", "lat_tenth"), crs = 4326, remove = FALSE)


# # Convex hulls
# hull_data <- decade_summs %>%
#   group_by(decade) %>% 
#   slice(chull(lat_tenth, lon_tenth))


# Map them
ggplot() +
  #geom_polygon(data = hull_data, aes(lon_tenth, lat_tenth, fill = decade), alpha = 0.2) +
  geom_sf(data = decade_summs, aes(size = strat_abund, color = decade), alpha = 0.5) +
  geom_sf(data = new_england, size = 0.3) +
  facet_wrap(~decade, nrow = 1) +
  scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
  coord_sf(xlim = c(-75.64, -67.21), ylim = c(40, 44.25)) +
  theme(legend.position = "bottom") +
  labs(size = "Stratified Abundance", x = "Lon", y = "Lat", title = "Black Sea Bass Stratified Abundance", subtitle = "By decade") +
  guides(color = guide_legend("Decade", title.position = "top", title.hjust = 0.5, nrow = 2),
         fill = guide_legend("Decade", title.position = "top", title.hjust = 0.5, nrow = 2),
         size = guide_legend(title.position = "top", title.hjust = 0.5, nrow = 2))

```



### Oceanographic Data

Looking at it from a thermal preference angle, there is the possibility to look at temperature suitability. This could be summer or winter displays of temperature envelopes that fit their thermal preferences.

**Reference on Habitat Suitability**
https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0147627&type=printable

```{r}
# Load Monthly OISST Data
sst_path <- cs_path("res", "OISST/oisst_mainstays")

library(terra)
month_sst <- terra::rast(str_c(sst_path, "monthly_averages/oisst_monthly.nc"))

# rename layers using time
month_times <- str_sub(time(month_sst), 1, 10)
names(month_sst) <- month_times
plot(month_sst$`1981-09-30`)


# Cropping to Gulf of Maine  
gom_sf <- read_sf(poly_path)

# Change extent to match crs 4326
ext(month_sst) <- ext(c(-180,180,-90,90))
crs(month_sst) <- crs("4326")


gom_monthly <- terra::crop(month_sst, gom_sf)



# Make Seasonal Averages
seasonal_avgs <- function(year_data, season){
  
  # Get the names (year and month)
  year_info <- str_sub(names(year_data), 1,4)
  month_info <- str_sub(names(year_data), 6,7)
  
  # Set the months to subset
  season_months <- switch(season,
  "winter" = c("12", "01", "02"),
  "spring" = c("03", "04", "05"),
  "summer" = c("06", "07", "08"),
  "fall"   = c("09", "10", "11"))
  
  # Identify the months that match
  month_subset <- which(month_info %in% season_months)
  
  # Subset the months
  season_subset <- subset(year_data, month_subset)
  
  # Average it
  season_avg <- mean(season_subset)
  #names(season_avg) <- str_c(season, year_info, sep = "-")
  return(season_avg)
}


# The years with full data
sst_years <- c(1982:2021)

# Split them out by year from the full stack
sst_years_dat <- map(sst_years, function(x){
  subset(gom_monthly, which(str_sub(names(gom_monthly), 1,4) == x))}) %>% 
  setNames(sst_years)

# Average the seasons for each year
winter_yearly <- map(sst_years_dat, seasonal_avgs, "winter")
spring_yearly <- map(sst_years_dat, seasonal_avgs, "spring")
summer_yearly <- map(sst_years_dat, seasonal_avgs, "summer")
fall_yearly   <- map(sst_years_dat, seasonal_avgs, "fall")

# Re-stack them
winter_avgs <- terra::rast(winter_yearly)
spring_avgs <- terra::rast(spring_yearly)
summer_avgs <- terra::rast(summer_yearly)
fall_avgs <- terra::rast(fall_yearly)


# global function replaces cellstats
seasonal_timelines <- bind_rows(list(
  "winter" = terra::global(winter_avgs, "mean", na.rm = T) %>% rownames_to_column(var = "year"), 
  "spring" = terra::global(spring_avgs, "mean", na.rm = T) %>% rownames_to_column(var = "year"), 
  "summer" = terra::global(summer_avgs, "mean", na.rm = T) %>% rownames_to_column(var = "year"), 
  "fall"   = terra::global(fall_avgs, "mean", na.rm = T)   %>% rownames_to_column(var = "year")),
  .id = "season"
)
  

# Plot seasonal averages
ggplot(seasonal_timelines, aes(year, mean, color = season)) +
  geom_line(aes(group = season))
```



