---
title: "Bodymass Size Spectra Surfaces"
author: "Adam A. Kemberling"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float:
        collapsed: FALSE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = F, 
                      warning = F, 
                      comment = "", 
                      fig.height = 6, 
                      fig.width = 6)

####  Packages  ####
library(targets)
library(here)
library(sf)
library(gmRi)
library(patchwork)
library(knitr)
library(tidyverse)
library(rnaturalearth)
library(plotly)

# Support functions
source(here("R/support/sizeSpectra_support.R"))
res_path <- cs_path("res")

# Set theme  
theme_set(theme_bw() +
            theme(legend.position = "bottom",
                  strip.text = element_text(color = "white", 
                                            face = "bold"),
                  strip.background = element_rect(
                    color = "white", 
                    #fill = "#36454F", # charcoal
                    fill = "#00736D",
                    size = 0.75, 
                    linetype="solid"), 
            legend.background = element_rect(fill = "transparent", 
                                             color = "black")))
```

`r use_gmri_style_rmd(css_file = "gmri_rmarkdown.css")`


#  What is a Bodymass Size Spectrum?

A community size spectrum is the relationship between abundance and body size. This shape and structure of this relationship has been shown to be temporally stable and reflect characteristics of the environments that support it (productivity levels). 

This relationship is often plotted on a log body size x log abundance scale which reveals a linear relationship between the two. The slope and intercept of this relationship have been used to make comparisons across ecosystems regarding their productivity and the efficiency of energy transfer within a community.


## Regions of Interest

```{r}
# Load the shapefiles for the polygons
new_england <- ne_states("united states of america", returnclass = "sf")
canada <- ne_states("canada", returnclass = "sf")

# Load the strata
survey_strata <- read_sf(str_c(res_path, "Shapefiles/BottomTrawlStrata/BTS_Strata.shp"))  %>% 
  janitor::clean_names() %>% 
  filter(strata >= 01010 ,
         strata <= 01760,
         strata != 1310,
         strata != 1320,
         strata != 1330,
         strata != 1350,
         strata != 1410,
         strata != 1420,
         strata != 1490) 


# Key to which strata = which regions
strata_key <- list(
  "Georges Bank"          = as.character(13:23),
  "Gulf of Maine"         = as.character(24:40),
  "Southern New England"  = str_pad(as.character(1:12), width = 2, pad = "0", side = "left"),
  "Mid-Atlantic Bight"    = as.character(61:76))


# Assign Areas
survey_strata <- survey_strata %>% 
  mutate(
    strata = str_pad(strata, width = 5, pad = "0", side = "left"),
    strata_num = str_sub(strata, 3, 4),
    area = case_when(
      strata_num %in% strata_key$`Georges Bank` ~ "Georges Bank",
      strata_num %in% strata_key$`Gulf of Maine` ~ "Gulf of Maine",
      strata_num %in% strata_key$`Southern New England` ~ "Southern New England",
      strata_num %in% strata_key$`Mid-Atlantic Bight` ~ "Mid-Atlantic Bight",
    TRUE ~ "Outside Major Study Areas"
  )) %>% 
  select(finstr_id, strata, strata_num, area, a2, str2, set, stratuma, str3, geometry)


# Map it
strata_plot <- ggplot() +
  geom_sf(data = new_england) +
  geom_sf(data = canada) +
  geom_sf(data = survey_strata, aes(fill = area)) +
  coord_sf(xlim = c(-77, -65), ylim = c(34, 45.75), expand = FALSE) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  labs(subtitle = "Survey Regions")

strata_plot
```

## Comparing Spectra Parameters {.tabset .tabset-pills}

```{r}
# Load the spectra coefficients
withr::with_dir(rprojroot::find_root('_targets.R'), 
                tar_load(size_spectrum_indices))

# Pull the ones that are regions and years
region_dat <- size_spectrum_indices %>% 
  filter(`group ID` == "single years * region") %>% mutate(Year = as.numeric(Year)) %>% 
  mutate(survey_area = factor(survey_area, levels = c("GoM", "GB", "MAB", "SNE")))

# Put them in list for lata
region_indices <- region_dat %>% 
  split(.$survey_area)



slopes <- region_dat %>% 
  ggplot(aes(Year, l10_slope_strat)) +
  geom_line(size = 0.5, color = "gray50") +
  geom_point(size = 1,color = "gray10") +
  facet_wrap(~survey_area, ncol = 1) +
  labs(subtitle = "Size Spectrum Slope", y = "")

ints <- region_dat %>% 
  ggplot(aes(Year, l10_int_strat)) +
  geom_line(size = 0.5, color = "gray50") +
  geom_point(size = 1,color = "gray10") +
  facet_wrap(~survey_area, ncol = 1) +
  labs(subtitle = "Size Spectrum Intercept", y ="")

slopes | ints
```

## Biomass and Abundance Trends

```{r}
# Load the data used to estimate spectra
withr::with_dir(rprojroot::find_root('_targets.R'), 
                tar_load(nefsc_1g_binned))      


# Get yearly biomass and abundance
yearly_summs <- nefsc_1g_binned %>% 
  group_by(Year, survey_area) %>% 
  summarise(strat_biomass = sum(strat_total_lwbio_s, na.rm = T),
            strat_abund = sum(strat_total_abund_s, na.rm = T),
            .groups = "drop") %>% 
  ungroup() %>% 
  mutate(survey_area = factor(survey_area, levels = c("GoM", "GB", "SNE",  "MAB")))



abund_p <- ggplot(yearly_summs, aes(Year, strat_abund)) +
  geom_line(size = 0.5, color = "gray50") +
  geom_point(size = 1,color = "gray10") +
  facet_wrap(~survey_area, ncol = 1, scales = "free") +
  scale_y_continuous(labels = comma_format()) +
  labs(subtitle = "Stratified Abundance", y ="")


biom_p <- ggplot(yearly_summs, aes(Year, strat_biomass)) +
  geom_line(size = 0.5, color = "gray50") +
  geom_point(size = 1,color = "gray10") +
  facet_wrap(~survey_area, ncol = 1, scales = "free") +
  scale_y_continuous(labels = comma_format()) +
  labs(subtitle = "Stratified Biomass", y ="")



abund_p | biom_p
```



## Big Fish & Small Fish Trends


```{r}
fish_sizes <- nefsc_1g_binned %>% 
  mutate(fish_size = ifelse(ind_weight_kg <= 2, "Small <= 2kg", "Large > 2kg")) %>% 
  group_by(Year, survey_area, fish_size) %>% 
  summarise(strat_biomass = sum(strat_total_lwbio_s, na.rm = T),
            strat_abund = sum(strat_total_abund_s, na.rm = T),
            .groups = "drop") %>% 
  mutate(survey_area = factor(survey_area, levels = c("GoM", "GB", "SNE",  "MAB")),
         fish_size = factor(fish_size, levels = c("Small <= 2kg", "Large > 2kg")))
```


```{r, fig.width=8}
ggplot(fish_sizes, aes(Year, strat_abund)) +
  geom_line(size = 0.5, color = "gray50") +
  geom_point(size = 1,color = "gray10") +
  facet_grid(fish_size~survey_area, scales = "free") +
  scale_y_continuous(labels = comma_format()) +
  labs(subtitle = "Stratified Abundance", y ="")
```


```{r, fig.width=8}
ggplot(fish_sizes, aes(Year, strat_biomass)) +
  geom_line(size = 0.5, color = "gray50") +
  geom_point(size = 1,color = "gray10") +
  facet_grid(fish_size~survey_area, scales = "free") +
  scale_y_continuous(labels = comma_format()) +
  labs(subtitle = "Stratified Biomass", y ="")

```




## Visualizing the Spectra

The body size spectra were estimated for each of these regions independently. Abundances were area-stratified, then grouped into log bodymass bins. 3d Surfaces below display how the shape of the spectra change over time. 

These surfaces are followed by the resulting size spectrum parameters of the annual size spectra.



```{r}
      
# Split the data into the groups we want to compare:
areas_binned <- nefsc_1g_binned %>% 
  split(.$survey_area)


# Get annual totals within each area
areas_aggregated <- areas_binned %>% 
  map(function(area_data){
    area_data %>% 
      split(.$Year) %>% 
      map_dfr(function(yr_data){
        aggregate_l10_bins(yr_data) %>%
          mutate(log10_bins = as.numeric(log10_bins),
                 norm_strat_abund = log10(norm_strat_abund),
                 norm_strat_abund = ifelse(is.na(norm_strat_abund), 0, norm_strat_abund))
          
      }, .id = "Year")
  })
```


```{r surface plot function}
####  Surface Plots
spectra_surface <- function(l10_totals, plot_title){
  
  # Pull data for matrix
  dat <- select(l10_totals, left_lim, Year, norm_strat_abund)
  
  # Make matrix to plot
  spectrum_mat <- as.matrix(xtabs(norm_strat_abund~left_lim + Year, data = dat))

  # Build plotly plot
  surf_plot <- plot_ly(
      x = as.numeric(colnames(spectrum_mat)), 
      y = as.numeric(rownames(spectrum_mat)), 
      z = spectrum_mat
    ) %>% 
  add_surface() %>%
  layout(
    title = str_c(plot_title, " - Bodymass Spectrum"),
    scene = list(
      xaxis = list(title = "Year"),
      yaxis = list(title = "Log10( BodyMass (g) )"),
      zaxis = list(title = "Log10( Abundance) "),
      camera = list(eye = list(x = 1, y = 1, z = 1.25))
    ))
  
  #return plot
  return(surf_plot)
  
}


# Spectrum parameter plots
plot_params <- function(area_abbrev){
  slope_p <- ggplot(region_indices[[area_abbrev]], aes(Year, l10_slope_strat)) +
    geom_line() +
    geom_point() +
    labs(y = "Size Spectrum Slope", x = "")
  
  int_p <- ggplot(region_indices[[area_abbrev]], aes(Year, l10_int_strat)) +
    geom_line() +
    geom_point() +
    labs(y = "Size Spectrum Intercept")
  
  
  bin_data <- areas_aggregated[[area_abbrev]] %>% 
    mutate(Year = as.numeric(Year),
           l10_bins = factor(left_lim))
  bin_p <- ggplot(bin_data, aes(x = Year, y = norm_strat_abund, color = l10_bins)) +
    geom_line(aes(group = log10_bins)) +
    geom_point() + 
    scale_color_viridis_d(direction = -1) +
    guides(color = guide_legend(title.position = "top", title.hjust = 0.5)) +
    labs(x = "Year", y = "Log( Abundance )", color = "Body Size Bin log10(g)")
  
  layout <- "
  AA##
  AACC
  BBCC
  BB##
  "
  plot_mosaic <- slope_p / int_p / bin_p
  return( plot_mosaic)
}


```


###  Gulf of Maine


```{r}
spectra_surface(l10_totals = areas_aggregated$GoM, "Gulf of Maine")
```

```{r, fig.height=8}
plot_params("GoM")
```



### Georges Bank

```{r}
spectra_surface(l10_totals = areas_aggregated$GB, "Georges Bank")
```

```{r, fig.height=8}
plot_params("GB")
```


### Southern New England

```{r}
spectra_surface(l10_totals = areas_aggregated$SNE, "Southern New England")
```

```{r, fig.height=8}
plot_params("SNE")
```


### Mid-Atlantic Bight


```{r}
spectra_surface(l10_totals = areas_aggregated$MAB, "Mid-Atlantic Bight")
```

```{r, fig.height=8}
plot_params("MAB")
```



`r insert_gmri_footer()`




`r insert_gmri_footer()`
