---
title: "Community Size-Structure of the Northwest Atlantic Groundfish Communities, Response to Direct Disturbance and a Changing Environment"
author: 
    name: "Adam A. Kemberling"
    title: "Quantitative Research Associate"
    url: https://github.com/adamkemberling
    affiliation: Gulf of Maine Research Institute
description: | 
  Results from a Individual Size Distribution Analysis of the Northeast US Groundfish Community
date: "`r Sys.Date()`"
# format: docx
format:
  html:
    self-contained: true
    code-fold: true
    code-tools: true
    df-print: kable
    toc: true
    toc-depth: 2
editor: source
execute: 
  echo: false
  warning: false
  message: false
  fig.height: 6
  fig.width: 7
  fig.align: center
  comment: ""
bibliography: references.bib
csl: ices-journal-of-marine-science.csl
---

```{r}
#| label: load packages and functions

####  Packages  ####
library(targets)
library(rnaturalearth)
library(here)
library(sf)
library(gmRi)
library(patchwork)
library(gt)
library(knitr)
library(tidyverse)
library(ggstream)
library(broom)
# library(ggforce)
library(bcp)
library(scales)
library(corrplot)
library(readxl)
library(Hmisc)



# Support functions
source(here("R/support/sizeSpectra_support.R"))

# Set theme
theme_set(theme_gmri() + theme(
  panel.grid.major = element_line(linewidth = 0.5, linetype = 1, color = "gray"),
  panel.grid.major.x = element_line(linewidth = 0.5, linetype = 3, color = "gray"),
  panel.grid.minor = element_line(linewidth = 0.5, linetype = 3, color = "gray90"),
  axis.line = element_line(color = "black"),
  axis.line.y = element_line(color = "black"), 
  strip.background = element_rect(colour = gmri_cols("teal")),
  legend.position = "bottom")
  )

# Map polygons
us_poly <- ne_states("united states of america", returnclass = "sf")
canada <- ne_states("canada", returnclass = "sf")

# levels for faceting areas
area_levels <- c("Northeast Shelf, Full", "GoM", "GB", "SNE", "MAB")
area_levels_long <- c("Northeast Shelf, Full", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight")
```

```{r}
#| label: functions-cell

# Function to process summaries for various factor combinations
get_group_summaries <- function(data_in, ...){
  
  # Do some grouping to get totals
  group_totals <- data_in %>% 
    group_by(...) %>% 
    summarise(total_survey_catch = sum(numlen, na.rm = T),
              total_lw_bio = sum(sum_weight_kg, na.rm = T),
              total_strat_abund = sum(strat_total_abund_s, na.rm = T),
              total_strat_lw_bio = sum(strat_total_lwbio_s, na.rm = T), 
              .groups = "drop") 
  
  # length bins
  group_lengths <- data_in %>% 
    group_by(..., length_bin) %>% 
    summarise(lenbin_survey_catch = sum(numlen),
              lenbin_lw_bio       = sum(sum_weight_kg),
              lenbin_strat_abund  = sum(strat_total_abund_s),
              lenbin_strat_lw_bio = sum(strat_total_lwbio_s), 
              .groups = "drop") %>% 
    left_join(group_totals) %>% 
    mutate(
      perc_total_catch  = (lenbin_survey_catch - total_survey_catch) * 100,
      perc_lw_bio       = (lenbin_lw_bio - total_lw_bio) * 100,
      perc_strat_abund  = (lenbin_strat_abund - total_strat_abund) * 100,
      perc_strat_lw_bio = (lenbin_strat_lw_bio - total_strat_lw_bio) * 100)
  
  # weight bins
  group_weights <- data_in %>% 
    group_by(..., weight_bin) %>% 
    summarise(wtbin_survey_catch = sum(numlen),
              wtbin_lw_bio       = sum(sum_weight_kg),
              wtbin_strat_abund  = sum(strat_total_abund_s),
              wtbin_strat_lw_bio = sum(strat_total_lwbio_s),
              .groups = "drop") %>% 
    left_join(group_totals) %>% 
    mutate(
      perc_total_catch  = (wtbin_survey_catch / total_survey_catch) * 100,
      perc_lw_bio       = (wtbin_lw_bio / total_lw_bio) * 100,
      perc_strat_abund  = (wtbin_strat_abund / total_strat_abund) * 100,
      perc_strat_lw_bio = (wtbin_strat_lw_bio / total_strat_lw_bio) * 100)

return(list("length_bins" = drop_na(group_lengths),
            "weight_bins" = drop_na(group_weights)))
}



# Function to process summaries for various factor combinations
log2_bin_summaries <- function(data_in, ...){
  
  # Do some grouping to get totals
  group_totals <- data_in %>% 
    group_by(...) %>% 
    summarise(total_survey_catch = sum(numlen, na.rm = T),
              total_lw_bio = sum(sum_weight_kg, na.rm = T),
              total_strat_abund = sum(strat_total_abund_s, na.rm = T),
              total_strat_lw_bio = sum(strat_total_lwbio_s, na.rm = T), 
              .groups = "drop") 
  
  # weight bins
  group_weights <- data_in %>% 
    mutate(weight_bin = bin_label) %>% 
    group_by(..., left_lim) %>% 
    summarise(wtbin_survey_catch = sum(numlen),
              wtbin_lw_bio       = sum(sum_weight_kg),
              wtbin_strat_abund  = sum(strat_total_abund_s),
              wtbin_strat_lw_bio = sum(strat_total_lwbio_s),
              .groups = "drop") %>% 
    left_join(group_totals) %>% 
    mutate(
      perc_total_catch  = (wtbin_survey_catch / total_survey_catch) * 100,
      perc_lw_bio       = (wtbin_lw_bio / total_lw_bio) * 100,
      perc_strat_abund  = (wtbin_strat_abund / total_strat_abund) * 100,
      perc_strat_lw_bio = (wtbin_strat_lw_bio / total_strat_lw_bio) * 100)

return(list("weight_bins" = drop_na(group_weights)))
}




# Function to grab the correlation data and lag data
get_ccf_vector <- function(x,y){
  
  # Run the ccf
  ccf_data <- ccf(x,y,plot= F , lag.max = 15)
  
  # Get the signif:
  # https://www.squaregoldfish.co.uk/programming/r_acf_significance.md/
  # Not 100% sure n is the same for ccf as it is for acf, but...
  significance_level <- qnorm((1 + 0.95)/2)/sqrt(sum(!is.na(x)))
  
  data.frame(
    "acf" = ccf_data$acf,
    "lag" = ccf_data$lag,
    "sigpos" = significance_level,
    "signeg" = significance_level*-1
  )
}


# Plotting the correlation matrix over a specific decade
plot_period_correlations <- function(x, label) {
  x %>% 
  ggplot(aes(var1, response_short, fill = r2)) +
  geom_tile() +
  geom_text(aes(label = signif)) +
  facet_grid(response_area ~ driver_type, scales = "free_x", space = "free") +
  scale_fill_distiller(palette = "RdBu", limits = c(-1, 1), breaks = c(-1,0,1)) +
  coord_cartesian(clip = "off") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        strip.text.y = element_text(angle = 0),
        legend.position = c(1.205, -.25),
        legend.title = element_text(vjust = 0.75),
        plot.caption = element_text(hjust = 0)) +
  guides(fill = guide_colorbar(title.position = "top", title.hjust = 0.5)) +
  labs(x = "Correlated Feature", y = "Normalized Abundance Spectra Coefficient",
       fill = "Correlation Coefficient",
       title = str_c(label, " | Correlation of Size Spectra to Hypothesized Drivers"),
       subtitle = "Correlation coefficients shown display same-year correlations.",
       caption = md("* Symbol signals a significant correlation at an alpha =  0.05"))
  
}

```

### Potential Journals:

1.  **ICES Journal of Marine Science**

# Abstract

The surface waters of the Northwest Atlantic Ocean are among the fastest warming on Earth. This area is highly-productive biologically, and there are concerns that ecological consequences will follow this rapid-warming. Research on the impacts of this rapid warming has primarily focused on high-profile and/or upper trophic level species. Ecological theory and laboratory studies suggest that elevated temperatures facilitate early maturation and smaller adult body-sizes. However, it is unclear whether that relationship might be mitigated against through adaptive behaviors in an open ocean environment. Here we've investigated ecosystem wide impacts on the individual size distribution (ISD) to track changes in community size structure. In cases where community responses are not adequate to counter the impacts of elevated temperatures, we anticipated a steepening of the size spectrum slope (ISD exponent). A steeper relationship relating to a reduction in larger sized individuals and an increased prevalence of smaller sized individuals. Using data from fisheries independent surveys we calculated the community size spectra for four regions along the US NE continental shelf. Correlation/regression analyses were then performed to then assess the degree to which these changes were in alignment to hypothesized bottom-up and top-down disturbances. At the regional scale, we found that community size structure changes (spectra slope) were the largest in the Northern regions, in the Gulf of Maine and Georges Bank. These areas are home to the coldest temperatures and the largest proportions of groundfish species in the community. Spectrum slope declines were most pronounced in the 80's and 90's, before the rapid warming of the last decade. The timing of these declines suggest that external factors drove the initial declines of larger-sized individuals within the communities, before elevated temperatures began to influence the ecosystem. Correlation analyses reveal that while fisheries landings are strongly correlated with these declines, bottom-up factors of zooplankton community metrics, Gulf Stream Index, and SST anomalies are also important. While the primary pressure of fisheries exploitation has declined dramatically over time, the recovery of larger-sized individuals has not been seen. That kind of recovery will likely depend on the elevated temperatures seen over the last decade.

# Introduction

## Normalized Size Spectra & Individual Size Distributions

Size is a defining characteristic of species and mediates many ecological interactions and metabolic pathways (@brown2000) . Size is a big factor in determining the mobility of an organism. Mobility then impacts the ability to evade predation, foraging success, efforts to locate and follow essential habitats, geographic ranging, and the metabolic costs associated with each of these behaviors (@hillaert2018). Body size also mediates vulnerability to aspects the immediate environment such as temperature through heat exchange or the threat of desiccation in terrestrial species (@gillooly2001; @heatwole1969). Body size even informs life history features like life span and the trophic position an individual might occupy through its impact on metabolism and resource use (@white2007).

Size structured environments are a fundamental organizational pattern globally that emerges from these relationships **add_citation**. Within strongly size-structured ecosystems, growth and maturity changes alter fitness and ultimately determine whether a species is successful in that environment **add_citation** . Ecological theory is rich with models relating how energy transfers from smaller prey species to larger predatory trophic levels, the allocation of energy for growth, and the trade offs of allocating energy towards those ends (@bertalanffy1938; @vonbertalanffy1957; **add_citations)**. A globally persistent pattern in ecology entangled in those relationships and their critiques is the decline in abundance with increasing body size (@damuth1981; @currie1993; @sheldon1972; @loeuille2006).

This relationship, between size and abundance, integrates multiple processes operating on the cellular, individual, and community levels simultaneously. The quantities for size and abundance are also some of the most readily collected data assets of any ecological community. This creates an opportunity to learn much about a system from a relatively low-effort in data collection. For these reasons, size spectrum analyses and individual size distribution (ISD) methods have gained increasing attention as an entry point to assessing ecosystem health and to detect system-wide disturbance (@shin2005'; @pomeranz2022; @white2007). An advantage of these models is they avoid the need to explicitly articulate each predator-prey interactions as they and can be estimated from the commonly collected measures of abundance and size. The "size spectrum" describes the distribution of biomass or abundance as a function of individuals' mass or size on a log--log scale (@guiet2016; @kerr2001) . Size spectra are described by two terms, the size spectrum slope & intercept. These two terms reveal a sense of the baseline productivity, and how energy flows through an ecosystem (in the form of biomass) from many smaller individuals to many fewer large individuals. The spectrum intercept has been linked to the productivity of the community, and is often connected to the prevailing environmental conditions (@boudreau1992; @rossberg2012).

**How are they used in practice:**

Size spectra condense the complexities of predator prey networks and their interactions into a handful of size-based indices. These indices capture the emergent properties of a system, and have become increasingly used as indicators of ecosystem health. Within the context of fisheries management, changes in spectrum slopes have been associated with fishing exploitation, primarily through the targeted removal of larger individuals (@bianchi2000; @shin2005). Numerical experiments have also linked changes in slope to environmental disturbances (@guiet2016). Size spectra have also been shown to express predictable relationships between ecosystems of similar productivity levels as well as from distinct temperatures (@guiet2016).

**Use Pomerantz paper & Edwards to extend into ISD**

## Temperature & Ecology

Temperature plays a critical role on biological life impacting many of the chemical reactions that underpin basic physiological function. Temperature has direct and indirect impacts on critical biological functions including the acquisition of biomass through feeding, the rates of biomass loss through metabolism, and the rates at which individuals mature and develop. Because of these relationships, most species have evolved thermal preferences around which these chemical reactions are optimized. Species that are unable to maintain their thermal preferences internally must be able to follow their thermal preference in the environment through locomotion or adapt to less-favorable conditions through changes in behavior or risk metabolic costs in failing to do so. In an era of anthropogenic climate change, there is an expectation that many species will be displaced from historic habitats in their efforts to follow their thermal preferences. Recent research in marine environments has shown evidence of this as species are now shifting to higher latitudes and to deeper depths in the pursuit of more favorable conditions (@kleisner2017; @pinsky2013). Others have suggested that temperature related impacts may not be seen through geographic distribution change, but through physiological changes, changes in seasonal phenology, or in dashed hopes of species recovery (@daan2005; @miller2018; @pershing2015; @meyer-gutbrod2021).

**Need to connect temperature to size here**

Direct quote from @guiet2016 , but nails the connection back to temp expectations:

> Because it controls chemical reactions, temperature controls metabolic rates which underpin maintenance, growth or reproduction (Clarke and Johnston, 1999; Kooijman, 2010) as well as the functional responses to food density (Rall et al., 2012). Guiet et al. (2016)... In addition to the impact of temperature on communities' intercepts (heights), the impact of temperature on the speed of the energy flow within communities may affect other properties, such as their resilience to perturbations or the intensity of trophic cascades (Andersen and Pedersen, 2009).

The potential for elevated temperatures to impact the size structure of an ecosystem has implications for the ecosystem resilience in the face of climate change, as well as the blue economies & natural resource systems that rely upon their good health.

### Temperature of the Gulf of Maine & NE Shelf

In addition to the ecological disturbances of industrial fisheries, the Northwest Atlantic is also one of the fastest warming locations in the global oceans. Sea surface temperatures in the Gulf of Maine since 1982 have been warming at rates faster than 96% of the world's oceans, with similar warming rates along the northwest Atlantic shelf (@pershing2018). This persistent elevated temperature regime of the area is a result of several forces, a combination of shifting ocean currents and the unique bathymetry of the region. A Northward shift in the Gulf Stream directly increased the regional temperatures through increased transport of warm Gulf Stream water into areas like the Gulf of Maine. The Northward Gulf Stream shift is associated with a higher frequency of warm core rings, and the obstruction of cold-water Scotian Shelf current flow that would otherwise counter the influence of the Gulf Stream on the region's temperatures (@gangopadhyay2019; @meyer-gutbrod2021). The combination of these oceanographic changes has led to a warmer continental shelf habitat.

The rapid warming in the northwest Atlantic is a major factor in the redistribution of marine species along the US east coast. Species have responded by adjusting the timing and locations of their seasonal migrations and shifting their geographic ranges (@nye2009; @staudinger2019). There is evidence that warming has hampered fisheries recoveries as well. Adding a metabolic tax to physiological pathways like growth and metabolism. Species like Black Sea Bass, Atlantic shortfin squid, and Blue crab have been high-profile examples of species expanding their ranges to follow their thermal preferences. While species like the American lobster have shown declines at their southern range near Long Island Sound, with much doubt whether they will recover under the present temperature trends. The recent regime shift in the physical oceanography has also shown to be a catalyst for biological shifts as well (@meyer-gutbrod2021; @perretti2017).

While these examples show that species can respond to changes in the physical environment around them through movement & behavior, research elsewhere suggests that physiological responses integrated across species will manifest as changes in community size structure.

## Fishing Impacts on Size/Species Composition

**NEED fishing impacts on size structure references**

### Species Trends in the Northeast Atlantic Shelf

The continental shelf groundfish community in Northwest Atlantic has changed dramatically over the last century. Stocks that supported international fishing effort collapsed, and recovery efforts fell short of their objectives. Research on Georges Bank estimated that biomass more than halved in the 1960's (pre-dating federal monitoring efforts), and noted a species replacement of commercial groundfish target species by skate and dogfish (@fogarty1998). Industrial fishing is inherently size-selective, with larger individuals selectively removed from the population. This has an immediate impact on the community size-distribution with additional impacts on the future population as well. Larger individuals have a greater impact on population recovery, capable of holding more (and often of higher quality) eggs. Size-based harvest in fisheries has been shown to create selective pressures that promote characteristics of early maturation at smaller sizes **add_citation**.

**reference gb spectra early work**

### Purpose

With the understanding that populations depend on the health of their ecosystems, there is a need to have community-wide metrics to effectively understand and manage marine resources. Size based indices are metrics that can be estimated from the information that has historically been available from long-term survey efforts. These indices have been shown to be sensitive to the impacts of fishing, but should also capture environmentally driven stress as well. We estimated size spectrum relationships as SBI's for the groundfish populations for each sub-region of the Northeast US continental shelf. In the case of the NW Atlantic sustained increases in temperature should have a physiological impact on the community size structure.

This leads to our second hypothesis:

> H2. Warming alters the community through the direct influence of temperature on metabolism, growth, and population productivity.

# Methods

## Fish Data Source and Processing

```{r}
#| label: load-survdat

# 1. Biological data used as input
withr::with_dir(rprojroot::find_root('_targets.R'), 
                tar_load(catch_1g_labelled))   

# rename and format
catch_size_bins <- catch_1g_labelled

# Get total number of species
n_species <- length(unique(catch_size_bins$comname))
```

Data on the biomass, abundance, and size of fish on the Northeast U.S. Shelf were collected as part of the Northeast Fisheries Science Center's bottom trawl survey ( @grosslein1969, @Azarovits1981, @politis2014 ) This survey is conducted from Cape Hatteras, North Carolina to the Gulf of Maine each year in the spring and in the fall.  The survey follows a stratified random sampling design, with strata defined based on depth, bottom habitat, and latitude. Trawls are performed for a fixed duration at each station, reporting aggregate abundance and biomass for all species caught, and measuring individual lengths and weights for the catch of each species or a sub-sample if that catch is large. Correction factors were applied to aggregate species abundance and biomass to account for changes in vessels, gear, and doors used in the survey over time ( @sissenwine1978, @byrne1991, @miller2010 ) However, abundance and biomass at length needed to be estimated after these aggregate corrections. As such, abundance at length for each species was adjusted to match the corrected aggregate species abundance at each station, such that for each species, the sum of the resulting estimated abundance numbers across each length is equal to the corrected aggregate abundance.

```{r}
#| label: make region map

# Get region polygons:

# Get the paths to the shapefiles used to mask
trawl_paths <- gmRi::get_timeseries_paths(region_group = "nmfs_trawl_regions", box_location = "cloudstorage")

# Polygons for each region
all_poly <- read_sf(trawl_paths[["inuse_strata"]][["shape_path"]]) 

# Make the Map
ggplot() +
  geom_sf(data = us_poly) +
  geom_sf(data = canada) +
  geom_sf(data = all_poly, aes(fill = survey_area), alpha = 0.8) +
  coord_sf(xlim = c(-76.4, -64.4), ylim = c(35, 45.5), expand = F) +
  scale_fill_gmri() +
  theme_bw() +
  map_theme(legend.position = c(0.6, 0.125), 
            legend.background = element_rect(fill = "white")) +
  guides(fill = guide_legend(title = "", nrow = 2)) +
  labs(caption = "Regions assigned to individual survey strata.")
```

### Community Composition & Functional Groups

Analyses were performed using 68 species. These species were selected based on the availability of published weight-at-length relationships (Wigley et al. 2003) and represented 98.98% of the total biomass caught in the survey. Each species was assigned to a functional group based on life history and geography using the definitions of (@hare2016). Functional groups included coastal fish, diadromous fish, elasmobranch, groundfish, and pelagic fish (@tbl-functional-groups). Six species with available length-weight details did not have a functional group designation, these species were designated as **reef species**. Exploratory analyses showed that the pelagic species biomass was low in all regions, and is unlikely to be representative of true biomass trends due to gear selectivity.

```{r}
#| label: biomass-group-summaries

# This cell provides the group summaries for how much biomass exists in each size/functional group

# The handmade weight and length bins
# # Do some grouping by year survey area and functional group
# fgroup_area <- get_group_summaries(data_in = catch_size_bins, Year, survey_area, hare_group)
# fgroup_all <- get_group_summaries(data_in = catch_size_bins, Year, hare_group)
# fgroup_all <- fgroup_all %>% map(~mutate(.x, survey_area = "Northeast Shelf, Full"))


# Actual log10 bins
# Do some grouping by year survey area and functional group
fgroup_area <- log2_bin_summaries(
  data_in = catch_size_bins, Year, survey_area, hare_group)
fgroup_all  <- log2_bin_summaries(data_in = catch_size_bins, Year, hare_group)
fgroup_all  <- fgroup_all %>% map(~mutate(.x, survey_area = "Northeast Shelf, Full"))


# Join the overall values back in
fgroup_area <- map2(fgroup_area, fgroup_all, bind_rows)

```

Published length-weight relationships ( Wigley et al. 2003 ) were used to convert from length data, available for all individuals, into their corresponding biomass-at-length. To account for differences in sampling effort among survey strata, all corrected abundance-at-length data were area-stratified. Area-stratified biomass-at-length values were then computed as the product of area-stratified abundance-at-length and estimated weight-at-length. All analyses were performed using area-stratified abundances and their associated area-stratified biomass estimates, hereby referred to as simply abundance-at-length & biomass-at-length

<!--# Everything above here in methods describes the data source, and its pre-processing, Everything below details what is done with it -->

### 

## Community Composition Metrics

Our analyses used all data collected during the spring and fall surveys from 1970-2019. Data were grouped using survey-design strata into four sub-regions: Gulf of Maine, Georges Bank, Southern New England, Mid-Atlantic Bight. These sub-regions have been widely used in regional ecological studies (e.g., ). For each region, we developed several time series indicators:

1.  Community Composition Metrics (regional abundance and biomass by functional group)

2.  Average length and weight of the aggregate community and each functional group

3.  Exponent of the Individual Size Distribution Spectra

### 

### Body Size Changes

The average body length (cm) and body weight (kg) was calculated for all catch for each region and within each functional group. These averages were weighted by abundance-at-length. Data for body size trends were not truncated using any minimum or maximum size and reflect all available catch data for the `r n_species`, for which biomass-at-length could be estimated.

```{r}
#| label: body size data

# Load the average body size data
withr::with_dir(rprojroot::find_root('_targets.R'), 
                tar_load(mean_sizes_ss_groups))   

# Grouped on year and region
regional_size <- mean_sizes_ss_groups %>% 
  filter(`group ID` == "single years * region") %>% 
  mutate(Year = as.numeric(Year),
         survey_area = factor(survey_area, levels = area_levels))


# Grouped on year, region, & functional group
functional_group_size <- mean_sizes_ss_groups %>% 
  filter(`group ID` == "single years * region * functional group") %>% 
  mutate(Year = as.numeric(Year),
         survey_area = factor(survey_area, 
                              levels = area_levels))

```



### Normalized Abundance Spectra

Normalized abundance spectra were estimated following the methodology as described in @sprules. When fitting the normalized size spectra, biomass at length data was binned into equal spaced intervals (1) on a $log_{2}$ scale, with abundances totaled across all species. To normalize the spectra, the stratified abundance within each bin was then divided by the bin-width to account for the increasing bin-widths, a consequence of the log scale. Normalized abundance spectra were fit for each year and for each region independently, and for each year across all strata, using a regression of log2( area-stratified abundance, normalized by log2 bin widths ) and the log2( body-size bin midpoints ).

```{r}
#| label: load the size spectrum indices
withr::with_dir(rprojroot::find_root('_targets.R'), 
                tar_load(size_spectrum_indices))   


# Grab SS Groups we care about
region_indices <- size_spectrum_indices  %>% 
  filter(`group ID` == "single years * region") %>% 
  mutate(yr = as.numeric(as.character(Year)),
         survey_area = factor(survey_area, 
                              levels = area_levels),
         sig_fit = ifelse(log2_sig_strat < 0.05, "Significant", "Non-Significant"))

```

### 

### Individual Size Distribution Analysis

Length for individuals in the catch data are measured to the nearest cm, with smaller specimens measured to the nearest millimeter. Because individual biomass is estimated from those length measurements, there is a range of possible body mass values between the cm & mm increments. The relationship between length and mass in fishes is exponential and taxon specific, so biases resulting from using only the lower or upper end of those ranges is different for each taxon and increases for larger taxa. To account for this and reduce biases, we used the extended likelihood method (MLEbin) of @edwards2020.

$$
\begin{align*}
f(x) = \frac{ (\lambda + 1)x^{\lambda} }{ x^{\lambda+1}_{max} - x^{\lambda+1}_{min} }~~~~~~\lambda\neq1,
\\
\\
f(x) = \frac{1}{logx_{max} - logx_{min}}~~~~~~\lambda=1
\end{align*}
$$

The individual size distribution relationship was estimated for each survey region and for all years from 1970 to 2019. A minimum biomass of 1g was used for the lower bound ($x_{min}$) and a maximum biomass of 10kg was used as an upper bound ($x_{max}$) for the ISD's bounded power law probability density function @eq-ISD-pdf, where $x$ is body mass & $\lambda$ is the scaling exponent of the ISD. The biomass within these limits constitutes 97.83% of all estimated biomass for the `r n_species` used in this study. This truncation accounts for poor gear selectivity at the smallest and largest size ranges and imposes shared limits to the size distribution we'd expect to sample across these different areas. Exponents of size spectra were calculated using code modified from the sizeSpectra package in R (@edwards2017; @edwards2020).

## Drivers of Size Distribution Changes

Changes in size spectra were correlated against several hypothesized driving forces related to both environmental regimes and anthropogenic disturbances. Potential large-scale environmental drivers included regional SST anomalies & the larger-scale impact of the Gulf Stream Index (GSI). Fishing pressure represents the primary top-down anthropogenic driver in the region, and was investigated using the aggregate regional commercial landings data.

### Gulf Stream Index

Gulf stream index (GSI) values were obtained from the ecodata package in R (Bastille & Hardison 2018) . This package supplies GSI data at monthly intervals following the methodology of @pérez-hernández2014 and @joyce2019 , using as sea level height anomaly data from the Copernicus Marine Environment Monitoring Service

```{r}
#| label: gsi-data-prep
#| eval: true
####  2. Climate Drivers  ####

# From ecodata:

####  GSI

# Gulf Stream Index
gsi <- ecodata::gsi %>% 
  mutate(
    Time = str_pad(as.character(Time), width = 7, side = "right", pad = "0"),
    year = as.numeric(str_sub(Time, 1, 4)),
    month = str_sub(Time, -2, -1),
    Time = as.Date(str_c(year, month, "01", sep = "-")))
```

### Sea Surface Temperature Data

Global Sea surface temperature data was obtained via NOAA's optimally interpolated SST analysis (OISSTv2), providing daily temperature values at a 0.25° latitude x 0.25° longitude resolution (Reynolds et al. 2007). A daily climatology for every 0.25° pixel in the global data set was created using average daily temperatures spanning the period of 1982-2011. Daily anomalies were then computed as the difference between observed temperatures and the daily climatological average. OISSTv2 data used in these analyses were provided by the NOAA PSL, Boulder, Colorado, USA from their website at [https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.html](https://www.google.com/url?q=https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.html&sa=D&source=docs&ust=1675370947039361&usg=AOvVaw1ShFZYjArUTbnJyV_pkcaK).

Sea surface temperature data was regionally averaged to match the survey regions from the age-at-length data. SST anomalies were averaged by year for each region and over the entire sampling region to produce daily time series. These time series were then processed into annual timeseries of surface temperatures and anomalies. All region-averaging was done with area-weighting of the latitude/longitude grid cells to account for differences in cell-size in the OISSTv2 data.

```{r}
#| label: load temperature data

# Load the regional temperatures from {targets}
withr::with_dir(rprojroot::find_root('_targets.R'), 
                tar_load(regional_oisst))


# Get regional averages
temp_regimes <- regional_oisst %>% 
  filter(yr > 1981) %>% 
  mutate(
    regime = ifelse(sst_anom > 0, "hot", "cold"),
    survey_area = ifelse(
      survey_area == "all", 
      "Northeast Shelf, Full", 
      survey_area),
    survey_area = factor(survey_area, levels = area_levels)) 



```

### Commercial Fishing

Fishing pressure in the region was indexed using state and federal commercial fishing landings.  These data were obtained from the Greater Atlantic Regional Fisheries Office (GARFO) for statistical areas that are routinely used for fisheries reporting and management (Figure X).  Individual statistical areas were aggregated into regions that closely align with the survey areas we defined for the size spectra analyses (Figure X).

```{r}
#| label: garfo-data-prep
#| eval: true

#### 1. GARFO Landings  ####

# Load the GARFO landings data:
res_path <- gmRi::cs_path("res")

# Landings of finfish* sheet 5
landings <- read_xlsx(
  path = str_c(res_path, "GARFO_landings/KMills_landings by area 1964-2021_JUN 2022.xlsx"), sheet = 5) %>% 
  rename_all(tolower)


# Shapefiles for the fisheries stat zones
stat_zones <- read_sf(str_c(res_path, "Shapefiles/Statistical_Areas/Statistical_Areas_2010_withNames.shp"))


# Make a list of zones to roughly match the survey areas:
fish_zones <- list(
  "Gulf of Maine" = c(511:515, 464, 465),
  "Georges Bank" = c(521, 522, 525, 561, 562),
  "Southern New England" = c(611, 612, 613, 616, 526, 537, 538, 539),
  "Mid-Atlantic Bight" = c(614:615, 621, 622, 625, 626, 631, 632))


# Trim out what we don't need and label them
stat_zones <- stat_zones %>% 
  mutate(
    survey_area = case_when(
      Id %in% fish_zones$"Gulf of Maine" ~ "Gulf of Maine",
      Id %in% fish_zones$"Georges Bank" ~ "Georges Bank",
      Id %in% fish_zones$"Southern New England" ~ "Southern New England",
      Id %in% fish_zones$"Mid-Atlantic Bight" ~ "Mid-Atlantic Bight")) %>% 
  filter(survey_area %in% c(
    "Georges Bank", "Gulf of Maine", "Southern New England", "Mid-Atlantic Bight"))
  


# Add the labels into the landings data and remove what we don't need there:
landings <- landings %>% 
  mutate(
    survey_area = case_when(
      `stat area` %in% fish_zones$"Gulf of Maine" ~ "Gulf of Maine",
      `stat area` %in% fish_zones$"Georges Bank" ~ "Georges Bank",
      `stat area` %in% fish_zones$"Southern New England" ~ "Southern New England",
      `stat area` %in% fish_zones$"Mid-Atlantic Bight" ~ "Mid-Atlantic Bight")) %>% 
  filter(survey_area %in% c("Georges Bank", "Gulf of Maine", "Southern New England", "Mid-Atlantic Bight")) %>% 
  mutate(survey_area = factor(survey_area, area_levels_long))



# Get Summaries
landings_summ <- landings %>% 
  drop_na() %>% 
  rename(
    "weight_lb" = `landed lbs`,
    "live_lb" = `live lbs`) %>% 
  drop_na() %>% 
  group_by(year, survey_area) %>% 
  summarise( 
    across(
      .cols = c(value, weight_lb, live_lb), 
      .fns = list(mean = mean, 
                  total = sum), 
      .names = "{.fn}_{.col}"), 
    .groups = "drop")  


# Scale the landings to create an index by area
landings_summ <- landings_summ %>% 
  group_by(survey_area) %>% 
  mutate(total_wt_z = base::scale(total_weight_lb)) %>% 
  ungroup()
```

```{r}
#| label: drivers-table


####  Build a Common form for all Drivers:   ####
# Format: year, area, var, value


# table to join for swapping shorthand for long-hand names
area_df <- data.frame(
  area = c("Scotian Shelf", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight", "All"),
  survey_area = c("SS", "GoM", "GB", "SNE", "MAB", "Northeast Shelf, Full"))



#--------------------
#### Reshape all the inputs so that we can bind them together in one dataframe


# 1.Climate Features

# Join the climate modes together
clim_drivers <- bind_rows(
  list(
    # This is a list because other driverse went here at one point
    gsi
  )) %>% filter(EPU == "All")
  

# Put climate drivers on annual schedule
clim_idx <- clim_drivers %>% 
  group_by(year, area = EPU, var = Var) %>% 
  summarise(value = mean(Value, na.rm = T),
            .groups = "drop")

#--------------------
# 2. GARFO landings
landings_idx <- landings_summ %>% 
  group_by(year, area = survey_area) %>% 
  summarise(value = mean(total_wt_z, na.rm = T),
            .groups = "drop") %>% 
  mutate(var = "landings")


#--------------------
# 3. Temperature
sst_idx <- temp_regimes %>% 
  left_join(area_df, by = "survey_area") %>% 
  select(year = yr, area, value = sst_anom) %>% 
  mutate(var = "sst_anom")


#--------------------
#  4. Size Spectrum Slopes & Intercepts
ss_idx <- region_indices %>% 
  select(
    year = Year, 
    survey_area, 
    spectra_slope = log2_slope_strat, 
    #spectra_int = log2_int_strat,
    isd_slope = b) %>% 
  pivot_longer(
    # cols = c(spectra_slope, spectra_int, isd_slope), 
    cols = c(spectra_slope, isd_slope), 
    names_to = "var", 
    values_to = "value") %>% 
  mutate(year = as.numeric(year)) %>% 
  left_join(area_df, by = "survey_area") %>% 
  select(-survey_area)

#--------------------
# All Metrics Together
all_drivers <- bind_rows(
  list(
    clim_idx,
    landings_idx,
    sst_idx,
    ss_idx
  )
)

```


```{r}
#| label: driver-scaling

# In this chunk the drivers get reshaped to scale
# Scaling should cover the period of analysis 1982-2019
# SST and GSI (all index/anomaly fields) are not scaled


# # Reshape the Drivers Wide to allow independent scaling
# # Scale the columns independently using the variance in all years of data
# drivers_scaled <- all_drivers %>% 
#   mutate(id = str_replace_all(str_c(area, "_", var), " ", "_")) %>% 
#     select(-c(area, var)) %>% 
#     pivot_wider(names_from = id, values_from = value) %>% 
#   column_to_rownames(var = "year") %>% 
#   scale() %>% 
#   as.data.frame()

# Put the data in a wide form for scaling:
drivers_wide <- all_drivers %>% 
  filter(year >= 1982,
         year <= 2019) %>% 
  mutate(id = str_replace_all(str_c(area, "_", var), " ", "_")) %>% 
    select(-c(area, var)) %>% 
    pivot_wider(names_from = id, values_from = value) %>% 
  column_to_rownames(var = "year")


# Scale the landings, not SST or GSI
to_scale   <- select(drivers_wide, ends_with("landings")) %>% scale() %>% as.data.frame()
dont_scale <- select(drivers_wide, -ends_with("landings"))

# Join back together
drivers_scaled <- bind_cols(dont_scale, to_scale) %>% 
  rownames_to_column(var = "year")



# -------------------------------------------------
# Idea: Add functional_group_sizes into the correlation chart:

# Reshape the size information so we can combine it with the drivers:
# Values are scaled here
sizes_wide <- functional_group_size %>% 
  select(year = Year, survey_area, hare_group, avg_length = mean_len_cm, avg_weight = mean_wt_kg) %>% 
  pivot_longer(names_to = "var", 
               values_to = "value", 
               cols = c(avg_length, avg_weight)) %>% 
  inner_join(area_df, by = "survey_area") %>% #distinct(area)
  select(-survey_area) %>% 
  mutate(id = str_replace_all(str_c(area, "_", hare_group, "_", var), " ", "_")) %>% 
    select(-c(area, var, hare_group)) %>% 
    pivot_wider(names_from = id, values_from = value, values_fn = {mean}
                ) %>% 
  column_to_rownames(var = "year") %>% 
  scale() %>% 
  as.data.frame()



# Join regional size changes to the scaled drivers
drivers_and_sizes <- drivers_scaled %>% 
  left_join(rownames_to_column(sizes_wide, var = "year"), by = "year") %>%
  column_to_rownames(var = "year")

```


```{r}
#| label: set-driver-periods


# Pull different years into a list
driver_periods <- list(
  "1980s" = c(1982:1989),
  "1990s" = c(1990:1999),
  "2000s" = c(2000:2009),
  "2010s" = c(2010:2019),
  "1982-2019" = c(1982:2019)
  )


# Split them out into a list that contains data for each period
# Takes the years in driver_periods and pulls the rows that correspond
driver_period_list <- map(driver_periods, function(yrs){
  period_matrix <- drivers_and_sizes[rownames(drivers_and_sizes) %in% yrs, ] %>% drop_na()
  })

```

## Driver Impacts

Cross correlation function estimates were used to look at correlation changes at one year intervals to investigate potential lag effects between the hypothesized drivers and the community's exponent of size spectra. CCF estimates were performed between the dependent variables of the size spectra slope, size spectra intercept, and individual size distribution exponent with the independent variables of the annual gulf stream index, and the corresponding regional commercial landings and sea surface temperature anomalies.

```{r}
#| label: ccf-data
#| fig-height: 8
#| fig-width: 8

# Run CCF function to get lagged correlations,
# Put that data in one table to plot in a better comparison form


#------------ Processing Data for CCF. ------------------------
# Scale Everything Before CCF,
# Reshape the Drivers Wide to scale
# Scale the columns independently using the variance in all years, but within an area
# Are the columns scaling independently or together?
# Confirmed independently scaling


# Drivers scaled trimmed is a matrix of all the different timeseries
# Columns that end with slope or intercept are pulled out to get a column of the response variables (slope features)
# Columns that don't contain "spectra" or "year" capture the remaining features
# Next we pull out the region information for each
driver_ccf_prep <- drivers_scaled %>% 
  select(-All_sst_anom) %>% 
  #rownames_to_column(var = "year") %>% 
  pivot_longer(names_to = "spectra_param", 
               values_to = "spectra_values", 
               cols = ends_with("slope") | ends_with("int")) %>% 
  pivot_longer(names_to = "driver_var", 
               values_to = "driver_values", 
               cols = -matches("spectra|year")) %>% 
  mutate(
    # C. These are the areas associated with the Spectra Features
    spectra_area = case_when(
      str_detect(spectra_param, "All")          ~ "All",
      str_detect(spectra_param, "Georges")      ~ "Georges Bank",
      str_detect(spectra_param, "Gulf")         ~ "Gulf of Maine",
      str_detect(spectra_param, "Southern")     ~ "Southern New England",
      str_detect(spectra_param, "Mid-Atlantic") ~ "Mid-Atlantic Bight"),
    spectra_area = factor(spectra_area, 
      levels = c("All", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight")),
    
    # D. These are the areas of the drivers
    driver_area = case_when(
      str_detect(driver_var, "All")          ~ "All",
      str_detect(driver_var, "Georges")      ~ "Georges Bank",
      str_detect(driver_var, "Gulf")         ~ "Gulf of Maine",
      str_detect(driver_var, "Southern")     ~ "Southern New England",
      str_detect(driver_var, "Mid-Atlantic") ~ "Mid-Atlantic Bight"),
    driver_area = factor(driver_area, 
      levels = c("All", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight"))) 



# Filter it out so there is only cases where the driver area matches
driver_ccf_prep <- driver_ccf_prep %>% 
  filter(driver_area == spectra_area | driver_area == "All") %>% 
  arrange(year, driver_var)



# Walk through each xvariable, and see how it correlates with each yvar
# split on xvar
# then split on yvars
# make sure order is good
ccf_relationships <- driver_ccf_prep %>% 
  drop_na() %>% 
  split(.$spectra_param) %>% 
    map_dfr(function(x_param){
      x_param %>% 
        split(.$driver_var) %>% 
          map_dfr(function(driver_y_data){
            ccf_df <- get_ccf_vector(
              x = driver_y_data$spectra_values,
              y = driver_y_data$driver_values
            )
          }, .id = "driver_var")
  }, .id = "spectra_param")




# Build back out the labels for plotting
ccf_plot_data <- ccf_relationships %>% 
  mutate(
    # Flag what the driver type was
    driver_type = case_when(
      str_detect(driver_var, "landings") ~ "Commercial Landings",
      str_detect(driver_var, "sst") ~ "SST",
      str_detect(driver_var, "index") ~ "GSI",
      TRUE ~ "Missed Something"),
    param_feature = case_when(
      # Flag what the Size Distribution Parameter was
      str_detect(spectra_param, "int") ~ "Spectra Intercept",
      str_detect(spectra_param, "isd") ~ "ISD Exponent",
      str_detect(spectra_param, "slope") ~ "Spectra Slope",
      TRUE ~ "Missed Something"),
    spectra_region = case_when(
      # Flag what region the driver was coming from
      str_detect(spectra_param, "All") ~ "All",
      str_detect(spectra_param, "Georges") ~ "Georges Bank",
      str_detect(spectra_param, "Gulf") ~ "Gulf of Maine",
      str_detect(spectra_param, "Southern") ~ "Southern New England",
      str_detect(spectra_param, "Mid-Atlantic") ~ "Mid-Atlantic Bight"),
    # Flag when it crosses threshold
    sig_flag = ifelse(acf < signeg | acf > sigpos, T, F),
    # Set Factor Levels
    spectra_region = factor(spectra_region, levels = c("All", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight")))


# Limit to one Response:
ccf_plot_data <- filter(ccf_plot_data, param_feature == "ISD Exponent")


# Plot them
ccf_plot_data %>% 
  ggplot(aes(
    x = lag, 
    y = acf, 
    group = driver_var, 
    color = driver_type, 
    fill = driver_type)) +
  geom_col(alpha = 0.65) +
  geom_col(
    data = filter(ccf_plot_data, sig_flag),
    aes(lag, acf, fill = driver_type,  group = driver_var), color = "black") +
  geom_text(
    data = filter(ccf_plot_data, sig_flag),
    aes(lag, y = 0, label = lag, vjust = ifelse(acf < 0, 1.5,-1.5)), color = "black") +
  geom_line(aes(x = lag, y = sigpos), linetype = 2, color = "gray25") +
  geom_line(aes(x = lag, y = signeg), linetype = 2, color = "gray25") +
  geom_hline(yintercept = 0, color = "gray25", linewidth = 1) +
  geom_vline(xintercept = 0, color = "black", linewidth = 1) +
  scale_color_gmri() +
  scale_fill_gmri() +
  facet_grid(spectra_region~driver_type*param_feature) +
  scale_x_continuous(limits = c(-15, 15), breaks = seq(from = -15, to = 12, by = 1)) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 8)) +
  labs(x = "Driver Lag", y = "Correlation (CCF)", 
       fill = "Driver Type:", color = "Driver Type:",
       caption = "All predictors scaled over year range of 1982-2019. Lags exceeding significance threshold outlined and numbered.")

```

Following exploratory analyses with CCF functions, regression analyses were performed independently for each region. Candidate models were evaluated using AIC to determine the most parsimonious models for each regional system. Potential drivers included in the models were the annual Gulf Stream Index, the annual regionally averaged SST anomaly value, and the scaled annual commercial landings from that region. Gulf stream index and SST anomalies were not standardized prior to use in model fitting and evaluation as the former is already an index value and the latter is scaled to a thirty-year climatological reference period.

```{r}
#| label: regression-analysis

# Get the drivers organized, only use data from 1982 onward


# Put the data in a wide form for scaling:
drivers_wide <- all_drivers %>% 
  filter(year >= 1982,
         year <= 2019) %>% 
  mutate(id = str_replace_all(str_c(area, "_", var), " ", "_")) %>% 
    select(-c(area, var)) %>% 
    pivot_wider(names_from = id, values_from = value) %>% 
  column_to_rownames(var = "year")


# Scale the landings, not SST or GSI
to_scale   <- select(drivers_wide, ends_with("landings")) %>% scale() %>% as.data.frame()
dont_scale <- select(drivers_wide, -ends_with("landings"))

# Join back together
regression_df <- bind_cols(dont_scale, to_scale) 

# Build back out the metadata:
# Identify what region is associated with both: spectrum values & driver values
regression_df <- regression_df %>% 
  rownames_to_column(var = "year") %>% 
  pivot_longer(names_to = "spectra_param", 
               values_to = "spectra_values", 
               cols = ends_with("slope") | ends_with("int")) %>% 
  pivot_longer(names_to = "driver_var", 
               values_to = "driver_values", 
               cols = -matches("spectra|year")) %>% 
  mutate(
    year = as.numeric(year),
    # A. Flag what the driver type was
    driver_type = case_when(
      str_detect(driver_var, "landings") ~ "Commercial Landings",
      str_detect(driver_var, "sst") ~ "SST",
      str_detect(driver_var, "index") ~ "GSI",
      TRUE ~ "Missed Something"),
    # B. Flag what the spectrum feature was
    param_feature = case_when(
      # Flag what the Size Distribution Parameter was
      str_detect(spectra_param, "int") ~ "Spectra Intercept",
      str_detect(spectra_param, "isd") ~ "ISD Exponent",
      str_detect(spectra_param, "slope") ~ "Spectra Slope",
      TRUE ~ "Missed Something"),
    # C. These are the areas associated with the Spectra Features
    spectra_area = case_when(
      str_detect(spectra_param, "All") ~ "All",
      str_detect(spectra_param, "Georges") ~ "Georges Bank",
      str_detect(spectra_param, "Gulf") ~ "Gulf of Maine",
      str_detect(spectra_param, "Southern") ~ "Southern New England",
      str_detect(spectra_param, "Mid-Atlantic") ~ "Mid-Atlantic Bight"),
    spectra_area = factor(spectra_area, levels = c("All", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight")),
    
    # D. These are the areas of the drivers
    driver_area = case_when(
      str_detect(driver_var, "All") ~ "All",
      str_detect(driver_var, "Georges") ~ "Georges Bank",
      str_detect(driver_var, "Gulf") ~ "Gulf of Maine",
      str_detect(driver_var, "Southern") ~ "Southern New England",
      str_detect(driver_var, "Mid-Atlantic") ~ "Mid-Atlantic Bight"),
    driver_area = factor(driver_area, levels = c("All", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight")),
    
    # E. Add Decade in
    decade = floor_decade(year),
    regime = ifelse(year < 2010, "regime_1", "regime_2" )) 


# Filter it so predictors are only within matching regions, 
# or just overarching (GSI)
regression_df <- regression_df %>% 
  filter(driver_area == spectra_area | driver_type == "GSI") %>% 
  arrange(year, driver_var)




# Last wrangling:
# columns for: year, slope, isd, spectra region, drver_region, GSI, sst_anom, landings
# GSI is so annoying here, just rejoin it without the area column
regression_df <- regression_df %>% 
  select(-c(spectra_param, driver_var)) %>% 
  pivot_wider(names_from = "driver_type", values_from = "driver_values") %>% 
  pivot_wider(names_from = "param_feature", values_from = "spectra_values") %>% 
  select(-GSI) %>% 
  filter(driver_area != "All") %>% 
  left_join( select(clim_idx, year, GSI = value), by = c("year") ) %>% 
  mutate(spectra_area = fct_drop(spectra_area))




```

```{r}
#| label: regression-trees
#| eval: false

library(gbm)
set.seed(123)


#Not enough data
gbm.fit <- gbm(
  formula = `ISD Exponent` ~ SST + decade + `Commercial Landings` + GSI + driver_area,
  distribution = "gaussian",
  data = regression_df,
  interaction.depth = 1,
  shrinkage = 0.001,
  cv.folds = 5,
  n.cores = NULL, # will use all cores by default
  verbose = FALSE
)

```

# Results

### Community Abundance

Stratified abundance was highest in the Gulf of Maine, and decreased across regions moving from North to South. Abundance across all body sizes remained relatively stable in all four regions until the 1990's. At this time abundance in the Gulf of Maine began to steadily rise. This increase in abundance reversed briefly from 2005-2010, but resumed and continued to rise until its peak in 2016. Georges Bank abundances remained low and stable until after 2008, when numbers rapidly increased through 2014, before quickly falling back to numbers slightly above normal by the end of the decade. Abundances in Southern New England experienced higher inter-annual changes in abundance across all years. This area saw a less dramatic rise and fall that began just before 2007, again falling back to earlier levels by the end of the decade. The Mid Atlantic Bight displayed the most inter-annual variability and had relatively consistent abundances throughout, with no major periods of abundance growth or decline.

```{r}
#| label: abundance distributions
#| fig.width: 8
#| fig-height: 8

# panel plot the biomass by body size
fgroup_area$weight_bins <- fgroup_area$weight_bins %>% 
  mutate(strat_abund_mill = wtbin_strat_abund / 1e6,
         strat_lwbio_mill = wtbin_strat_lw_bio / 1e6,
         survey_area = factor(survey_area,
                              levels = area_levels))




# Just abundance, not by groups
strat_abundance_plot <- fgroup_area$weight_bins %>% 
  group_by(Year, survey_area, hare_group) %>% 
  summarise(abund_millions = sum(strat_abund_mill, na.rm = T),
            .groups = "drop") %>% 
  #filter(survey_area == "GoM") %>% 
  ggplot(aes(Year, abund_millions, fill = hare_group)) +
  geom_area(color = "gray30", linewidth = 0.1) +
  scale_fill_gmri() + 
  scale_x_continuous(expand = expansion(add = c(0,0))) +
  facet_wrap(~survey_area, ncol = 1, scales = "free") + 
  labs(y = "Abundance (millions)",  
       caption = "Estimated Abundance from Area-Stratified Catch Rates",
         x = "",
       fill = "Functional Group") +
  theme(legend.position = "bottom",
        panel.background = element_rect(colour = "black", size=2, fill = NA))



# Just Biomass, by area and functional group
strat_biomass_plot <- fgroup_area$weight_bins %>% 
  group_by(Year, survey_area, hare_group) %>% 
  summarise(abund_millions = sum(strat_lwbio_mill, na.rm = T),
            .groups = "drop") %>% 
  ggplot(aes(Year, abund_millions, fill = hare_group)) +
  geom_area(color = "gray30", linewidth = 0.1) +
  scale_fill_gmri() + 
  scale_x_continuous(expand = expansion(add = c(0,0))) +
  facet_wrap(~survey_area, ncol = 1, scales = "free") + 
  labs(y = "Biomass (million kg)",  
       caption = "Estimated Biomass from Area-Stratified Catch Rates",
       x = "",
       fill = "Functional Group") +
  theme(legend.position = "bottom",
        panel.background = element_rect(colour = "black", size=2, fill = NA))


# Biomass and Abundance Together
(strat_abundance_plot | strat_biomass_plot) + plot_layout(guides = "collect") & theme(legend.position = "bottom")


```

Abundance gains observed in Georges Bank and Gulf of Maine were primarily from groundfish species, with additional growth from diadromous species seen in the Gulf of Maine. Increases in abundance across all areas was mostly attibutable to individuals weighing less than .5kg. With some years driven in large-part by exceptional year-classes in just a handful of species e.g. haddock in Georges Bank. The observed abundance volatility in Southern New England and the Mid-Atlantic Bight conversely was largely the result of changes in abundance in pelagic species, whose abundance varied by several times the magnitude that of the other functional groups.

```{r}
#| label: abundance-by-group
#| fig.height: 8
#| fig.width: 10
#| eval: true


# Plot abundance by region and Group
fgroup_area$weight_bins %>% 
  # ggplot(aes(Year, strat_abund_mill, fill = weight_bin)) +
  ggplot(aes(Year, strat_abund_mill, fill = factor(left_lim))) +
    geom_area(color = "transparent") +
    facet_grid(hare_group~survey_area, scales = "free") +
    scale_fill_gmri() +
    scale_y_continuous(labels = comma_format()) +
    labs(y = "Stratified Abundance (millions)",  
         caption = "Stratified abundances of all functional groups and across regions, with fill colored by individual bodymass.",
         x = "", 
         fill = "Minimum Individual Weight 10^x (g)") +
    theme(legend.position = "bottom",
          axis.text.x = element_text(angle = 45, hjust = 1))

```

### Community Biomass

Similar to abundance, the overall biomass was highest in the two northern regions, the Gulf of Maine and Georges Bank. Roughly half of the biomass sampled in these regions can be attributed to groundfish species, with the second largest contributions coming from elasmobranchs. Within the groundfish biomass, larger individuals \>2kg in particular, declined during the 70's and 80's in these regions, never truly recovering. Beginning in the 2000's there were signs that groundfish abundances were increasing as evidenced by increasing numbers of smaller individuals, however in both regions this trend appears to have reversed by the mid 2010's. Elasmobranch biomass increased steadily throughout the survey time period across all regions, with the exception of southern New England. This area showed large 5-10 year swings in biomass, but no clear long-term trend. Larger elasmobranch were rare in all regions except for a period spanning the late 70's through the early 90's isolated to Georges Bank. Demersal species biomass was highest in the Gulf of Maine, dwarfing their contributions in other regions. Their biomass declined in the 70's, was flat until the late 90's, remaining relatively high until declining in the late 2010's. Pelagic species biomass was low in all regions, and is unlikely to be representative of true biomass trends due to gear selectivity.

```{r}
#| label: bodymass distributions
#| fig.width: 10
#| fig-height: 8

# panel plot the biomass by body size
fgroup_area$weight_bins %>% 
  complete(Year, hare_group, survey_area, left_lim) %>% 
  # ggplot(aes(Year, strat_abund_mill, fill = weight_bin)) +
  ggplot(aes(Year, strat_abund_mill, fill = factor(left_lim))) +
    geom_area(color = "transparent") +
    facet_grid(hare_group~survey_area, scales = "free") +
    scale_fill_gmri() +
    scale_y_continuous(labels = comma_format()) +
    labs(y = "Stratified Total Biomass (million kg)",  
         caption = "Stratified abundances of all functional groups and across regions, with fill colored by individual bodymass.",
         x = "", 
         fill = "Minimum Individual Weight 10^x (g)") +
    theme(legend.position = "bottom",
          axis.text.x = element_text(angle = 45, hjust = 1))


```

### Regional Variation in Species Composition

There was a distinct difference between Northern and Southern regions in the way biomass was distributed among the different functional groups. The largest contributors to biomass in the southern regions (southern New England & mid-Atlantic bight) was the elasmobranch community. While the northern regions (Gulf of Maine & Georges Bank) each had similar quantities of elasmobranch biomasses, there was also a comparable contribution of groundfish and in the Gulf of Maine there was a major component of demersal species as well.

### Body Size Trends

The average fish size in the Gulf of Maine (length and weight) declined the greatest of all regions over our study period. The average individual length was greatest in the 1970's in the 35-40cm range, falling to 28-33cm over the last decade. Body-weight fell dramatically in the 1980's, from around .75kg in the 1970's to .25-.30kg, roughly a third of what it had been. Georges Bank body sizes also declined during the study period, but less dramatically. Both of these Northern regions had brief period in the late 2000's where average length and weight rose, before falling again in the 2010's. The MAB region was the only region to see a long-term increase in both length and weight during the study period. SNE saw no long-term change in length, and a minor decline in average body-weight.

::: panel-tabset
#### All Species

```{r}
#| label: regional-size-trends
#| fig.height: 8
#| fig.weight: 8

# Re-factor
regional_size <- regional_size %>% 
  mutate(
    survey_area = factor(survey_area, area_levels),
    area_copy = survey_area)


# Length plot
avg_len_p <-  regional_size %>% 
    ggplot(aes(Year, mean_len_cm)) +
    geom_line(data = select(regional_size, -survey_area),
              aes(group = area_copy),
              alpha = 0.2, linewidth = 0.5) + 
    geom_line(aes(group = survey_area), 
              linewidth = 1,
              show.legend = F) + 
    scale_color_gmri() +
    facet_wrap(~survey_area, ncol = 1) + 
    labs(title = "Average Length",
         y = "Length (cm)",
         color = "Region")


# Weight plot
avg_wt_p <- regional_size %>% 
  ggplot(aes(Year, mean_wt_kg)) + 
  geom_line(data = select(regional_size, -survey_area),
            aes(group = area_copy),
            alpha = 0.2, linewidth = 0.5) + 
  geom_line(aes(group = survey_area#,
                  #color = area_copy
                ), 
            linewidth = 1,
              show.legend = F) + 
  scale_color_gmri() +
  facet_wrap(~survey_area, ncol = 1) + 
  labs(title = "Average Weight",
       y = "Weight (kg)",
       color = "Region")
  
# Plot side by side
(avg_len_p | avg_wt_p) + plot_annotation(caption = "Changes in the average length and bodymass of fishes in each of the study regions.")

```

#### Functional Group Length

```{r}
#| label: functional-group-length-trends
#| fig.width: 8
#| fig-height: 8


# Refactor areas
functional_group_size <- functional_group_size %>% 
  mutate(survey_area = factor(survey_area, area_levels),
      area_copy = survey_area)



functional_group_size %>% 
  ggplot(aes(Year, mean_len_cm)) +
  geom_line(aes(group = survey_area#,
                  #color = area_copy
                  ), 
              linewidth = 1,
              show.legend = F) + 
    geom_smooth(color = "orange", method = "loess", linewidth = 1) +
    facet_grid(hare_group~survey_area, scales = "free_y") + 
    theme(axis.text = element_text(size = 8)) +
    labs(title = "Average Length",
         y = "Length (cm)",
         color = "Region")


```

#### Functional Group Bodymass

```{r}
#| label: functional-group-weight-trends
#| fig.width: 8
#| fig-height: 8


# Average individual weight
functional_group_size %>% 
  ggplot(aes(Year, mean_wt_kg)) +
  geom_line(aes(group = survey_area), 
              linewidth = 1,
              show.legend = F) + 
    geom_smooth(color = "orange", method = "loess", linewidth = 1) +
    facet_grid(hare_group~survey_area, scales = "free_y") + 
    theme(axis.text = element_text(size = 8)) +
    labs(title = "Average Weight",
         y = "Weight (kg)",
         color = "Region")

```
:::

### Regional Size Spectra

At the start of our time series, back in the 1970's, there was a clear difference in the relative positions of spectra parameters among the different regions. Gulf of Maine and Georges Bank showed the least steep spectra slopes in the earlier time periods with slopes around -1 & -1.1 respectively. The relatively flat slopes in these regions both steepened over time, settling near -1.3 (GoM) and -1.5 (GB). Gulf of Maine experienced much of its decline during the 1980's and 1990's. There was a brief reversal in this trend during the 2000's, but slopes continued to steepen by 2010 and remained steep through 2019. Georges Bank did not experience as rapid of a decline, but experienced a similar long-term steepening. In contrast to the northern regions, SNE and MAB had steeper slopes in the -1.2 to -1.5 territory. The long term pattern for SNE was one of increasing volatility, but not so much a decline. The spectra slope for the MAB was less volatile, but similarly maintained a relatively stable wander around -1.4. By the end of the study period all regions had slopes that were at or near a similar level.




```{r}
#| label: size spectra results


# Make a copy so we can gray out the lines
region_indices <- mutate(region_indices, area_copy = survey_area)


# Plot the Regional Slopes
slope_timeline <- region_indices %>% 
  ggplot(aes(yr, log2_slope_strat, group = survey_area)) +
  geom_line(data = select(region_indices, -survey_area),
            aes(group = area_copy),
            alpha = 0.2, linewidth = 0.5) +
  geom_line(aes(group = survey_area), linewidth = 1) +
  scale_color_gmri() +
  facet_wrap(~survey_area, ncol = 1)  +
  labs(x = "Year", y = "Slope", title = "Normalized Abundance Spectra") 




# Exponent of ISD
isd_timeline <- region_indices %>% 
  ggplot(aes(yr, b, group = survey_area)) +
  geom_line(data = select(region_indices, -survey_area),
            aes(group = area_copy),
            alpha = 0.2, linewidth = 0.5) +
  geom_line(aes(group = survey_area), linewidth = 1) +
  #geom_pointrange(aes(group = survey_area, ymin = confMin, ymax = confMax), size = 0.1) +
  scale_color_gmri() +
  facet_wrap(~survey_area, ncol = 1) +
  labs(x = "Year", y = "b", title = "Exponent of ISD")
```


::: panel-tabset

#### Exponent of Size Spectra

```{r}
#| fig.height: 8
#| fig.weight: 8

isd_timeline + geom_smooth(color = "orange")
```

#### Binned Size Spectra

```{r}
#| fig.height: 8
#| fig.weight: 8

slope_timeline + geom_smooth(color = "orange")
```

#### Contrast

```{r}
#| fig.height: 8
#| fig.weight: 8

# Assemble the figure
slope_timeline +
  scale_y_continuous(limits = c(-1.35, -0.4))  | isd_timeline +
  scale_y_continuous(limits = c(-1.35, -0.4)) 


```

:::



## Size Spectra Drivers

### Drivers of Spectra Shape

**NOTE:** Correlation matrix is computed starting at the year where there are no NA values in any drivers. Currently with SST included that begins the matrix at 1982.



```{r}
#| label: period-correlations-processing

# Run Correlations for Each period and for the entire period
driver_period_correlations <- map(driver_period_list, function(drivers){
  
  #------ debug copy-paste line
  # Step 1. Get the correlation matrix (data is pre-scaled)
corr_matrix <- rcorr(as.matrix(drivers))

# R-squares
rho_df <- as.data.frame(corr_matrix$r) %>%
  rownames_to_column(var = "var1") %>%
  pivot_longer(cols = -var1, 
               names_to = "var2", 
               values_to = "r2")

# Significance Dataframe
sig_df <- as.data.frame(corr_matrix$P) %>%
  rownames_to_column(var = "var1") %>%
  pivot_longer(cols = -var1, 
               names_to = "var2", 
               values_to = "pval")

# Join
corr_both <- left_join(rho_df, sig_df, by = c("var1", "var2")) 


# Step 2. Tidy Up for ggplot
# Make correlation matrix & significance a dataframe for GGplot
# Goal: Flag the different Functional Group size changes somehow


# Mega Cleanup Code Chunk
# This first step adds metadata, but mostly creates a way to filter out correlations to be in the direction we want. no sst as response variables for example
# GOAL: All the drivers in one column, and all the responses in another, order doesn't matter
corr_tidy <- corr_both %>% 
  mutate(
    
    # 1.These flag whether or not the matrix variables in each column are the independent or dependent variables
    is_var1_driver   = ifelse(str_detect(var1, "landings|index|anom|avg"), T, F),
    is_var2_response = ifelse(str_detect(var2, "slope|int"), T, F),
    
    # 2. Column to Flag what type of driver it is:
    driver_type = case_when(
        str_detect(var1, "landings")    ~ "Top-Down",
        str_detect(var1, "stream|anom") ~ "Bottom-Up",
        str_detect(var1, "length")      ~ "Avg. Length",
        str_detect(var1, "weight")      ~ "Avg. Weight"),
    
    # 3. Column to Flag what area the driver data is from
    driver_area = case_when(
        str_detect(var1, "All")          ~ "All",
        str_detect(var1, "Georges")      ~ "Georges Bank",
        str_detect(var1, "Gulf")         ~ "Gulf of Maine",
        str_detect(var1, "Southern")     ~ "Southern New England",
        str_detect(var1, "Mid-Atlantic") ~ "Mid-Atlantic Bight"),
    
    # 4. Set Factor Levels for plot
      driver_area = factor(
        driver_area, 
        levels = c(
          "All", 
          "Gulf of Maine", 
          "Georges Bank", 
          "Southern New England", 
          "Mid-Atlantic Bight")),
      
    
    # 5. Shorthand Names for the Spectra Parameters - Response
    response_short = case_when(
      str_detect(var2, "int")   ~ "Spectra Intercept",
      str_detect(var2, "isd")   ~ "ISD Exponent",
      str_detect(var2, "slope") ~ "Spectra Slope"),
    
    
    # 6. These are the areas of the drivers
    response_area = case_when(
      str_detect(var2, "All")          ~ "All",
      str_detect(var2, "Georges")      ~ "Georges Bank",
      str_detect(var2, "Gulf")         ~ "Gulf of Maine",
      str_detect(var2, "Southern")     ~ "Southern New England",
      str_detect(var2, "Mid-Atlantic") ~ "Mid-Atlantic Bight"),
    
    
    # 7. Set levels for the driver areas
    response_area = factor(
      response_area, 
      levels = c(
        "All", 
        "Gulf of Maine", 
        "Georges Bank", 
        "Southern New England", 
        "Mid-Atlantic Bight")),
    
    # 8. Flag when the areas are the same or not:
    area_match = driver_area == response_area
  )
  

# Now we filter rows out to focus in on:
# 1. one direction driver & response
# 2. response and driver in same area, unless its GSI
corr_tidy_pairs <- corr_tidy %>% 
  filter( 
      is_var2_response,
      is_var1_driver,
      area_match | str_detect(var1, "index")) 
  
  
# From here it is just text formatting for plots:
# Because we are faceting we can have repeated names to simplify things
corr_tidy_pairs <- corr_tidy_pairs %>% 
  mutate(
      var1 = ifelse(str_detect(var1, "landings"), "Commercial_Landings", var1),
      var1 = ifelse(str_detect(var1, "sst"), "SST_Anomalies", var1),
      var1 = ifelse(str_detect(var1, "gulf_stream_index"), "Gulf Stream Index", var1),
      var1 = str_replace_all(var1, "_", " "),
      
      # Make a functional group column
      hare_group = case_when(
        str_detect(var1, "elasmobranch") ~ "Elasmobranch",
        str_detect(var1, "pelagic")      ~ "Pelagic",
        str_detect(var1, "coastal")      ~ "Coastal",
        str_detect(var1, "diadromous")   ~ "Diadromous",
        str_detect(var1, "groundfish")   ~ "Groundfish",
        str_detect(var1, "reef")         ~ "Reef",
        TRUE ~ ""),
      # Add the functional group column ahead of the average length and weight columns
      var1 = ifelse(str_detect(var1, "avg"), hare_group, var1),
      
      # Set The Factor Levels groups
      var1 = factor(
        var1,
        levels = c(
          "Gulf Stream Index",
          "SST Anomalies",
          "Groundfish",
          "Elasmobranch",
          "Coastal",
          "Diadromous",
          "Pelagic",
          "Reef",
          "Commercial Landings")),
      
      # Flag Significance Levels
      signif = ifelse(pval < 0.05, "*", ""))

  return(corr_tidy_pairs)
  
  #------- 

  
  
  
})

```



::: panel-tabset

#### All Years

```{r}
#| label: correlations-all
#| fig-height: 6
#| fig-width: 8

period <- "1982-2019"
plot_period_correlations(driver_period_correlations[[period]], period)
```



#### 1980s

```{r}
#| label: correlations-80s
#| fig-height: 6
#| fig-width: 8

period <- "1980s"
plot_period_correlations(driver_period_correlations[[period]], period)
```


#### 1990's

```{r}
#| label: correlations-90s
#| fig-height: 6
#| fig-width: 8

period <- "1990s"
plot_period_correlations(driver_period_correlations[[period]], period)
```


#### 2000's

```{r}
#| label: correlations-20s
#| fig-height: 6
#| fig-width: 8

period <- "2000s"
plot_period_correlations(driver_period_correlations[[period]], period)
```


#### 2010's

```{r}
#| label: correlations-21s
#| fig-height: 6
#| fig-width: 8

period <- "2010s"
plot_period_correlations(driver_period_correlations[[period]], period)

```



:::


# WORKING HERE


## Exponent of Size Distribution Regressions


::: panel-tabset

### Gulf of Maine

```{r}
library(MuMIn)
library(gtsummary)
library(broom.helpers)
library(car)

# change na. action
options(na.action = "na.fail")

# Pull GOM to test:
gom_df <- filter(regression_df, spectra_area == "Gulf of Maine") 

# fit model with all parameters
all_parms <- lm(`ISD Exponent` ~ SST + GSI + `Commercial Landings` + regime, data = gom_df)

# # Refine the dredge
# # the dredge function fits all combinations
# # of the variables in the all.parms model fit above
# # only allow a maximum of 3 and minimum of 1 parameters in each model
# # fit all models but do not include models with both regime and decade
# # grab best model
results <- dredge(all_parms,# subset= !(regime && decade), 
                  m.lim = c(1,4), fixed = c("(Intercept)"))
# subset
# subset(results, delta == 0)

# The "best" model
bmod <- lm(`ISD Exponent` ~ `Commercial Landings`, data = gom_df)

# A model that makes sense
no_decade_mod <- lm(`ISD Exponent` ~ `Commercial Landings` +  GSI, data = gom_df)


# Use Model Selection Manually with MuMIn
# model.sel(bmod, no_decade_mod)


```

::: panel-tabset

#### Model Summary

```{r}
# output summary - brrom.helpers and gtsummary

# Best Model
tbl1 <- bmod %>% 
  tbl_regression() %>% 
  add_glance_table() %>% 
  bold_labels()

tbl1

# To compare two models:

# # What I would guess to be better
# tbl2 <- no_decade_mod %>% 
#   tbl_regression() %>% 
#   add_glance_table() %>% 
#   bold_labels()
  
  
# tbl_merge(list(tbl1, tbl2)) %>%
#   modify_table_body(~.x %>% arrange(row_type == "glance_statistic"))
```



#### VIF

```{r}
all_parms <- lm(`ISD Exponent` ~ SST + GSI + `Commercial Landings` + regime, 
                data = gom_df)
vif(all_parms) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "Predictor") %>% 
  rename(VIF = ".") %>% 
  gt() 

```


#### Actual vs. Fitted

```{r}
augment(bmod, gom_df ) %>% 
  ggplot(aes(year)) +
  geom_line(aes(y = `ISD Exponent`, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = .fitted, color = "Predicted"), linewidth = 1) +
  scale_color_gmri() +
  labs(color = "Color")
```



#### Residual Autocorrelation

```{r}
plot(residuals(bmod),type="b", main = "Gulf of Maine Residuals"); abline(h=0,lty=3) 
```
 

:::







### Georges Bank

```{r}
# Pull GB to test:
gb_df <- filter(regression_df, spectra_area == "Georges Bank") %>% 
  drop_na()

# fit model with all parameters
all_parms <- lm(`ISD Exponent` ~ SST + GSI + `Commercial Landings` + regime, data = gb_df)

# Refine the dredge
# only allow a maximum of 3 and minimum of 1 parameters in each model
# fit all models but do not include models with both regime and decade
results <- dredge(all_parms, #subset= !(regime && decade), 
                  m.lim = c(1,3), 
                  fixed = c("(Intercept)"))

#grab best model
# subset(results, delta == 0)
bmod <- lm(`ISD Exponent` ~ GSI, data = gb_df)

```


::: panel-tabset

#### Model Summary

```{r}
# output summary - brrom.helpers and gtsummary

# Best Model
tbl1 <- bmod %>% 
  tbl_regression() %>% 
  add_glance_table() %>% 
  bold_labels()

tbl1

```



#### VIF

```{r}
all_parms <- lm(`ISD Exponent` ~ SST + GSI + `Commercial Landings` + regime, 
                data = gb_df)
vif(all_parms) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "Predictor") %>% 
  rename(VIF = ".") %>% 
  gt() 

```


#### Actual vs. Fitted

```{r}
augment(bmod, drop_na(gb_df)) %>% 
  ggplot(aes(year)) +
  geom_line(aes(y = `ISD Exponent`, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = .fitted, color = "Predicted"), linewidth = 1) +
  scale_color_gmri() +
  labs(color = "Color")
```



#### Residual Autocorrelation

```{r}
plot(residuals(bmod),type="b", main = "Gulf of Maine Residuals"); abline(h=0,lty=3) 
```
 

:::

### Southern New England

```{r}
# Pull SNE to test:
sne_df <- filter(regression_df, spectra_area == "Southern New England") %>% 
  drop_na()

# fit model with all parameters
all_parms <- lm(`ISD Exponent` ~ SST + GSI + `Commercial Landings` + decade, data = sne_df)

# Refine the dredge
# only allow a maximum of 3 and minimum of 1 parameters in each model
# fit all models but do not include models with both regime and decade
results <- dredge(all_parms, #subset= !(regime && decade), 
                  m.lim = c(1,3), 
                  fixed = c("(Intercept)"))

#grab best model
# subset(results, delta == 0)
bmod <- lm(`ISD Exponent` ~ `Commercial Landings`, data = sne_df)

```


::: panel-tabset

#### Model Summary

```{r}
# output summary - brrom.helpers and gtsummary

# Best Model
tbl1 <- bmod %>% 
  tbl_regression() %>% 
  add_glance_table() %>% 
  bold_labels()

tbl1

```



#### VIF

```{r}
all_parms <- lm(`ISD Exponent` ~ SST + GSI + `Commercial Landings` + regime, 
                data = sne_df)
vif(all_parms) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "Predictor") %>% 
  rename(VIF = ".") %>% 
  gt() 

```


#### Actual vs. Fitted

```{r}
augment(bmod, sne_df ) %>% 
  ggplot(aes(year)) +
  geom_line(aes(y = `ISD Exponent`, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = .fitted, color = "Predicted"), linewidth = 1) +
  scale_color_gmri() +
  labs(color = "Color")
```



#### Residual Autocorrelation

```{r}
plot(residuals(bmod),type="b", main = "Gulf of Maine Residuals"); abline(h=0,lty=3) 
```
 

:::

### Mid-Atlantic Bight

```{r}
# Pull MAB to test:
mab_df <- filter(regression_df, spectra_area == "Mid-Atlantic Bight") %>% 
  drop_na()

# fit model with all parameters
all_parms <- lm(`ISD Exponent` ~ SST + GSI + `Commercial Landings` + regime, data = mab_df)

# Refine the dredge
# only allow a maximum of 3 and minimum of 1 parameters in each model
# fit all models but do not include models with both regime and decade
results <- dredge(all_parms, #subset= !(regime && decade), 
                  m.lim = c(1,3), 
                  fixed = c("(Intercept)"))

#grab best model
# subset(results, delta == 0)
bmod <- lm(`ISD Exponent` ~ regime, data = mab_df)

```


::: panel-tabset

#### Model Summary

```{r}
# output summary - brrom.helpers and gtsummary

# Best Model
tbl1 <- bmod %>% 
  tbl_regression() %>% 
  add_glance_table() %>% 
  bold_labels()

tbl1

```



#### VIF

```{r}
all_parms <- lm(`ISD Exponent` ~ SST + GSI + `Commercial Landings` + regime, 
                data = mab_df)
vif(all_parms) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "Predictor") %>% 
  rename(VIF = ".") %>% 
  gt() 

```


#### Actual vs. Fitted

```{r}
augment(bmod, mab_df ) %>% 
  ggplot(aes(year)) +
  geom_line(aes(y = `ISD Exponent`, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = .fitted, color = "Predicted"), linewidth = 1) +
  scale_color_gmri() +
  labs(color = "Color")
```



#### Residual Autocorrelation

```{r}
plot(residuals(bmod),type="b", main = "Gulf of Maine Residuals"); abline(h=0,lty=3) 
```
 

:::


:::



### Binned Spectra Summaries

Exploring the same model evaluation steps with AIC gives these important predictors in each area:
 * Gulf of Maine: Landings & Decades.  
 * Georges Bank: GSI & Regime.  
 * Southern New England: SST & Decades.  
 * Mid-Atlantic Bight: Decades.  

```{r}
#| eval: false


# Pull GOM to test:
slope_df <- filter(regression_df, spectra_area == "Mid-Atlantic Bight") %>% drop_na()

# fit model with all parameters
all_parms_binned <- lm(`Spectra Slope` ~ SST * GSI + `Commercial Landings`  + regime, 
                       data = slope_df)
results <- dredge(all_parms_binned, 
                  #subset= !(regime && decade), 
                  m.lim = c(1,3), 
                  fixed = c("(Intercept)"))

#grab best model
subset(results, delta == 0)
slope_bmod <- lm(`Spectra Slope` ~ decade + `Commercial Landings`, data = slope_df)

```










## Normalized Size Spectra Regressions



# Discussion

-   Top-down and bottom up influences on both carrying capacity (intercept) and transfer efficiency (slope)

Some of the major drivers suggested here operate on both, but to varying degrees. Here are some potential mechanisms:

**Literature suggests:** - Intercept (a proxy for productivity and carrying capacity) is primarily determined by bottom up features like: nutrient availability, temperature

-   Slope (a measure of energy transfer efficiency and static biomass distribution) has been shown to be sensitive to the physical removal of species through fishing.

**Temperature Mechanisms:** - Temperature's impact on growth via genetic plasticity impacts both the available biomass at the primary producer level (impacting ecosystem carrying capacity), as well as the Linf of larger species (recruitment rate). - Temperature also impacts the efficiency of energy (as biomass) being transferred between individuals via predator & prey interactions. More energy per-capita is expended in the form of increased metabolic rates and/or behavioral changes. This metabolic tax should steepen the spectrum slope by removing available energy at a system wide level. - Temperature impacts behavior via physiological impacts on metabolism and foraging rates as well as through the avoidance of temperature stresses.

## Separating Complimentary Forces Impacting Growth

The data used for this analysis was collected as part of a survey program which began out of concern that fisheries were already being over-harvested. Early estimates from scientists at that time suggested that by the 1970's total biomass of Georges Bank had already been halved, and elasmobranch species had begun to replace the over-harvested gadoid species [@fogarty1998]. Having such a large disturbance which pre-dates our time series is suggestive that the measured steepening of size spectrum slope we observed in this area and the adjacent Gulf of Maine are potentially the tail-ends of a longer and more severe ecosystem decline. While metrics of overall fishing pressure do not align exactly with trawl survey coverage, historical records and anecdotal evidence fro that time suggest that groundfish fishing pressure in these areas are a fraction of their what their impacts were in the 1960's and 1970's.

## Forces Preventing the Recovery of Large Individuals

This begs the question of why larger adult numbers never began to recover in these regions. Looking at abundance and biomass information from the survey there was evidence of strong recruitment among smaller individuals \< 1kg, but there has since not been a recovery in fishes larger than 1kg outside of the elasmobranchs. Work by [@pershing2015] suggested this prolonged recovery period may be due to a lack of accounting for temperature change in fisheries management. At the time of that research, the regional temperatures had only begun rising, and could have been considered at that time an acute stressor. Since that time the region has experienced nearly a decade of sustained above-average temperatures. There are signs that the success seen in recruitment and survival of even the smaller size classes is declining. While temperature change has been associated with changes in growth rates and size-at-age, so too have size-selective fishing practices, making it difficult to disentangle the importance of exploitation & temperature on the overall community size structure when body size integrates these two forces [@shackell2007].

### Potential Drivers Timeseries:

The following panels show the historical changes in each of the drivers. Landings have been scaled by average total landings within each region across all years. SST and GSI have not been scaled. This is different from how they are implemented in the regression analyses, when the landings were scaled over the 1982-2019 period.

::: panel-tabset
#### Surface Temperature Anomalies

```{r}
#| label: temperature-index
#| fig.height: 8


# Plot the timelines
temp_regimes %>% 
  mutate(anom_direction = ifelse(sst_anom > 0, "Positive", "Negative")) %>% 
  ggplot( aes(yr, sst_anom)) +
  geom_col(aes(fill = anom_direction),  
           linewidth = 0.75, alpha = 0.8) +
  geom_hline(yintercept = 0, linewidth = 0.5, lty = 1) +
  scale_x_continuous(expand = expansion(add = c(0.25,0.25))) +
  scale_y_continuous(labels = number_format(suffix = " \u00b0C")) +
  scale_fill_gmri() +  
  facet_wrap(~survey_area, ncol = 1) +
  theme(
    legend.title = element_blank(),
    legend.background = element_rect(fill = "transparent"),
    legend.key = element_rect(fill = "transparent", color = "transparent")) + 
  labs(
    title = "NW-Atlantic Regional SST Anomalies",
    x = "", 
    y = "Surface Temperature Anomaly",
    caption = "Anomalies calculated using 1982-2011 reference period.") +
  guides(fill = "none",
         label = "none",
         color = guide_legend(keyheight = unit(0.5, "cm")))
```

#### GARFO Landings

```{r}
#| label: landings-index
#| fig.height: 8

####  Landings  ####
landings_summ %>% 
  mutate(anom_direction = ifelse(total_wt_z > 0, "Positive", "Negative")) %>% 
  ggplot(aes(year, total_wt_z, group = survey_area)) +
  geom_col(aes(fill = anom_direction),  
           linewidth = 0.75, alpha = 0.8) +
  geom_hline(yintercept = 0, linewidth = 0.5, lty = 1) +
  scale_fill_gmri() +  
  facet_wrap(~survey_area, ncol = 1) +
  scale_x_continuous(breaks = seq(1950, 2030, by = 10)) +
  labs(x = "Year", y = "Finfish Landings Index (z)",
       title = "GARFO Finfish Landings")  +
  guides(fill = "none")
```

#### Climate Modes

```{r}
#| label: climate-modes

####  Climate Drivers  ####
clim_drivers %>% 
  group_by(year = lubridate::year(Time), Var, EPU) %>% 
  summarise(Value = mean(Value, na.rm = T)) %>% 
  mutate(anom_direction = ifelse(Value > 0, "Positive", "Negative"),
         Var = str_to_title(Var)) %>% 
  ggplot(aes(year, Value, group = EPU)) +
  geom_col(aes(fill = anom_direction),  
           linewidth = 0.75, alpha = 0.8) +
  geom_hline(yintercept = 0, linewidth = 0.5, lty = 1) +
  scale_fill_gmri() +  
  facet_wrap(~Var, ncol = 1) +
  scale_y_continuous(labels = scales::comma_format()) +
  scale_x_continuous(limits = c(1960, 2020), 
                     breaks = seq(1950, 2030, by = 10)) +
  labs(x = "Year", y = "Index",
       title = "Environmental Drivers") +
  guides(fill = "none")
```
:::

# Supplemental Materials

::: panel-tabset
## Species Functional Groups

```{r}
#| label: tbl-functional-groups
#| tbl-cap: "Common and scientific names for the species that constitute each functional group used in our analyses. X markers are used to indicate which regions each species has been caught in the data."


# Tidy up a key of what is labeled what
spec_key <- catch_size_bins %>% 
  distinct(comname, scientific_name, hare_group) %>% 
  mutate(
    hare_group = ifelse(hare_group %in% c("elasmobranch", "groundfish"), hare_group, str_c(hare_group)),
    hare_group = ifelse(is.na(hare_group), "Not Classified", hare_group),
    hare_group = str_to_title(hare_group),
    hare_group = fct_drop(hare_group),
    hare_group = factor(hare_group, levels = c("Coastal", "Diadromous", "Elasmobranch", "Groundfish",  "Pelagic", "Reef", "Not Classified"))) %>% 
  group_by(hare_group) %>% 
  mutate(n = str_c("(",n(),")")) %>% 
  ungroup()


# Reformat the data to check presence absence
catch_size_bins %>% 
  distinct(comname, survey_area) %>% 
  mutate(Presence = "X") %>% 
  complete(comname, survey_area) %>% 
  left_join(spec_key, by = "comname") %>% 
  mutate(comname = str_to_title(comname),
         Presence = ifelse(is.na(Presence) == FALSE, "X", ""),
         Presence = ifelse(is.na(Presence) == TRUE, "", Presence)) %>% 
  left_join(area_df, by = "survey_area") %>% 
  select(hare_group, n, comname, scientific_name, area, Presence) %>% 
  pivot_wider(names_from = area, values_from = Presence) %>% 
  arrange(hare_group, comname) %>% 
  group_by(hare_group,n) %>% 
  gt()  %>% 
  gt::tab_header(title = md("**Functional Group Assignments and Regional Presence/Absence**")) %>% 
  tab_stubhead(label = md("**Functional Group**")) %>% 
  cols_label(
    comname = md("*Common Name*"),
    scientific_name = md("*Scientific Name*"),
    `Gulf of Maine` = md("*Gulf of Maine*"),
    `Georges Bank` = md("*Georges Bank*"),
    `Southern New England` = md("*Southern New England*"),
    `Mid-Atlantic Bight` = md("*Mid-Atlantic Bight*")
    ) %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_row_groups()) %>% 
  gt::tab_source_note(source_note = md("*Functional group assignments adapted from Hare et al. 2010*"))

```

## GARFO Landings Summary

```{r}
#| label: garfo-landings-summary

# Add the labels into the landings data and remove what we don't need there:
landings %>% 
  filter(between(year, 1960, 2019),
         !str_detect(sppname, "CONFIDENTIAL")) %>% 
  mutate(
    decade = floor_decade(year),
    sppname = str_to_title(sppname)) %>% 
  group_by(survey_area, decade,  sppname) %>% 
  summarise(avg_landings_lb = mean(`landed lbs`, na.rm = T),
            across(.cols = c(landings_lb = `landed lbs`, 
                             value = value), 
                   .fns = sum, 
                   .names = "total_{.col}"),
            .groups = "drop") %>% 
  group_by(survey_area, decade) %>% 
  slice_max(order_by = total_landings_lb, n = 3) %>% 
  gt() %>% 
  gt::tab_header(title = md("**Top Commercial Fisheries Landings of Northeastern US (by weight)**")) %>% 
  tab_stubhead(label = md("*Harvest Region*")) %>% 
  fmt_number(columns = c(avg_landings_lb, total_landings_lb, total_value),
             use_seps = T, 
             sep_mark = ",",
             suffixing = T) %>% 
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_row_groups()) %>% 
  cols_label(
    decade            = md("*Decade*"),
    sppname           = md(""),
    avg_landings_lb   = md("*Avg. Annual Landings (lb.)*"),
    total_landings_lb = md("*Total Landings (lb.)*"),
    total_value       = md("*Total Value ($)*")) %>%
  gt::tab_source_note(source_note = md("*Landings data obtained from the Greater Atlantic Regional Fishing Office (GARFO)*"))
  
```
:::

### GARFO Regions

```{r}
#| label: GARFO-map
#| eval: false


# Map it out? - takes forever from Chesapeake bay
garfo_map <- ggplot(stat_zones) +
  geom_sf(aes(fill = survey_area), alpha = 0.8) +
  geom_sf(data = us_poly) +
  geom_sf(data = canada) +
  coord_sf(xlim = c(-76.4, -64.4), ylim = c(35, 45.5), expand = F) +
  scale_fill_gmri() +
  theme_bw() +
  map_theme(legend.position = c(0.6, 0.125),
            legend.background = element_rect(fill = "white")) +
  guides(fill = guide_legend(title = "", nrow = 2)) +
  labs(caption = "Map of the allocation of GARFO statistical zones into the four regions used in this study.")
# ggsave(garfo_map, filename = here::here("R/nmfs_size_spectra/images/garfo_regions.png"), bg = "white")


```

# References
