---
title: "Community Size-Structure of the Northwest Atlantic Groundfish Communities, Response to Direct Disturbance and a Changing Environment"
author: 
    name: "Adam A. Kemberling"
    title: "Quantitative Research Associate"
    url: https://github.com/adamkemberling
    affiliation: Gulf of Maine Research Institute
description: | 
  Results from a Individual Size Distribution Analysis of the Northeast US Groundfish Community
date: "`r Sys.Date()`"
# format: docx
format:
  html:
    self-contained: true
    code-fold: true
    code-tools: true
    df-print: kable
    toc: true
    toc-depth: 2
editor: source
execute: 
  echo: false
  warning: false
  message: false
  fig.height: 6
  fig.width: 7
  fig.align: center
  comment: ""
bibliography: references.bib
csl: ices-journal-of-marine-science.csl
---

```{r}
#| label: load packages and functions

####  Packages  ####
library(car)
library(targets)
library(rnaturalearth)
library(here)
library(sf)
library(gmRi)
library(patchwork)
library(gt)
library(knitr)
library(tidyverse)
library(ggstream)
library(broom)
library(bcp)
library(scales)
library(corrplot)
library(readxl)
library(Hmisc)
library(MuMIn)
library(gtsummary)
library(broom.helpers)
# changepoint package
library(EnvCpt)

# Package Conflicts
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")

# Support functions
source(here("R/support/sizeSpectra_support.R"))

# Resource Path
res_path <- gmRi::cs_path("res")

# GGplot theme
theme_set(theme_gmri() + theme(
  panel.grid.major = element_line(linewidth = 0.5, linetype = 1, color = "gray"),
  panel.grid.major.x = element_line(linewidth = 0.5, linetype = 3, color = "gray"),
  panel.grid.minor = element_line(linewidth = 0.5, linetype = 3, color = "gray90"),
  axis.line = element_line(color = "black"),
  axis.line.y = element_line(color = "black"), 
  strip.background = element_rect(colour = gmri_cols("teal")),
  legend.position = "bottom"))

# Map polygons
us_poly <- ne_states("united states of america", returnclass = "sf")
canada <- ne_states("canada", returnclass = "sf")

# levels for faceting areas
area_levels <- c("Northeast Shelf, Full", "GoM", "GB", "SNE", "MAB")
area_levels_long <- c("Northeast Shelf, Full", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight")

# table to join for swapping shorthand for long-hand names
area_df <- data.frame(
  area = c("Scotian Shelf", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight", "All"),
  survey_area = c("SS", "GoM", "GB", "SNE", "MAB", "Northeast Shelf, Full"),
  area_titles = c("Scotian Shelf", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight", "Northeast Shelf, Full"))
```



```{r}
#| label: functions-cell

# Function to process summaries for various factor combinations
get_group_summaries <- function(data_in, ...){
  
  # Do some grouping to get totals
  group_totals <- data_in %>% 
    group_by(...) %>% 
    summarise(total_survey_catch = sum(numlen, na.rm = T),
              total_lw_bio = sum(sum_weight_kg, na.rm = T),
              total_strat_abund = sum(strat_total_abund_s, na.rm = T),
              total_strat_lw_bio = sum(strat_total_lwbio_s, na.rm = T), 
              .groups = "drop") 
  
  # length bins
  group_lengths <- data_in %>% 
    group_by(..., length_bin) %>% 
    summarise(lenbin_survey_catch = sum(numlen),
              lenbin_lw_bio       = sum(sum_weight_kg),
              lenbin_strat_abund  = sum(strat_total_abund_s),
              lenbin_strat_lw_bio = sum(strat_total_lwbio_s), 
              .groups = "drop") %>% 
    left_join(group_totals) %>% 
    mutate(
      perc_total_catch  = (lenbin_survey_catch - total_survey_catch) * 100,
      perc_lw_bio       = (lenbin_lw_bio - total_lw_bio) * 100,
      perc_strat_abund  = (lenbin_strat_abund - total_strat_abund) * 100,
      perc_strat_lw_bio = (lenbin_strat_lw_bio - total_strat_lw_bio) * 100)
  
  # weight bins
  group_weights <- data_in %>% 
    group_by(..., weight_bin) %>% 
    summarise(wtbin_survey_catch = sum(numlen),
              wtbin_lw_bio       = sum(sum_weight_kg),
              wtbin_strat_abund  = sum(strat_total_abund_s),
              wtbin_strat_lw_bio = sum(strat_total_lwbio_s),
              .groups = "drop") %>% 
    left_join(group_totals) %>% 
    mutate(
      perc_total_catch  = (wtbin_survey_catch / total_survey_catch) * 100,
      perc_lw_bio       = (wtbin_lw_bio / total_lw_bio) * 100,
      perc_strat_abund  = (wtbin_strat_abund / total_strat_abund) * 100,
      perc_strat_lw_bio = (wtbin_strat_lw_bio / total_strat_lw_bio) * 100)

return(list("length_bins" = drop_na(group_lengths),
            "weight_bins" = drop_na(group_weights)))
}



# Function to process summaries for various factor combinations
log2_bin_summaries <- function(data_in, ...){
  
  # Do some grouping to get totals
  group_totals <- data_in %>% 
    group_by(...) %>% 
    summarise(total_survey_catch = sum(numlen, na.rm = T),
              total_lw_bio = sum(sum_weight_kg, na.rm = T),
              total_strat_abund = sum(strat_total_abund_s, na.rm = T),
              total_strat_lw_bio = sum(strat_total_lwbio_s, na.rm = T), 
              .groups = "drop") 
  
  # weight bins
  group_weights <- data_in %>% 
    mutate(weight_bin = bin_label) %>% 
    group_by(..., left_lim) %>% 
    summarise(wtbin_survey_catch = sum(numlen),
              wtbin_lw_bio       = sum(sum_weight_kg),
              wtbin_strat_abund  = sum(strat_total_abund_s),
              wtbin_strat_lw_bio = sum(strat_total_lwbio_s),
              .groups = "drop") %>% 
    left_join(group_totals) %>% 
    mutate(
      perc_total_catch  = (wtbin_survey_catch / total_survey_catch) * 100,
      perc_lw_bio       = (wtbin_lw_bio / total_lw_bio) * 100,
      perc_strat_abund  = (wtbin_strat_abund / total_strat_abund) * 100,
      perc_strat_lw_bio = (wtbin_strat_lw_bio / total_strat_lw_bio) * 100)

return(list("weight_bins" = drop_na(group_weights)))
}




# Function to grab the correlation data and lag data
get_ccf_vector <- function(x,y){
  
  # Run the ccf
  ccf_data <- ccf(x,y,plot= F , lag.max = 15)
  
  # Get the signif:
  # https://www.squaregoldfish.co.uk/programming/r_acf_significance.md/
  # Not 100% sure n is the same for ccf as it is for acf, but...
  significance_level <- qnorm((1 + 0.95)/2)/sqrt(sum(!is.na(x)))
  
  data.frame(
    "acf" = ccf_data$acf,
    "lag" = ccf_data$lag,
    "sigpos" = significance_level,
    "signeg" = significance_level*-1
  )
}


# Plotting the correlation matrix over a specific decade
plot_period_correlations <- function(x, label) {
  x %>% 
    filter(response_short != "Spectra Slope",
           response_area != "All") %>% 
    ggplot(aes(var1, response_short, fill = r2)) +
    geom_tile() +
    geom_text(aes(label = signif)) +
    facet_grid(response_area ~ driver_type, scales = "free_x", space = "free") +
    scale_fill_distiller(palette = "RdBu", limits = c(-1, 1), breaks = c(-1,0,1)) +
    coord_cartesian(clip = "off") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          axis.text.y = element_blank(),
          strip.text.y = element_text(angle = 0),
          legend.position = c(1.205, -.25),
          legend.title = element_text(vjust = 0.75),
          plot.caption = element_text(hjust = 0)) +
    guides(fill = guide_colorbar(title.position = "top", title.hjust = 0.5)) +
    labs(x = "Correlated Feature", y = "Size Spectrum Slope",
         fill = "Correlation Coefficient",
         title = str_c(label, " | Correlation of Size Spectra to Hypothesized Drivers"),
         subtitle = "Correlation coefficients shown display same-year correlations.",
         caption = md("* Symbol signals a significant correlation at an alpha =  0.05"))
  
}

```

### Potential Journals:

1.  **ICES Journal of Marine Science**

# Abstract

The surface waters of the Northwest Atlantic Ocean are among the fastest warming on Earth. This area is highly-productive biologically, and there are concerns that ecological consequences will follow this rapid-warming. Research on the impacts of this rapid warming has primarily focused on high-profile and/or upper trophic level species. Ecological theory and laboratory studies suggest that elevated temperatures facilitate early maturation and smaller adult body-sizes. However, it is unclear whether that relationship might be mitigated against through adaptive behaviors in an open ocean environment. Here we've investigated ecosystem wide impacts on the individual size distribution (ISD) to track changes in community size structure. In cases where community responses are not adequate to counter the impacts of elevated temperatures, we anticipated a steepening of the size spectrum slope (ISD exponent). A steeper relationship relating to a reduction in larger sized individuals and an increased prevalence of smaller sized individuals. Using data from fisheries independent surveys we calculated the community size spectra for four regions along the US NE continental shelf. Correlation/regression analyses were then performed to then assess the degree to which these changes were in alignment to hypothesized bottom-up and top-down disturbances. At the regional scale, we found that community size structure changes (spectra slope) were the largest in the Northern regions, in the Gulf of Maine and Georges Bank. These areas are home to the coldest temperatures and the largest proportions of groundfish species in the community. Spectrum slope declines were most pronounced in the 80's and 90's, before the rapid warming of the last decade. The timing of these declines suggest that external factors drove the initial declines of larger-sized individuals within the communities, before elevated temperatures began to influence the ecosystem. Correlation analyses reveal that while fisheries landings are strongly correlated with these declines, bottom-up factors of zooplankton community metrics, Gulf Stream Index, and SST anomalies are also important. While the primary pressure of fisheries exploitation has declined dramatically over time, the recovery of larger-sized individuals has not been seen. That kind of recovery will likely depend on the elevated temperatures seen over the last decade.

# Introduction

## Rapid Warming of Northeast Shelf

The Northwest Atlantic is one of the fastest warming locations in the global oceans. Sea surface temperatures in the Gulf of Maine since 1982 have been warming at rates faster than 96% of the world’s oceans, with similar warming rates along the northwest Atlantic shelf (Pershing et al. (2018)). The persistent elevated temperature regime of the area is a result of several forces: a combination of shifting ocean currents and the unique bathymetry of the region. A Northward shift in the Gulf Stream directly increased the regional temperatures through increased transport of warm Gulf Stream water into areas like the Gulf of Maine. The Northward Gulf Stream shift is associated with a higher frequency of warm core rings, and the obstruction of cold-water Scotian Shelf current flow that otherwise counters the influence of the Gulf Stream on the region’s temperatures (Gangopadhyay et al. (2019); Meyer-Gutbrod et al. (2021)). The combination of these oceanographic changes has led to a warmer continental shelf habitat, which has caused alarm among scientists studying the ecology of the region.


Temperature plays a critical role on biological function, affecting many of the chemical reactions that underpin basic physiological function. Reactions impacting critical actions like locomotion and feeding behavior, metabolism and development, and even maturation rates. Because of these relationships, species have evolved thermal preferences around which these functions each operate efficiently. Individuals that are unable to maintain their thermal preferences internally risk metabolic costs unless they can follow their thermal preference in the environment through locomotion or adapt to less-favorable conditions through changes in behavior. In an era of anthropogenic climate change, there is an expectation that many species will be displaced from historic habitats in their efforts to follow their thermal preferences. Recent research in marine environments has shown evidence of this, as species are now shifting to higher latitudes and to deeper depths in the pursuit of more favorable conditions (Nye et al. (2009); Pinsky et al. (2013)). Other work suggests that temperature related impacts may manifest through physiological changes and changes in seasonal phenology, hindering species conservation efforts (Miller et al. (2018); Pershing et al. (2015); Meyer-Gutbrod et al. (2021)).
Need to connect temperature to size here:
Direct quote from Guiet et al. (2016) , but nails the connection back to temp expectations:


> Because it controls chemical reactions, temperature controls metabolic rates which underpin maintenance, growth or reproduction (Clarke and Johnston, 1999; Kooijman, 2010) as well as the functional responses to food density (Rall et al., 2012). Guiet et al. (2016)… In addition to the impact of temperature on communities’ intercepts (heights), the impact of temperature on the speed of the energy flow within communities may affect other properties, such as their resilience to perturbations or the intensity of trophic cascades (Andersen and Pedersen, 2009).


The potential for elevated temperatures to impact the size structure of an ecosystem has implications for the ecosystem resilience in the face of climate change, as well as the blue economies & natural resource systems that rely upon their good health.


## Size Spectra Theory

Size is a defining characteristic of species and mediates many ecological interactions and metabolic pathways (Brown et al. 2000) . Size is a major determining factor for the mobility of an organism which directly impacts the ability to evade predation, find foraging success, and efforts to locate and follow seasonal habitats. Size and physical dimensions also impact the metabolic costs associated with each of these behaviors (Hillaert et al. 2018),  mediating exchange rates with the immediate environment like heat loss or desiccation in terrestrial species (Gillooly et al. (2001); Heatwole et al. 1969). Body size even informs life history features like life span and the trophic position an individual might occupy through its impact on metabolism and resource use (White et al. 2007). 
Size structured environments are a fundamental organizational pattern that emerges from these relationships add_citation. Within strongly size-structured ecosystems, growth and maturity changes alter fitness and ultimately determine whether a species is successful in that environment add_citation . Ecological theory is rich with models relating how energy transfers from smaller prey species to larger predatory trophic levels, the allocation of energy for growth, and the trade offs of allocating energy towards those ends (Bertalanffy (1938); Bertalanffy (1957)); add_size_theory_citations). A globally persistent pattern in ecology entangled in those relationships and their critiques is the decline in abundance with increasing body size (Damuth (1981); Currie (1993); Sheldon et al. (1972); Loeuille and Loreau (2006)).
The relationship between size and abundance integrates processes operating on the cellular, individual, and community levels simultaneously. By measuring how this relationship changes, scientists can detect whether the fundamental energy transfer pathways from small individuals to larger predators have been impacted or altered, an outward indication of ecosystem-wide changes. The quantities for size and abundance are often the most readily collected data assets of any ecological community. This creates an opportunity to learn much about a system from a relatively low-effort in data collection. For these reasons, size spectrum analyses and individual size distribution (ISD) methods have gained increasing attention as an entry point to assessing ecosystem health and to detect system-wide disturbance (Shin et al. (2005)‘; Pomeranz et al. (2022); White et al. (2007)). Spectra models are also taxon agnostic and avoid the need to explicitly articulate each predator-prey interaction. The “size spectrum” describes the distribution of abundance or biomass as a function of individuals’ mass on a log–log scale (Guiet et al. (2016); Kerr and Dickie (2001)) . Size spectra condense the complexities of predator prey networks and their interactions into taxon-agnostic size-based indices. These indices capture the emergent properties of a system, and have become increasingly used as indicators of ecosystem health. Within the context of fisheries management, changes in spectrum slopes have been associated with fishing exploitation, primarily through the targeted removal of larger individuals (Bianchi et al. (2000); Shin et al. (2005)). Numerical experiments have also linked changes in slope to environmental disturbances (Guiet et al. (2016)). Size spectra have also been shown to express predictable relationships between ecosystems of similar productivity levels as well as from distinct temperature regimes (Guiet et al. (2016)).
Use Sprules and Barth Paper to discuss applications
Use Pomerantz paper & Edwards to extend into ISD


## External Drivers of Size Structure

Our study area’s ecology is one that has historically been studied through the fishing grounds it once supported and their challenges. The focus has traditionally been on the success/failure of single species management actions, however there are growing efforts to better understand the full ecological impact of this endeavor. Fishing practices are often size-selective, with harvesters targeting larger individuals for a higher yield on any time and capital invested. Larger individuals are also commonly the older individuals in a population, having lived longer to achieve those sizes. With the unfortunate side-effect of having an outsized impact on the reproductive potential of the population. Larger individuals have a greater impact on population resilience and recovery, capable of holding more (and often of higher quality) eggs which increase the odds of recruitment success. Heavily fished populations have been shown to exhibit changes to growth rates, suggesting an additional genetic impact from size-selective fishing on the target-species. The direct removal of larger individuals carries a predictable impact on the size distributions of the target species, but may also change that of non-target species as well. Non-target species may be impacted as by-catch from the fishery, or may alternatively find relief from predation if the target species is a direct predator. This can even create circumstances where prey species can crowd out, or outcompete the juveniles of their natural predators, reinforcing the new community. Industrial fishing practices are inherently size-selective, with an outsized footprint relative to small-scale fisheries. Large removals have an immediate and measurable impact on the community size-distribution with potential additional impacts on the future population as well. Size-based harvest in fisheries has been shown to create selective pressures that promote characteristics of early maturation at smaller sizes add_citation. These consequences may independently or in combination create circumstances where a community is unable to return to conditions before these  disturbances occurred.
reference gb/ss spectra early work (Duplisea & Kerr 1995, or Kerr book)


The ecological community in our study area has with no uncertainty been altered through human behaviors over the last century. Early research estimated that biomass had more than halved in some areas by the 1960’s, pre-dating federal monitoring efforts and inspiring their formation (Fogarty and Murawski (1998)). Key stocks that supported international fishing efforts collapsed, and recovery efforts have in many species fallen short. Work exploring retrospective patterns in stock assessment predictions have highlighted that these shortfalls can be partially attributed to un-accounted for influences of a changing environment. However, other research has noted that the community structure itself may have changed, and with it, how it responds to outside pressure. That same early research noted that species replacement of commercial target species by skate and dogfish was happening as their numbers quickly began to rise. A wealth of anecdotal evidence and records from commercial landings support the conclusion that the region does not support either the fisheries or the fishing  communities in the same way today. The Lobster industry remains King in the Gulf of Maine, another side-effect of the decline in large predators in the area. With other similar declines in large-fish populations up and down the coast, scientists have recently brought the focus of their attention away from single-species management to more holistic and ecosystem-wide approaches.

## Regime Shifts in Marine Systems

**Goals:**

Want to hit: dynamic tipping points, hysteresis, and non-stationary functional relationships. Show examples of the idea on single species, suggest that we size spectra can be used to check the whole community for tipping points

When an ecosystem experiences an abrupt change in its internal dynamics in response to internal or external pressures, it is said to have crossed a tipping point. When these tipping points are crossed, it often becomes difficult or impossible to return to the previous state. This is true even when the pressures that caused the shift have ceded and favorable conditions have returned, a phenomenon known as hysteresis. This phenomenon has been observed in individual population trends, and is a major factor undermining sustainable management of fisheries (Blocker 2023; Sguottti 2019) . Another feature that may accompany a regime shift is the altering of functional relationships with other biotic or abiotic features of the environment. When this occurs, the productivity of a species and/or its resilience to stress may be altered, and assumptions underpinning sustainable harvest thresholds should be updated (Blocker 2023). At the community level the alternative community states may support a different composition of species.  The community itself may respond differently to the same stressors/stimulus. In the Northeast Shelf Region there is evidence to suggest that through a combination of human influences and environmental changes that the ecology of some areas may have crossed ecological tipping points. This study explores whether evidence for an ecological tipping point can be seen through the perspective of the size spectrum, and whether there is evidence for hysteresis, or non-stationarity in the community’s relationships to external forces.


### Purpose

Understanding that human populations depend on the health of their ecosystems, there is a need to better understand when we’ve crossed ecological tipping points. One way to capture these changes is through the use of community-wide metrics that capture the inner-efficiencies of many trophic interactions. Size based indices are metrics that can be estimated from the information that has historically been available from long-term survey efforts. These indices have been shown to be sensitive to the impacts of fishing, but also capture environmental dynamics as well. We estimated size spectrum relationships to use as size based indicators of the ecological communities within each sub-region of the Northeast US continental shelf. Based on ecological theory we believe that the sustained increases in temperature in the NW Atlantics should have a physiological impact on the community size structure. We hypothesize that rapid warming alters the community through the direct influence of temperature on metabolism, growth, and population productivity. We also hypothesize that size spectra will uniquely reflect the physical environment of the regions, with less steep size spectra in regions with colder temperatures and of higher primary productivity. Understanding that temperature is not the sole mechanism for altering community size structure, we will explore how the temporal structures in size spectra have changed in relation to major external drivers. We posit that distinct patterns of deviation from that signature may reflect a fundamental change in the ecology of that region, an indication of a possible ecological regime change, and an altered relationship to human and environmental disturbances.


# Methods


## Fish Data Source and Processing


```{r}
#| label: load-survdat

# 1. Biological data used as input
withr::with_dir(rprojroot::find_root('_targets.R'), tar_load(catch_log2_labelled))   

# rename and format
catch_size_bins <- catch_log2_labelled %>% fill_func_groups()

# Get total number of species
n_species <- length(unique(catch_size_bins$comname))
```

Data on the biomass, abundance, and size of fish on the Northeast U.S. Shelf were collected as part of the Northeast Fisheries Science Center’s bottom trawl survey (Grosslein 1969, Azarovits 1981, Politis 2014). This survey is conducted from Cape Hatteras, North Carolina to the Gulf of Maine each year in the spring and in the fall.  The survey follows a stratified random sampling design, with strata defined based on depth, bottom habitat, and latitude. Trawls are performed for a fixed duration at each station, reporting aggregate abundance and biomass for all species caught, and measuring individual lengths and weights for the catch of each species or a sub-sample if that catch is large. Correction factors were applied to aggregate species abundance and biomass to account for changes in vessels, gear, and doors used in the survey over time (Sissenwine and Bowman 1978, Byrne and Forrester 1991, Miller et al. 2010). However, abundance and biomass at length needed to be estimated after these aggregate corrections. As such, abundance at length for each species was adjusted to match the corrected aggregate species abundance at each station, such that for each species, the sum of the resulting estimated abundance numbers across each length is equal to the corrected aggregate abundance. 

```{r}
#| label: study-area-maps

#------------- Trawl Strata --------------

# Get the paths to the shapefiles used to mask
trawl_paths <- gmRi::get_timeseries_paths(region_group = "nmfs_trawl_regions", box_location = "cloudstorage")

# Polygons for each region
all_poly <- read_sf(trawl_paths[["inuse_strata"]][["shape_path"]]) 

# Make the Map
strata_map <- ggplot() +
  geom_sf(data = us_poly) +
  geom_sf(data = canada) +
  geom_sf(data = all_poly, aes(fill = survey_area), alpha = 0.8) +
  coord_sf(xlim = c(-76.4, -64.4), ylim = c(35, 45.5), expand = F) +
  scale_fill_gmri() +
  theme_bw() +
  map_theme(legend.position = c(0.6, 0.125), 
            legend.background = element_rect(fill = "white"),
            plot.caption = element_text(hjust = 0)) +
  guides(fill = guide_legend(title = "", nrow = 2)) 

# ----------- Landings Statistical Zones -----------------


# Shapefiles for the fisheries stat zones
stat_zones <- read_sf(str_c(res_path, "Shapefiles/Statistical_Areas/Statistical_Areas_2010_withNames.shp"))

# Make a list of zones to roughly match the survey areas:
fish_zones <- list(
  "Gulf of Maine" = c(511:515, 464, 465),
  "Georges Bank" = c(521, 522, 525, 561, 562),
  "Southern New England" = c(611, 612, 613, 616, 526, 537, 538, 539),
  "Mid-Atlantic Bight" = c(614:615, 621, 622, 625, 626, 631, 632))


# Trim out what we don't need and label them
stat_zones <- stat_zones %>% 
  mutate(
    survey_area = case_when(
      Id %in% fish_zones$"Gulf of Maine" ~ "Gulf of Maine",
      Id %in% fish_zones$"Georges Bank" ~ "Georges Bank",
      Id %in% fish_zones$"Southern New England" ~ "Southern New England",
      Id %in% fish_zones$"Mid-Atlantic Bight" ~ "Mid-Atlantic Bight")) %>% 
  filter(survey_area %in% c(
    "Georges Bank", "Gulf of Maine", "Southern New England", "Mid-Atlantic Bight"))

# Map it out? - takes forever from Chesapeake bay
garfo_map <- ggplot(stat_zones) +
  geom_sf(aes(fill = survey_area), alpha = 0.8) +
  geom_sf(data = us_poly) +
  geom_sf(data = canada) +
  coord_sf(xlim = c(-76.4, -64.4), ylim = c(35, 45.5), expand = F) +
  scale_fill_gmri() +
  theme_bw() +
  map_theme(legend.position = c(0.6, 0.125),
            legend.background = element_rect(fill = "white"),
            plot.caption = element_text(hjust = 0)) +
  guides(fill = guide_legend(title = "", nrow = 2)) 
# ggsave(garfo_map, filename = here::here("R/nmfs_size_spectra/images/garfo_regions.png"), bg = "white")


# Map them together
(strata_map | garfo_map) + plot_layout(guides = "collect")
```

### Community Composition & Functional Groups

Analyses were performed using 68 species. These species were selected based on the availability of published weight-at-length relationships (Wigley et al. 2003) and represented 98.98% of the total biomass caught in the survey. Each species was assigned to a functional group based on life history and geography using the definitions of (Hare et al. (2016)). Functional groups included coastal fishes, diadromous fishes, elasmobranchs, groundfish, pelagic fishes, and reef fish species (Table 1.). Six species with available length-weight details did not have a functional group designation, these species were designated as reef species. Exploratory analyses showed that the pelagic species biomass was low in all regions, and is unlikely to be representative of true biomass trends due to gear selectivity.

```{r}
#| label: biomass-group-summaries

# This cell provides the group summaries for how much biomass exists in each size/functional group

# Actual log10 bins
# Do some grouping by year survey area and functional group
fgroup_area <- log2_bin_summaries(data_in = catch_size_bins, Year, survey_area, hare_group)
fgroup_all  <- log2_bin_summaries(data_in = catch_size_bins, Year, hare_group)
fgroup_all  <- fgroup_all %>% map(~mutate(.x, survey_area = "Northeast Shelf, Full"))


# Join the overall values back in
fgroup_area <- map2(fgroup_area, fgroup_all, bind_rows)

# Add new units in millions
fgroup_area$weight_bins <- fgroup_area$weight_bins %>%
  mutate(strat_abund_mill = wtbin_strat_abund / 1e6,
         strat_lwbio_mill = wtbin_strat_lw_bio / 1e6,
         survey_area = factor(survey_area,
                              levels = area_levels))

```

Published length-weight relationships ( Wigley et al. 2003 ) were used to convert from length data, available for all individuals, into their corresponding biomass-at-length. To account for differences in sampling effort among survey strata, all corrected abundance-at-length data were area-stratified. Area-stratified biomass-at-length values were then computed as the product of area-stratified abundance-at-length and estimated weight-at-length. All analyses were performed using area-stratified abundances and their associated area-stratified biomass estimates, hereby referred to as simply abundance-at-length & biomass-at-length



## Community Metrics

Our analyses used all data collected during the spring and fall surveys from 1970-2019. Data were grouped using survey-design strata into four sub-regions: Gulf of Maine, Georges Bank, Southern New England, Mid-Atlantic Bight (Figure 1.). These sub-regions have been widely used in regional ecological studies (e.g., ). For each region, we developed the following time series of ecological indicators:


1.  Annual mean abundance and biomass by functional group

2.  Annual mean length and weight of the aggregate community and for each functional group

3.  Annual estimates of the community size spectrum slope



### Quantifying Body Size Changes

Changes in the size structure of the community was measured using the average length and weight across all species, each weighted by their species and size-specific catch rates. The average body length (cm) and body weight (kg) was calculated for each region and within each functional group. 

$$X_j = \frac{ \sum_{}^{}  n_iX_{ij}  }{  \sum~w_i  }$$ Where $X$ is the body-size metric, $j$ is the year, & $n$ is the abundance for each station $i$.

Data for body size trends were not truncated using any minimum or maximum size and reflect all available catch data for the 68 species in this study for which biomass-at-length could be estimated.

```{r}
#| label: body size data

# Load the average body size data
withr::with_dir(rprojroot::find_root('_targets.R'), 
                tar_load(mean_sizes_ss_groups))   

# Overall across all regions
shelf_size <- mean_sizes_ss_groups %>% 
  filter(`group ID` == "single years") %>% 
  mutate(survey_area = "Northeast Shelf, Full")

# Grouped on year and region
regional_size <- mean_sizes_ss_groups %>% 
  filter(`group ID` == "single years * region") %>% 
  bind_rows(shelf_size) %>% 
  mutate(Year = as.numeric(Year),
         survey_area = as.character(survey_area)) %>% 
  left_join(area_df) %>% 
  mutate(area_titles = factor(area_titles, levels = area_levels_long),
         area_copy = survey_area)




# 2. Grouped on year, region, & functional group
functional_group_size <- mean_sizes_ss_groups %>% 
  filter(`group ID` == "single years * region * functional group") %>% 
  mutate(Year = as.numeric(Year),
         survey_area = factor(survey_area, 
                              levels = area_levels))

# Refactor areas
functional_group_size <- functional_group_size %>% 
  mutate(survey_area = as.character(survey_area))  %>% 
  left_join(area_df) %>% 
  mutate(area_titles = factor(area_titles, levels = area_levels_long),
         area_copy = survey_area)

```

### Estimating Size Spectra

```{r}
#| label: load the size spectrum indices
withr::with_dir(rprojroot::find_root('_targets.R'), 
                tar_load(size_spectrum_indices))   

# Grab the overall shelf-wide SS
shelf_ss <- size_spectrum_indices %>% filter(`group ID` == "single years") %>% 
  mutate(survey_area = "Northeast Shelf, Full")


# Grab SS Groups for each region
region_indices <- size_spectrum_indices  %>% 
  filter(`group ID` == "single years * region") %>% 
  bind_rows(shelf_ss) %>% 
  mutate(yr = as.numeric(as.character(Year)),
         survey_area = factor(survey_area, levels = area_levels),
         sig_fit = ifelse(log2_sig_strat < 0.05, "Significant", "Non-Significant"))

# Make a copy so we can gray out the lines in plots
# Merge the long names for labels "area_titles"
region_indices <- region_indices  %>% 
  mutate(survey_area = as.character(survey_area))  %>% 
  left_join(area_df) %>% 
  mutate(area_titles = factor(area_titles, levels = area_levels_long),
         area_copy = survey_area)

```

Abundance size spectra were calculated using the area-weighted abundance at length information from the catch data. Lengths of individuals in the catch data are measured to the nearest cm, with smaller specimens measured to the nearest millimeter. Because individual biomass is estimated from those length measurements, there is a range of possible body mass values between the increments used (cm and mm). The relationship between length and mass in fishes is exponential and taxon specific. Using only the lower or upper end of those ranges introduces biases which are different for each species and increase with larger body-sizes. To account for these biases, we used the extended likelihood method (MLEbin) of Edwards et al. (2020). This method estimates the exponent of size spectra (b) for a bounded power law relationship between abundance and length-estimated biomass. The exponent of this relationship is analogous to slope estimates from linear regression on logarithmic axes, with a steeper slope (i.e. a more negative b) indicating fewer large-bodied individuals and/or more small-bodied fishes (Carvalho 2021) . This method has been shown to be the most accurate for estimating the exponent of size spectra when tested against alternative methods that require arbitrary decisions around binning the data and make estimates difficult to compare directly across studies (White 2008; Sprules & Barth 2016; Edwards 2017).


$$
\begin{align*}
f(x) = \frac{ (\lambda + 1)x^{\lambda} }{ x^{\lambda+1}_{max} - x^{\lambda+1}_{min} }~~~~~~\lambda\neq1,
\\
\\
f(x) = \frac{1}{logx_{max} - logx_{min}}~~~~~~\lambda=1
\end{align*}
$${#eq-ISD}

Using this method, the size spectra exponent (b) was estimated for each region from 1970 to 2019. A minimum biomass of 1g was used for the lower bound and a maximum biomass of 10kg was used as an upper bound for the ISD’s bounded power law probability density function Equation 1, where X is body mass & Λ  is the scaling exponent of the abundance density function. This body-size range constitutes 97.83% of all the biomass for the 68 species included in this study. This body-size range truncation was done to account for poor gear selectivity at the smallest and largest size ranges. This measure also imposes shared bounds to the size range covered by our size spectra, reflecting only the ranges  we’d expect to consistently sample across these different areas. Exponents of size spectra (b) were calculated using code modified from the sizeSpectra package (Edwards et al. (2017); Edwards et al. (2020)) and implemented using the R statistical programming language.

## Drivers of Size Spectra Changes

The impact of external factors on the changes in size spectra were explored using multiple regression analyses. Annual variation in size spectrum slopes was modeled using several hypothesized drivers including the environmental state and measures of anthropogenic disturbances. The large-scale environmental drivers included were regional SST anomalies & the larger-scale impact of the Gulf Stream Index (GSI). Fishing pressure represents the primary top-down anthropogenic driver in the region, and was investigated using the aggregate regional commercial landings data. Bottom-up ecological interactions were explored using zooplankton indices from the EcoMon plankton survey. Each of these drivers (climate, productivity, fishing) have independently been shown in other works to have measurable impacts on size spectra. Our exploratory investigations here are an effort to evaluate which best explain the trends within our study region, and whether those relationships have changed over the study period.


### Gulf Stream Index

Data for the Gulf stream index (GSI) was obtained from the ecodata package in R (Bastille & Hardison 2018) . This package supplies GSI data at monthly intervals following the methodology of Pérez-Hernández and Joyce (2014) and Joyce et al. (2019) , using as sea level height anomaly data from the Copernicus Marine Environment Monitoring Service.

```{r}
#| label: gsi-data-prep
#| eval: true

####  2. Climate Drivers  ####

# From ecodata: GSI & Zooplankton

####  GSI

# Gulf Stream Index
gsi <- ecodata::gsi %>% 
  mutate(
    Time = str_pad(as.character(Time), width = 7, side = "right", pad = "0"),
    year = as.numeric(str_sub(Time, 1, 4)),
    month = str_sub(Time, -2, -1),
    Time = as.Date(str_c(year, month, "01", sep = "-")))
```

### Sea Surface Temperature Data

Global sea surface temperature (SST) data were obtained via NOAA’s Optimum Interpolation SST analysis (OISSTv2), which provides daily SST values at a 0.25° latitude x 0.25° longitude resolution (Reynolds et al. 2007). A daily climatology for every 0.25° pixel in the global data set was created using average daily temperatures spanning the period of 1982-2011. Daily anomalies were then computed as the difference between observed temperatures and the daily climatological average. OISSTv2 data used in these analyses were obtained from the NOAA Physical Sciences Laboratory, Boulder, Colorado, USA from their website at https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.html.

SST data were regionally averaged to match the four survey sub-regions described above (Figure 1a),  producing daily time series for each area. These were then averaged into annual time series of surface temperatures and anomalies. All averaging between satellite grid units was done with area-weighting of the grid cells to account for differences in cell size with latitude/longitude in the OISSTv2 data. 

```{r}
#| label: load temperature data

# Load the regional temperatures from {targets}
withr::with_dir(rprojroot::find_root('_targets.R'), 
                tar_load(regional_oisst))


# Get regional averages
temp_regimes <- regional_oisst %>% 
  filter(yr > 1981) %>% 
  mutate(
    regime = ifelse(sst_anom > 0, "hot", "cold"),
    survey_area = ifelse(
      survey_area == "all", 
      "Northeast Shelf, Full", 
      survey_area),
    survey_area = factor(survey_area, levels = area_levels)) 



```

### Commercial Fishing

Fishing pressure in the region was indexed using state and federal commercial fishing landings.  These data were obtained from the Greater Atlantic Regional Fisheries Office (GARFO) for statistical areas that are routinely used for fisheries reporting and management (Figure 1., right).  Individual statistical areas were aggregated into regions that closely align with the survey areas we defined for the size spectra analyses (Figure 1 b.). Each region’s commercial landings timeline was scaled using z-score normalization

```{r}
#| label: garfo-data-prep
#| eval: true

####  GARFO Landings  ####

# Load the GARFO landings data:
# Landings of finfish* sheet 5
landings <- read_xlsx(
  path = str_c(res_path, "GARFO_landings/KMills_landings by area 1964-2021_JUN 2022.xlsx"), sheet = 5) %>% 
  rename_all(tolower)


# Add the labels into the landings data and remove what we don't need there:
landings <- landings %>% 
  mutate(
    survey_area = case_when(
      `stat area` %in% fish_zones$"Gulf of Maine" ~ "Gulf of Maine",
      `stat area` %in% fish_zones$"Georges Bank" ~ "Georges Bank",
      `stat area` %in% fish_zones$"Southern New England" ~ "Southern New England",
      `stat area` %in% fish_zones$"Mid-Atlantic Bight" ~ "Mid-Atlantic Bight")) %>% 
  filter(survey_area %in% c("Georges Bank", "Gulf of Maine", "Southern New England", "Mid-Atlantic Bight")) %>% 
  mutate(survey_area = factor(survey_area, area_levels_long))



# Get Summaries
landings_summ <- landings %>% 
  drop_na() %>% 
  rename(
    "weight_lb" = `landed lbs`,
    "live_lb" = `live lbs`) %>% 
  drop_na() %>% 
  group_by(year, survey_area) %>% 
  summarise( 
    across(
      .cols = c(value, weight_lb, live_lb), 
      .fns = list(mean = ~mean(.x , na.rm = T), 
                  total = ~sum(.x , na.rm = T)), 
      .names = "{.fn}_{.col}"), 
    .groups = "drop")  


# Scale the landings to create an index by area
landings_summ <- landings_summ %>% 
  group_by(survey_area) %>% 
  mutate(total_wt_z = base::scale(total_weight_lb)) %>% 
  ungroup()
```


### Zooplankton Indices

Zooplankton data was from the National Oceanographic and Atmospheric Administration Marine Resources Monitoring, Assessment and Prediction (MARMAP) program and Ecosystem Monitoring (EcoMon) cruises detailed extensively in Kane (2007), Kane (2011), and Morse et al. (2017). Abundance anomalies were computed from the expected abundance on the day of sample collection. The small copepod index is computed by averaging the individual abundance anomalies of Pseudocalanus spp., Centropages hamatus, Centropages typicus, and Temora longicornis. The “large-copepod” anomaly values are the abundance anomaly of Calanus finmarchicus, the largest copepod in the Northeast U.S. region. Anomalies for both large and small zooplankton groups were averaged within ecological production units by the data provider, producing annual timeseries for the Gulf of Maine, Georges Bank, and Mid-Atlantic Bight. The timeseries for the Mid-Atlantic Bight EPU was used as a predictor for both the Mid-Atlantic Bight and Southern New England regions in our analyses. This decision was made to address how the spatial coverage of this EPU spans the joint area of our two southernmost regions. 


```{r}
#| label: zp-data-prep

# ADD Zooplankton: small and large zooplankton anomalies for the EPUS
# Don't need to scale

# Load original data
zp_sli <- ecodata::zoo_sli_anom %>% filter(EPU != "SS")

# Reformat to match the other indices
zp_index <- zp_sli %>% 
  #EPUS combine SNE and MAB, we can repeat the values here
  bind_rows(filter(zp_sli, EPU == "MAB") %>% mutate(EPU = "SNE")) %>% 
  mutate(
    survey_area = case_when(
      EPU == "GB" ~ "Georges Bank",
      EPU == "GOM" ~ "Gulf of Maine",
      EPU == "MAB" ~ "Mid-Atlantic Bight",
      EPU == "SNE" ~ "Southern New England"),
    Units = str_c("zp_", Var)) %>% 
  select(year = Time, survey_area, Value, Units) %>% 
  pivot_wider(names_from = Units, values_from = Value)


```







```{r}
#| label: combine-drivers-table


####  Build a Common form for all Drivers:   ####
# Format: year, area, var, value

#---
#### Reshape all the inputs so that we can bind them together in one dataframe


# 1.Climate Features

# Join the climate modes together
clim_drivers <- gsi %>% filter(EPU == "All")
  

# Put climate drivers on annual schedule
clim_idx <- clim_drivers %>% 
  group_by(year, area = EPU, var = Var) %>% 
  summarise(value = mean(Value, na.rm = T),
            .groups = "drop")

#---
# 2. GARFO landings
landings_idx <- landings_summ %>% 
  group_by(year, area = survey_area) %>% 
  summarise(value = mean(total_wt_z, na.rm = T),
            .groups = "drop") %>% 
  mutate(var = "landings")


#---
# 3. Temperature
sst_idx <- temp_regimes %>% 
  left_join(area_df, by = "survey_area") %>% 
  select(year = yr, area, value = sst_anom) %>% 
  mutate(var = "sst_anom")


#---
# 4. Zooplankton
zp_idx <- zp_index %>% 
  pivot_longer(cols = c(zp_small, zp_large), 
               names_to = "var", values_to = "value") %>% 
  rename(area = survey_area)



#---
#  4. Size Spectrum Slopes & Intercepts
ss_idx <- region_indices %>% 
  select(
    year = Year, 
    survey_area, 
    spectra_slope = log2_slope_strat, 
    #spectra_int = log2_int_strat,
    isd_slope = b) %>% 
  pivot_longer(
    # cols = c(spectra_slope, spectra_int, isd_slope), 
    cols = c(spectra_slope, isd_slope), 
    names_to = "var", 
    values_to = "value") %>% 
  mutate(year = as.numeric(year)) %>% 
  left_join(area_df, by = "survey_area") %>% 
  select(-survey_area)

#---
# All Metrics Together
all_drivers <- bind_rows(
  list(
    clim_idx,
    landings_idx,
    sst_idx,
    zp_idx,
    ss_idx))

```




```{r}
#| label: driver-scaling

# In this chunk the drivers get reshaped to scale
# Scaling should cover the period of analysis 1982-2019
# SST and GSI (all index/anomaly fields) are not scaled


# Put the data in a wide form for scaling:
drivers_wide <- all_drivers %>% 
  filter(year >= 1982,
         year <= 2019) %>% 
  mutate(id = str_replace_all(str_c(area, "_", var), " ", "_")) %>% 
    select(-c(area, area_titles, var)) %>% 
    pivot_wider(names_from = id, values_from = value) %>% 
  column_to_rownames(var = "year")


# Scale the landings, not SST or GSI
to_scale   <- select(drivers_wide, ends_with("landings")) %>% scale() %>% as.data.frame()
dont_scale <- select(drivers_wide, -ends_with("landings"))

# Join back together
drivers_scaled <- bind_cols(dont_scale, to_scale) %>% 
  rownames_to_column(var = "year")


```


## Exploring Influence of Drivers on Size Spectra

The hypothesized drivers for size spectrum changes were evaluated using multiple regression analysis. Candidate models for each region were informed by the any observed temporal structures in the size spectra changes through time, and with the inclusion of important lag-effects of predictors.These modeling decisions were informed by the following steps performed prior to multiple regression analysis with the above drivers.


### Temporal Lags in Hypothesized Drivers

In addition to exploring their contemporaneous impacts, the importance of any time-lags on predictors was also evaluated. Important lagged relationships were identified through the use of cross correlation function (CCF) estimates. These estimates looked at correlations at one-year lag intervals to investigate potential time-delayed effects between the hypothesized drivers and the community’s size spectra. CCF were performed between the dependent variables of the size spectra slope with the independent variables of the annual gulf stream index, the corresponding regional commercial landings and sea surface temperature anomalies, and both small and large zooplankton indices. This was performed for annual lags up to 10 years. Significant lags  identified in the CCF analysis were then included as additional candidate predictors in the multiple regression analysis of spectrum slope changes.




### Non-Stationary Behavior Across Response Regimes

To highlight potential signatures of any non-stationary or non-linear discontinuous dynamics, a hallmark of regime-shift dynamics (Blocker 2023), we explored discontinuous changes in the size spectra using breakpoint analyses. Breakpoint analysis was performed on the annual time series of spectrum slopes to explore their temporal structures for evidence of any potential regime shifts. Breakpoint analyses were performed using the envcpt package (Killick et al. 2021). This package applies an automatic model selection process testing support for candidate model structures including constant/piecewise changes to the mean, variance, trends, & autocorrelations and any identified changepoints located using the pruned exact linear time algorithm (Rikardsen 2004). Best supported model structures from this procedure were then used to inform whether (if any) regional spectra expressed distinct periods of behavior. In the events where changepoints were identified, this was used as evidence that there may have been a community regime change. Following this thinking it is plausible for there to be inconsistencies or non-stationarity in the predictors' impact on the community spectra, as different community regimes may be differently vulnerable to top-down pressures like commercial fishing. To allow for changes in these relationships across different regimes, an additional interaction term was added to the regression model candidates for that region. 



### Regime Specific Correlations

Based on any breakpoints detected, single driver correlations were prepared to visually explore how each driver was correlated within the different regimes. Additionally, spectra slope changes were also correlated with the body-size changes of each functional group as a window into what body-size changes might be informing the change in the size structure.

```{r}
#| label: driver-correlation-periods

# Pull different years into a list
driver_periods <- list(
  "1982-1998"     = c(1982:1998),
  "1999-2006"     = c(1999:2006),
  "2007-2019"     = c(2007:2019),
  "1982-2019"     = c(1982:2019))


# -------------------------------------------------
# Idea: Add functional_group_sizes into the correlation chart:

# Reshape the size information so we can combine it with the drivers:
# Values are scaled here
sizes_wide <- functional_group_size %>% 
  select(year = Year, 
         survey_area, 
         hare_group, 
         avg_weight = mean_wt_kg) %>% 
  pivot_longer(names_to = "var", 
               values_to = "value", 
               cols = c(avg_weight)) %>% 
  inner_join(area_df, by = "survey_area") %>% 
  select(-survey_area) %>% 
  mutate(id = str_replace_all(str_c(area, "_", hare_group, "_", var), " ", "_")) %>% 
  select(-c(area, area_titles, var, hare_group)) %>% 
  pivot_wider(names_from = id, values_from = value, values_fn = {mean}) %>% 
  column_to_rownames(var = "year") %>% 
  scale() %>% 
  as.data.frame()



# Join regional size changes to the scaled drivers
drivers_and_sizes <- drivers_scaled %>% 
  left_join(rownames_to_column(sizes_wide, var = "year"), by = "year") %>%
  column_to_rownames(var = "year")


# Split them out into a list that contains data for each period
# Takes the years in driver_periods and pulls the rows that correspond
driver_period_list <- map(driver_periods, function(yrs){
  period_matrix <- drivers_and_sizes[rownames(drivers_and_sizes) %in% yrs, ] %>% drop_na()})

```




### Multiple Regression Modeling of Hypothesized Drivers

Following these explorations into temporal structures, multiple regression models were developed and evaluated independently for each region. Potential drivers included in the models were the annual Gulf Stream Index, the annual regionally averaged SST anomaly value, regionally averaged small and large copepod indices, and the scaled annual commercial landings from that region. Gulf stream index, zooplankton indices, and SST anomalies were not standardized prior to use in model fitting and evaluation as the two former are already an index value and the latter is scaled to a thirty-year climatological reference period. In the event that CCF analysis identified an important time lag for a region’s independent variables, these specific time-lagged predictors were included as predictors in the candidate models, otherwise time-lags on predictors were not evaluated. For any region(s) that showed support for changepoints in their temporal structure, the periods separated by those breaks were included as an interaction term on all other predictors. In the absence of any identified changepoints, no multi-year periods were included as potential regime impacts. All candidate models for a region were evaluated using AIC and AIC weights to rank the most parsimonious models for each region’s dynamics. 






```{r}
#| label: regression-analysis-data


# Use scaled drivers for the regressions
# Need to drop zooplankton because they are non-unique since we split the MAB EPU and repeated it
regression_df <- drivers_scaled %>% 
  select(!ends_with(c("zp_small","zp_large")))

# Build back out the metadata:
# Identify what region is associated with both: spectrum values & driver values
regression_df <- regression_df %>% 
  pivot_longer(names_to = "spectra_param", 
               values_to = "spectra_values", 
               cols = ends_with("slope") | ends_with("int")) %>% 
  pivot_longer(names_to = "driver_var", 
               values_to = "driver_values", 
               cols = -matches("spectra|year")) %>% 
  mutate(
    year = as.numeric(year),
    # A. Flag what the driver type was
    driver_type = case_when(
      str_detect(driver_var, "landings") ~ "Commercial Landings",
      str_detect(driver_var, "sst") ~ "SST",
      str_detect(driver_var, "index") ~ "GSI",
      TRUE ~ "Missed Something"),
    # B. Flag what the spectrum feature was
    param_feature = case_when(
      # Flag what the Size Distribution Parameter was
      str_detect(spectra_param, "int") ~ "Spectra Intercept",
      str_detect(spectra_param, "isd") ~ "ISD Exponent",
      str_detect(spectra_param, "slope") ~ "Spectra Slope",
      TRUE ~ "Missed Something"),
    # C. These are the areas associated with the Spectra Features
    spectra_area = case_when(
      str_detect(spectra_param, "All") ~ "All",
      str_detect(spectra_param, "Georges") ~ "Georges Bank",
      str_detect(spectra_param, "Gulf") ~ "Gulf of Maine",
      str_detect(spectra_param, "Southern") ~ "Southern New England",
      str_detect(spectra_param, "Mid-Atlantic") ~ "Mid-Atlantic Bight"),
    spectra_area = factor(spectra_area, levels = c("All", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight")),
    
    # D. These are the areas of the drivers
    driver_area = case_when(
      str_detect(driver_var, "All") ~ "All",
      str_detect(driver_var, "Georges") ~ "Georges Bank",
      str_detect(driver_var, "Gulf") ~ "Gulf of Maine",
      str_detect(driver_var, "Southern") ~ "Southern New England",
      str_detect(driver_var, "Mid-Atlantic") ~ "Mid-Atlantic Bight"),
    driver_area = factor(driver_area, levels = c("All", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight")),
    
    # E. Add Decade in
    decade = floor_decade(year),
    #
    #regime = ifelse(year < 2010, "regime_1", "regime_2" )
    ) 


# Filter it so predictors are only within matching regions, 
# or just overarching (GSI)
regression_df <- regression_df %>% 
  filter(driver_area == spectra_area | driver_type == "GSI") %>% 
  arrange(year, driver_var)




# Last wrangling:
# columns for: year, slope, isd, spectra region, drver_region, GSI, sst_anom, landings
# GSI is so annoying here, just rejoin it without the area column
regression_df <- regression_df %>% 
  select(-c(spectra_param, driver_var)) %>% 
  pivot_wider(names_from = "driver_type", values_from = "driver_values") %>% 
  pivot_wider(names_from = "param_feature", values_from = "spectra_values") %>% 
  select(-GSI) %>% 
  filter(driver_area != "All") %>% 
  left_join( select(clim_idx, year, GSI = value), by = c("year") ) %>% 
  mutate(spectra_area = fct_drop(spectra_area))




# Add zooplankton here so it can be checked in the regressions
# Zooplankton values are duplicated in MAB and SNE b/c its the same EPU
regression_df <- left_join(
    regression_df, 
    zp_index %>% rename(spectra_area = survey_area), 
    by = join_by(year, spectra_area))

```


# Results


```{r}
#| label: abundance stacked bars


 


# # Just abundance, not by groups
# strat_abundance_plot <- fgroup_area$weight_bins %>% 
#   filter(survey_area == "Northeast Shelf, Full") %>% 
#   group_by(Year, survey_area, hare_group) %>% 
#   summarise(abund_millions = sum(strat_abund_mill, na.rm = T),
#             .groups = "drop") %>% 
#   #filter(survey_area == "GoM") %>% 
#   ggplot(aes(Year, abund_millions, fill = hare_group)) +
#   #geom_area(color = "gray30", linewidth = 0.1) +
#   geom_col(color = "gray30", linewidth = 0.1, position = "stack") +
#   scale_fill_gmri() + 
#   scale_x_continuous(expand = expansion(add = c(0,0))) +
#   scale_y_continuous(labels = comma_format()) +
#   facet_wrap(~survey_area, ncol = 1, scales = "free") + 
#   labs(y = "Abundance (millions)",  
#        x = NULL,
#        fill = "Functional Group") +
#   theme(legend.position = "bottom")
# 
# 
# 
# 
# # Abundance Totals
# strat_abundance_plot

```


```{r}
#| label: biomass stacked bars

# 
# # Just Biomass, by area and functional group
# strat_biomass_plot <- fgroup_area$weight_bins %>% 
#   filter(survey_area == "Northeast Shelf, Full") %>% 
#   group_by(Year, survey_area, hare_group) %>% 
#   summarise(abund_millions = sum(strat_lwbio_mill, na.rm = T),
#             .groups = "drop") %>% 
#   ggplot(aes(Year, abund_millions, fill = hare_group)) +
#   # geom_area(color = "gray30", linewidth = 0.1) +
#   geom_col(color = "gray30", linewidth = 0.1, position = "stack") +
#   scale_fill_gmri() + 
#   scale_x_continuous(expand = expansion(add = c(0,0))) +
#   scale_y_continuous(labels = comma_format()) +
#   facet_wrap(~survey_area, ncol = 1, scales = "free") + 
#   labs(y = "Biomass (million kg)",  
#        x = NULL,
#        fill = "Functional Group") +
#   theme(legend.position = "bottom")
# 
# 
# strat_biomass_plot

```

### Community Abundance




```{r}
#| label: abundance and biomass summary numbers
#---- Overall Abundance Trends


# Total abundance and biomass for the whole shelf
shelf_summary <- catch_size_bins %>% 
  group_by(Year) %>% 
  summarise(
    lwbio_mill = sum(strat_total_lwbio_s, na.rm = T) / 1e6,
    abund_mill = sum(strat_total_abund_s, na.rm = T) / 1e6)

# Total Abundance and Biomass for the regions
region_summary <- catch_size_bins %>% 
  left_join(area_df) %>% 
  group_by(Year, area_titles) %>% 
  summarise(
    lwbio_mill = sum(strat_total_lwbio_s, na.rm = T) / 1e6,
    abund_mill = sum(strat_total_abund_s, na.rm = T) / 1e6,
    .groups = "drop") %>% 
  mutate(area_titles = factor(area_titles, levels = area_levels_long))


# Total abundance and biomass for functional Groups within the regions
functional_group_summary <- catch_size_bins %>% 
  left_join(area_df) %>% 
  group_by(Year, area_titles, hare_group) %>% 
  summarise(
    lwbio_mill = sum(strat_total_lwbio_s, na.rm = T) / 1e6,
    abund_mill = sum(strat_total_abund_s, na.rm = T) / 1e6,
    .groups = "drop") %>% 
  mutate(area_titles = factor(area_titles, levels = area_levels_long))
```



Stratified abundance estimates have been gradually increasing on the Northeast Shelf since 1970. Abundance growth accelerated after 2007 to peak levels in 2014. Abundance then began declining, which continued through 2019 (Figure 2).  


```{r}
#| label: abundance totals by area
#| eval: true
#| fig.width: 8
#| fig-height: 6

# Abundance 1.
shelf_abund_p <- shelf_summary %>% 
  mutate(area_titles = "Northeast Shelf, Full") %>% 
  ggplot() +
  geom_line(aes(Year, abund_mill), color = gmri_cols("dark gray"), linewidth = 1) +
  #geom_col(aes(Year, abund_mill), fill = gmri_cols("dark gray")) +
  scale_y_continuous(labels = label_comma()) +
  facet_wrap(~area_titles) +
  labs(y = "Abundance (millions)")



# Abundance 2.
region_abund_p <- region_summary %>% 
  ggplot() +
  geom_col(aes(Year, abund_mill, fill = area_titles), show.legend = F) +
  scale_fill_gmri() +
  scale_y_continuous(labels = label_comma()) +
  facet_wrap(~area_titles, ncol = 1, scales = "free") +
  labs(y = "Abundance (millions)")

shelf_abund_p | region_abund_p
```



In the Gulf of Maine, fish abundance remained relatively-low until the 1990’s, at which point it began to steadily rise–closely aligning with the pattern for the Northeast Shelf. This increase in abundance reversed briefly during 2002-2006, but continued to rise after 2006 before hitting a regional peak in 2016. Populations then declined through 2019 similar to the shelf-wide trend. Georges Bank abundances were consistently low from 1970 around 2010. By 2014 abundance had roughly quadrupled, propelled by strong recruitment classes of haddock. After 2014, abundance quickly fell to numbers more similar to the 1970-2010 levels by the end of the decade. Community abundance in Southern New England displayed higher inter-annual changes across all years compared to both Gulf of Maine and Georges Bank. Abundance in this area showed a less dramatic rise and fall than the Northern regions.  Abundances began increasing rapidly here in 2007, before falling back to earlier levels by the end of the 2010’s decade. The Mid Atlantic Bight displayed the most inter-annual variability and had no major trends in fish abundance.


#### Regional Abundances


::: {.panel-tabset}

### Northeast Shelf

```{r}
#| fig.width: 8
#| fig-height: 6


shelf_fgroup_summary <- catch_size_bins %>% 
  mutate(area_titles = "Northeast Shelf, Full") %>% 
  group_by(Year, area_titles, hare_group) %>% 
  summarise(
    lwbio_mill = sum(strat_total_lwbio_s, na.rm = T) / 1e6,
    abund_mill = sum(strat_total_abund_s, na.rm = T) / 1e6,
    .groups = "drop")


# Make Plot for Functional Groups
shelf_fgroup_abund_p <- shelf_fgroup_summary %>% 
  ggplot() +
  geom_col(aes(Year, abund_mill, fill = hare_group), show.legend = F) +
  scale_fill_gmri() +
  scale_y_continuous(labels = label_comma()) +
  facet_wrap(~hare_group, ncol = 1, scales = "free") +
  labs(y = "Abundance (millions)")


# Place Functional Groups Beside Shelf
shelf_abund_p | shelf_fgroup_abund_p

```


### Gulf of Maine

```{r}
#| fig.width: 8
#| fig-height: 6


area_option <- "Gulf of Maine"

plot_area_abund_summary <- function(area_option){
  
  area_abund_p <- region_summary %>% 
  filter(area_titles == area_option) %>% 
  ggplot() +
  geom_line(aes(Year, abund_mill), color = gmri_cols("dark gray"), linewidth = 1) +
  #geom_col(aes(Year, abund_mill), fill = gmri_cols("dark gray")) +
  scale_y_continuous(labels = label_comma()) +
  facet_wrap(~area_titles) +
  labs(y = "Abundance (millions)")
  

# Make Plot for Functional Groups
area_fgroup_abund_p <- functional_group_summary %>% 
  filter(area_titles == area_option) %>% 
  ggplot() +
  geom_col(aes(Year, abund_mill, fill = hare_group), show.legend = F) +
  scale_fill_gmri() +
  scale_y_continuous(labels = label_comma()) +
  facet_wrap(~hare_group, ncol = 1, scales = "free") +
  labs(y = "Abundance (millions)")


# Place Functional Groups Beside Shelf
p_out <- area_abund_p | area_fgroup_abund_p
return(p_out)
  
}

plot_area_abund_summary(area_option = "Gulf of Maine")
```


### Georges Bank


```{r}
#| fig.width: 8
#| fig-height: 6


plot_area_abund_summary(area_option = "Georges Bank")
```

### Southern New England

```{r}
#| fig.width: 8
#| fig-height: 6


plot_area_abund_summary(area_option = "Southern New England")
```


### Mid-Atlantic Bight

```{r}
#| fig.width: 8
#| fig-height: 6

plot_area_abund_summary(area_option = "Mid-Atlantic Bight")
```


:::



Groundfish species were the dominant functional group driving the abundance and biomass trends in the Gulf of Maine and Georges Bank, with the two southern regions showing a more balanced abundance distribution among the five functional groups (Figure 2). 



#### Total Abundance by Individual Body-Size

```{r}
#| label: bodymass biomass distributions
#| fig.width: 10
#| fig-height: 8

# panel plot the biomass by body size
plot_abundance_vertical <- function(p_area = "GoM", show_facets = F, ylab = "Abundance (million)"){
  
  p <- fgroup_area$weight_bins %>% 
    complete(Year, hare_group, survey_area, left_lim) %>% 
    filter(survey_area == p_area) %>% 
    ggplot(aes(Year, strat_abund_mill, fill = left_lim)) +
      geom_col(color = "gray30", linewidth = 0.1) +
      facet_grid(hare_group~survey_area, scales = "free") +
      scale_fill_distiller(
        palette = "RdYlGn", 
        direction = -1, 
        breaks = seq(0,14, 2), 
        labels = math_format(2^.x)) +
      guides(
        fill = guide_colorbar(
          title.position = "top", title.hjust = 0.5, barwidth = unit(4, "in"),
          frame.colour = "black", ticks.colour = "black")) +
      scale_y_continuous(labels = comma_format()) +
      labs(y = ylab,  
           x = NULL, 
           fill = "Minimum Individual Weight 2^x (g)") +
      theme(legend.position = "bottom",
            axis.text.x = element_text(angle = 45, hjust = 1),
            plot.margin = margin(t = 0, r = 0, b = 0, l = 2))
  if(show_facets == F){p <- p + theme(strip.text.y = element_blank(), strip.background.y = element_blank())}
  return(p)
  
}



# Build the pieces
shelf_abunds <- plot_abundance_vertical("Northeast Shelf, Full", show_facets = F)
gom_abunds   <- plot_abundance_vertical("GoM", F, ylab = NULL)
gb_abunds    <- plot_abundance_vertical("GB", F, ylab = NULL)
sne_abunds   <- plot_abundance_vertical("SNE", F, ylab = NULL)
mab_abunds   <- plot_abundance_vertical("MAB", T, ylab = NULL)

# patch together
( shelf_abunds | gom_abunds |gb_abunds |sne_abunds |gb_abunds |mab_abunds ) + plot_layout(guides = "collect")


```



### Community Biomass





```{r}
#| label: biomass totals by area
#| fig.width: 8
#| fig-height: 6

#----- Overall Biomass


# Repeat for Biomass
# Abundance 1.
shelf_bio_p <- shelf_summary %>% 
  mutate(area_titles = "Northeast Shelf, Full") %>% 
  ggplot() +
  geom_line(aes(Year, lwbio_mill), color = gmri_cols("dark gray"), linewidth = 1) +
  #geom_col(aes(Year, abund_mill), fill = gmri_cols("dark gray")) +
  scale_y_continuous(labels = label_comma()) +
  facet_wrap(~area_titles) +
  labs(y = "Biomass (million kg)")

# Abundance 2.
region_bio_p <- region_summary %>% 
  ggplot() +
  geom_col(aes(Year, lwbio_mill, fill = area_titles), show.legend = F) +
  scale_fill_gmri() +
  scale_y_continuous(labels = label_comma()) +
  facet_wrap(~area_titles, ncol = 1, scales = "free") +
  labs(y = "Biomass (million kg)")

shelf_bio_p | region_bio_p

```



Similar to abundance, the overall biomass was highest in the two northern regions, the Gulf of Maine and Georges Bank. Roughly half of the biomass sampled in these regions can be attributed to groundfish species, with the second largest contributions coming from elasmobranchs. Within the groundfish biomass, larger individuals \>2kg in particular, declined during the 70's and 80's in these regions, never truly recovering. Beginning in the 2000's there were signs that groundfish abundances were increasing as evidenced by increasing numbers of smaller individuals, however in both regions this trend appears to have reversed by the mid 2010's. Elasmobranch biomass increased steadily throughout the survey time period across all regions, with the exception of southern New England. This area showed large 5-10 year swings in biomass, but no clear long-term trend. Larger elasmobranch were rare in all regions except for a period spanning the late 70's through the early 90's isolated to Georges Bank. Demersal species biomass was highest in the Gulf of Maine, dwarfing their contributions in other regions. Their biomass declined in the 70's, was flat until the late 90's, remaining relatively high until declining in the late 2010's. Pelagic species biomass was low in all regions, and is unlikely to be representative of true biomass trends due to gear selectivity.





#### Regional Biomass



There was a distinct difference between Northern and Southern regions in the way biomass was distributed among the different functional groups. The largest contributors to biomass in the southern regions (southern New England & mid-Atlantic bight) was the elasmobranch community. While the northern regions (Gulf of Maine & Georges Bank) each had similar quantities of elasmobranch biomasses, there was also a comparable contribution of groundfish and in the Gulf of Maine there was a major component of demersal species as well.



::: {.panel-tabset}

### Northeast Shelf

```{r}
#| fig.width: 8
#| fig-height: 6

# Make Plot for Functional Groups
shelf_fgroup_bio_p <- shelf_fgroup_summary %>% 
  ggplot() +
  geom_col(aes(Year, lwbio_mill, fill = hare_group), show.legend = F) +
  scale_fill_gmri() +
  scale_y_continuous(labels = label_comma()) +
  facet_wrap(~hare_group, ncol = 1, scales = "free") +
  labs(y = "Biomass (million kg)")


# Place Functional Groups Beside Shelf
shelf_bio_p | shelf_fgroup_bio_p

```


### Gulf of Maine

```{r}
#| fig.width: 8
#| fig-height: 6


area_option <- "Gulf of Maine"

plot_area_bio_summary <- function(area_option){
  
  area_bio_p <- region_summary %>% 
  filter(area_titles == area_option) %>% 
  ggplot() +
  geom_line(aes(Year, lwbio_mill), color = gmri_cols("dark gray"), linewidth = 1) +
  scale_y_continuous(labels = label_comma()) +
  facet_wrap(~area_titles) +
  labs(y = "Biomass (million kg)")
  

# Make Plot for Functional Groups
area_fgroup_bio_p <- functional_group_summary %>% 
  filter(area_titles == area_option) %>% 
  ggplot() +
  geom_col(aes(Year, lwbio_mill, fill = hare_group), show.legend = F) +
  scale_fill_gmri() +
  scale_y_continuous(labels = label_comma()) +
  facet_wrap(~hare_group, ncol = 1, scales = "free") +
  labs(y = "Biomass (million kg)")


# Place Functional Groups Beside Shelf
p_out <- area_bio_p | area_fgroup_bio_p
return(p_out)
  
}

plot_area_bio_summary(area_option = "Gulf of Maine")
```


### Georges Bank


```{r}
#| fig.width: 8
#| fig-height: 6


plot_area_bio_summary(area_option = "Georges Bank")
```

### Southern New England

```{r}
#| fig.width: 8
#| fig-height: 6


plot_area_bio_summary(area_option = "Southern New England")
```


### Mid-Atlantic Bight

```{r}
#| fig.width: 8
#| fig-height: 6


plot_area_bio_summary(area_option = "Mid-Atlantic Bight")
```


:::



#### Total Biomass by Individual Body-Size

```{r}
#| label: bodymass abundance distributions
#| fig.width: 10
#| fig-height: 8

# panel plot the biomass by body size
plot_bodymass_vertical <- function(p_area = "GoM", show_facets = F, ylab = "Biomass (million kg)"){
  
  p <- fgroup_area$weight_bins %>% 
    complete(Year, hare_group, survey_area, left_lim) %>% 
    filter(survey_area == p_area) %>% 
    ggplot(aes(Year, strat_lwbio_mill, fill = left_lim)) +
      geom_col(color = "gray30", linewidth = 0.1) +
      facet_grid(hare_group~survey_area, scales = "free") +
      scale_fill_distiller(
        palette = "RdYlGn", 
        direction = -1, 
        breaks = seq(0,14, 2), 
        labels = math_format(2^.x)) +
      guides(
        fill = guide_colorbar(
          title.position = "top", title.hjust = 0.5, barwidth = unit(4, "in"),
          frame.colour = "black", ticks.colour = "black")) +
      scale_y_continuous(labels = comma_format()) +
      labs(y = ylab,  
           x = NULL, 
           fill = "Minimum Individual Weight 2^x (g)") +
      theme(legend.position = "bottom",
            axis.text.x = element_text(angle = 45, hjust = 1),
            plot.margin = margin(t = 0, r = 0, b = 0, l = 2))
  if(show_facets == F){p <- p + theme(strip.text.y = element_blank(), strip.background.y = element_blank())}
  return(p)
  
}



# Build the pieces
shelf_bmass <- plot_bodymass_vertical("Northeast Shelf, Full", show_facets = F)
gom_bmass   <- plot_bodymass_vertical("GoM", F, ylab = NULL)
gb_bmass    <- plot_bodymass_vertical("GB", F, ylab = NULL)
sne_bmass   <- plot_bodymass_vertical("SNE", F, ylab = NULL)
mab_bmass   <- plot_bodymass_vertical("MAB", T, ylab = NULL)

# patch together
( shelf_bmass | gom_bmass |gb_bmass |sne_bmass |gb_bmass |mab_bmass ) + plot_layout(guides = "collect")


```



### Body Size Trends

The average fish size in the Gulf of Maine (length and weight) declined the greatest of all regions over our study period. The average individual length was greatest in the 1970's in the 35-40cm range, falling to 28-33cm over the last decade. Body-weight fell dramatically in the 1980's, from around .75kg in the 1970's to .25-.30kg, roughly a third of what it had been. Georges Bank body sizes also declined during the study period, but less dramatically. Both of these Northern regions had brief period in the late 2000's where average length and weight rose, before falling again in the 2010's. The MAB region was the only region to see a long-term increase in both length and weight during the study period. SNE saw no long-term change in length, and a minor decline in average body-weight.

::: panel-tabset
#### All Species

```{r}
#| label: regional-size-trends
#| fig.height: 6

# Re-factor
regional_size <- regional_size %>% 
  filter(Year < 2020) %>% 
  mutate(
    survey_area = factor(survey_area, area_levels),
    area_copy = survey_area)


# Get the average length and weight at each decade
avg_size_decades <- regional_size %>% 
  mutate(
    decade_left = as.numeric(as.character(floor_decade(Year))),
    decade_right = as.numeric(as.character(floor_decade(Year+10))) - 1,
    decade_center = (decade_left + decade_right)/2,
    decade_lab = str_c(decade_left, "-", decade_right)) %>% 
  group_by(area_titles, decade_left, decade_right, decade_center, decade_lab) %>% 
  summarise(
    mean_len_cm = mean(mean_len_cm),
    mean_wt_kg = mean(mean_wt_kg),
    .groups = "drop"
  )


# Length plot
avg_len_p <-  regional_size %>% 
    ggplot() +
    geom_line(aes(Year, mean_len_cm, group = area_titles), 
              linewidth = 1,
              show.legend = F, color = "gray60") + 
  geom_segment(
    data = avg_size_decades,
    aes(x = decade_left, xend = decade_right, y = mean_len_cm, yend = mean_len_cm), linewidth = 1.5) +
  geom_text(
    data = avg_size_decades,
    aes(x = decade_center, y = mean_len_cm, 
        label = str_c(round(mean_len_cm, 1)) ), 
    vjust = -1.05, size = 4, fontface = "bold") +
    scale_color_gmri() +
    facet_wrap(~area_titles, ncol = 1) + 
    labs(title = "Average Length",
         y = "Length (cm)",
         color = "Region")


# Weight plot
avg_wt_p <- regional_size %>% 
  ggplot(aes(Year, mean_wt_kg)) + 
  geom_line(aes(group = area_titles), 
            linewidth = 1,
            show.legend = F,
            color = "gray60") + 
  geom_segment(
    data = avg_size_decades,
    aes(x = decade_left, xend = decade_right, y = mean_wt_kg, yend = mean_wt_kg), linewidth = 1.5) +
  geom_text(
    data = avg_size_decades,
    aes(x = decade_center, y = mean_wt_kg, 
        label = str_c(round(mean_wt_kg, 1)) ), 
    vjust = -1.05, size = 4, fontface = "bold") +
  scale_color_gmri() +
  facet_wrap(~area_titles, ncol = 1) + 
  labs(title = "Average Weight",
       y = "Weight (kg)",
       color = "Region")
  
# Plot side by side
(avg_len_p | avg_wt_p) 

```

#### Functional Group Avg. Length

```{r}
#| label: functional-group-length-trends
#| fig-height: 6

# Get the average length and weight at each decade
avg_fgroup_decades <- functional_group_size %>% 
  
  mutate(
    decade_left = as.numeric(as.character(floor_decade(Year))),
    decade_right = as.numeric(as.character(floor_decade(Year+10))) - 1,
    decade_center = (decade_left + decade_right)/2,
    decade_lab = str_c(decade_left, "-", decade_right)) %>% 
  group_by(area_titles, hare_group, decade_left, decade_right, decade_center, decade_lab) %>% 
  summarise(
    mean_len_cm = mean(mean_len_cm),
    mean_wt_kg = mean(mean_wt_kg),
    .groups = "drop")



# Plot the Length changes
functional_group_size %>% 
    ggplot() +
    geom_line(aes(Year, mean_len_cm, group = area_titles), 
              linewidth = 1,
              show.legend = F, color = "gray60") + 
  geom_segment(
    data = avg_fgroup_decades,
    aes(x = decade_left, xend = decade_right, y = mean_len_cm, yend = mean_len_cm), linewidth = 1.5) +
  geom_text(
    data = avg_fgroup_decades,
    aes(x = decade_center, y = mean_len_cm, 
        label = str_c(round(mean_len_cm, 1)) ), 
    vjust = -1.05, size = 3.5, fontface = "bold") +
    scale_color_gmri() +
    facet_grid(hare_group~area_titles, scales = "free_y") +  
    labs(title = "Average Length",
         y = "Length (cm)",
         color = "Region")







```

#### Functional Group Avg. Weight

```{r}
#| label: functional-group-weight-trends
#| fig-height: 6

# Average Weights
functional_group_size %>% 
    ggplot() +
    geom_line(aes(Year, mean_wt_kg, group = area_titles), 
              linewidth = 1,
              show.legend = F, color = "gray60") + 
  geom_segment(
    data = avg_fgroup_decades,
    aes(x = decade_left, xend = decade_right, y = mean_wt_kg, yend = mean_wt_kg), linewidth = 1.5) +
  geom_text(
    data = avg_fgroup_decades,
    aes(x = decade_center, y = mean_wt_kg, 
        label = str_c(round(mean_len_cm, 1)) ), 
    vjust = -1.05, size = 3.5, fontface = "bold") +
    scale_color_gmri() +
    facet_grid(hare_group~area_titles, scales = "free_y") +  
    labs(title = "Average Weight",
         y = "Weight (kg)",
         color = "Region")

```
:::

### Regional Size Spectra

At the start of our time series, back in the 1970's, there was a clear difference in the relative positions of spectra parameters among the different regions. Gulf of Maine and Georges Bank showed the least steep spectra slopes in the earlier time periods with slopes around -1 & -1.1 respectively. The relatively flat slopes in these regions both steepened over time, settling near -1.3 (GoM) and -1.5 (GB). Gulf of Maine experienced much of its decline during the 1980's and 1990's. There was a brief reversal in this trend during the 2000's, but slopes continued to steepen by 2010 and remained steep through 2019. Georges Bank did not experience as rapid of a decline, but experienced a similar long-term steepening. In contrast to the northern regions, SNE and MAB had steeper slopes in the -1.2 to -1.5 territory. The long term pattern for SNE was one of increasing volatility, but not so much a decline. The spectra slope for the MAB was less volatile, but similarly maintained a relatively stable wander around -1.4. By the end of the study period all regions had slopes that were at or near a similar level.

```{r}
#| label: size spectra results


# Make decadal averages of spectra slopes
indices_decades <- region_indices %>% 
  mutate(
    Year = as.numeric(Year),
    decade_left = as.numeric(as.character(floor_decade(Year))),
    decade_right = as.numeric(as.character(floor_decade(Year + 10))) - 1,
    decade_center = (decade_left + decade_right)/2,
    decade_lab = str_c(decade_left, "-", decade_right)) %>% 
  group_by(area_titles, decade_left, decade_right, decade_center, decade_lab) %>% 
  summarise( mean_b = mean(b, na.rm = T),
             mean_l2slope = mean(log2_slope_strat, na.rm = T), 
             .groups = "drop")

```



::: panel-tabset
#### Exponent of Size Spectra

```{r}
#| fig.height: 6

# 1. Exponent of ISD, MLE method
isd_timeline <- region_indices %>% 
  ggplot(aes(yr, b, group = area_titles)) +
    geom_line(aes(group = area_titles), color = "gray60", linewidth = 1) +
    #geom_pointrange(aes(group = area_titles, ymin = confMin, ymax = confMax), size = 0.1) +
    geom_segment(
      data = indices_decades,
      aes(x = decade_left, xend = decade_right, y = mean_b, yend = mean_b), linewidth = 1.5) +
    geom_text(
      data = indices_decades,
      aes(x = decade_center, y = mean_b, 
          label = str_c(round(mean_b, 1)) ), 
      vjust = -1.05, size = 3.5, fontface = "bold") +
    scale_y_continuous(expand = expansion(add = c(0,.1))) +
    scale_color_gmri() +
    facet_wrap(~area_titles, ncol = 1) +
    labs(x = "Year", y = "Abundance Size Spectrum Slope (b)")

isd_timeline 
#+ geom_smooth(color = "orange", se = FALSE)
```

#### Log2 Binned Size Spectra

```{r}
#| fig.height: 6

# 2. Plot the Regional Slopes, using the binned method
slope_timeline <- region_indices %>% 
  ggplot(aes(yr, log2_slope_strat, group = area_titles)) +
    geom_line(aes(group = area_titles), linewidth = 1, color = "gray70") +
    geom_segment(
      data = indices_decades,
      aes(x = decade_left, xend = decade_right, y = mean_l2slope, yend = mean_l2slope), linewidth = 1.5) +
    geom_text(
      data = indices_decades,
      aes(x = decade_center, y = mean_l2slope, 
          label = str_c(round(mean_l2slope, 1)) ), 
      vjust = -1.05, size = 3.5, fontface = "bold") +
    scale_y_continuous(expand = expansion(add = c(0,.1))) +
    scale_color_gmri() +
    facet_wrap(~area_titles, ncol = 1)  +
    labs(x = "Year", y = "Normalized Abundance Spectrum Slope") 

slope_timeline #+ geom_smooth(color = "orange", se = FALSE)
```

:::

## Size Spectra Drivers

### Lagged Impacts of Predictors

Exploratory analysis on potentially important predictor time lags through the use of cross correlation functions raised 6 potential lagged-driver candidates for the driver regression analyses. CCF identified potential time lags in Georges Bank predictors, and in the Mid-Atlantic Bight. Lagged predictors added to Georges Bank regression model selection process included:  A 2-year lag on the small zooplankton index, a 1-year lag on the large zooplankton index, a 1-year lag of SST, 4-year lag & 1-year lags of the GSI, and 4 & 5 year lags of commercial landings. Lagged predictors added to Mid-Atlantic Bight regression model selection process included: 5 & 7 year lags on the small zooplankton index and a 2-year lag on SST. 


```{r}
#| label: ccf-results
#| fig-height: 8
#| fig-width: 8

# Run CCF function to get lagged correlations,
# Put that data in one table to plot in a better comparison form


#------------ Processing Data for CCF. ------------------------
# Scale Everything Before CCF,
# Reshape the Drivers Wide to scale
# Scale the columns independently using the variance in all years, but within an area
# Are the columns scaling independently or together?
# Confirmed independently scaling


# Drivers scaled trimmed is a matrix of all the different timeseries
# Columns that end with slope or intercept are pulled out to get a column of the response variables (slope features)
# Columns that don't contain "spectra" or "year" capture the remaining features
# Next we pull out the region information for each
driver_ccf_prep <- drivers_scaled %>% 
  select(-All_sst_anom) %>% 
  #rownames_to_column(var = "year") %>% 
  pivot_longer(names_to = "spectra_param", 
               values_to = "spectra_values", 
               cols = ends_with("slope") | ends_with("int")) %>% 
  pivot_longer(names_to = "driver_var", 
               values_to = "driver_values", 
               cols = -matches("spectra|year")) %>% 
  mutate(
    # C. These are the areas associated with the Spectra Features
    spectra_area = case_when(
      str_detect(spectra_param, "All")          ~ "All",
      str_detect(spectra_param, "Georges")      ~ "Georges Bank",
      str_detect(spectra_param, "Gulf")         ~ "Gulf of Maine",
      str_detect(spectra_param, "Southern")     ~ "Southern New England",
      str_detect(spectra_param, "Mid-Atlantic") ~ "Mid-Atlantic Bight"),
    spectra_area = factor(spectra_area, 
      levels = c("All", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight")),
    
    # D. These are the areas of the drivers
    driver_area = case_when(
      str_detect(driver_var, "All")          ~ "All",
      str_detect(driver_var, "Georges")      ~ "Georges Bank",
      str_detect(driver_var, "Gulf")         ~ "Gulf of Maine",
      str_detect(driver_var, "Southern")     ~ "Southern New England",
      str_detect(driver_var, "Mid-Atlantic") ~ "Mid-Atlantic Bight"),
    driver_area = factor(driver_area, 
      levels = c("All", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight"))) 



# Filter it out so there is only cases where the driver area matches
driver_ccf_prep <- driver_ccf_prep %>% 
  filter(driver_area == spectra_area | driver_area == "All") %>% 
  arrange(year, driver_var)



# Walk through each xvariable, and see how it correlates with each yvar
# split on xvar
# then split on yvars
# make sure order is good
ccf_relationships <- driver_ccf_prep %>% 
  drop_na() %>% 
  split(.$spectra_param) %>% 
    map_dfr(function(x_param){
      x_param %>% 
        split(.$driver_var) %>% 
          map_dfr(function(driver_y_data){
            ccf_df <- get_ccf_vector(
              x = driver_y_data$spectra_values,
              y = driver_y_data$driver_values
            )
          }, .id = "driver_var")
  }, .id = "spectra_param")




# Build back out the labels for plotting
ccf_plot_data <- ccf_relationships %>% 
  mutate(
    # Flag what the driver type was
    driver_type = case_when(
      str_detect(driver_var, "landings") ~ "Commercial Landings",
      str_detect(driver_var, "sst") ~ "SST",
      str_detect(driver_var, "index") ~ "GSI",
      str_detect(driver_var, "small") ~ "ZP-S",
      str_detect(driver_var, "large") ~ "ZP-L",
      TRUE ~ "Missed Something"),
    param_feature = case_when(
      # Flag what the Size Distribution Parameter was
      str_detect(spectra_param, "int") ~ "Spectra Intercept",
      str_detect(spectra_param, "isd") ~ "ISD Exponent",
      str_detect(spectra_param, "slope") ~ "Spectra Slope",
      TRUE ~ "Missed Something"),
    spectra_region = case_when(
      # Flag what region the driver was coming from
      str_detect(spectra_param, "All") ~ "All",
      str_detect(spectra_param, "Georges") ~ "Georges Bank",
      str_detect(spectra_param, "Gulf") ~ "Gulf of Maine",
      str_detect(spectra_param, "Southern") ~ "Southern New England",
      str_detect(spectra_param, "Mid-Atlantic") ~ "Mid-Atlantic Bight"),
    # Flag when it crosses threshold
    sig_flag = ifelse(acf < signeg | acf > sigpos, T, F),
    # Set Factor Levels
    spectra_region = factor(spectra_region, levels = c("All", "Gulf of Maine", "Georges Bank", "Southern New England", "Mid-Atlantic Bight")))


# Limit to one Response:
ccf_plot_data <- filter(ccf_plot_data, param_feature == "ISD Exponent") %>% 
  filter(spectra_region != "All")

# Plot them
ccf_plot_data %>% 
  ggplot(aes(
    x = lag, 
    y = acf, 
    group = driver_var, 
    color = driver_type, 
    fill = driver_type)) +
  geom_col(alpha = 0.65) +
  geom_col(
    data = filter(ccf_plot_data, sig_flag),
    aes(lag, acf, fill = driver_type,  group = driver_var), color = "black") +
  geom_text(
    data = filter(ccf_plot_data, sig_flag),
    aes(lag, y = 0, label = lag, vjust = ifelse(acf < 0, 1.5,-1.5)), color = "black") +
  geom_line(aes(x = lag, y = sigpos), linetype = 2, color = "gray25") +
  geom_line(aes(x = lag, y = signeg), linetype = 2, color = "gray25") +
  geom_hline(yintercept = 0, color = "gray25", linewidth = 1) +
  geom_vline(xintercept = 0, color = "black", linewidth = 1) +
  scale_color_gmri() +
  scale_fill_gmri() +
  facet_grid(spectra_region~driver_type*param_feature) +
  scale_x_continuous(limits = c(-15, 15), breaks = seq(from = -15, to = 12, by = 1)) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 8)) +
  labs(x = "Driver Lag", y = "Correlation (CCF)", 
       fill = "Driver Type:", color = "Driver Type:",
       caption = "All predictors scaled over year range of 1982-2019. Lags exceeding significance threshold outlined and numbered.")

```





### Spectra Slope Changepoints


Breakpoint analyses showed support for ecological regime change in one region, the Gulf of Maine. Evidence of breakpoints were found in 1999 & 2007 – when slope values increased briefly, before reversing course and falling further. The best-supported breakpoint model structure included breaks in the linear trend at 1999 & 2007, and a 2-year autocorrelation term. Suggesting that the community spectra behaved differently over 3 regimes with strong autocorrelation to its state two years prior. For Georges Bank spectra slopes showed a single trend structure, with slopes steepening over time. This was also the case for the Mid-Atlantic Bight, but with slopes becoming more shallow with time. In Southern New England a single mean (intercept) model best reflected the lack of change. With no strong support for either a long-term trend or break points. Based on these exploratory results, three regime periods were added to the Gulf of Maine’s dataset for multiple regression analysis. The three regimes were 1982-1998, 1999-2006, 2007-2019. These regimes were included as an interaction term with each of the original hypothesized drivers, permitting potentially different influences of the drivers within each regime. The two-year lag on the spectrum slope was also included based on its support in the most-parsimonious changepoint analysis model. No breakpoints or autocorrelation terms were added to models for the other regions.


```{r}
#| label: regional changepoint analysis

# Pull Regions to test independently:
regression_df <- regression_df %>% 
  rename(
    landings = `Commercial Landings`,
    b = `ISD Exponent`)
gom_df <- filter(regression_df, spectra_area == "Gulf of Maine") 
gb_df <- filter(regression_df, spectra_area == "Georges Bank") %>% drop_na()
sne_df <- filter(regression_df, spectra_area == "Southern New England") 
mab_df <- filter(regression_df, spectra_area == "Mid-Atlantic Bight") %>% drop_na()


# Check each for different changepoint structures
gom_cpt <- envcpt(gom_df$b, verbose = F)
gb_cpt  <- envcpt(gb_df$b, verbose = F)
sne_cpt <- envcpt(sne_df$b, verbose = F)
mab_cpt <- envcpt(mab_df$b, verbose = F)


# Pulling out details of best supported model
cpt_res <- list(
  "gom" =  list("cpt_res" = gom_cpt),
  "gb"  =  list("cpt_res" = gb_cpt),
  "sne" =  list("cpt_res" = sne_cpt),
  "mab" =  list("cpt_res" = mab_cpt)) %>% map(function(x){
    best <- names(which.min(AIC(x$cpt_res)))
    best_summ <- x$cpt_res[[best]]
    x$best <- best
    x$best_summ <- best_summ
    return(x)
})


# # Plot GOM, the only one with changepoints
# plot(gom_cpt, main = "Gulf of Maine, Break Point Analysis")



# Put the rest of the results into a delta AIC table:
bind_rows(map(cpt_res, ~tibble("Most Supported Model" = .x[["best"]])), .id = "Region") %>% kable()


```


Changepoint analyses suggest there were as many as 3 distinct regimes in the Gulf of Maine slope changes. For all other regions there was not strong support for changepoints. 


## Spectra Slope Regression Analysis

# WORKING HERE

fix the coefficients so the interactions now reflect the net result

Fix it so delta AIC is actually delta AICc


::: panel-tabset

### Gulf of Maine

Model rankings using AICc & delta-AICc (Figure 5.) for the Gulf of Maine’s size spectrum slope models best support a regression model containing two predictors with a stationary effect across all years, and two predictors whose effects across were allowed to vary across the regime periods identified by the changepoint analysis (non-stationary effects). The two stationary drivers impacts were the negative impact of the small zooplankton index (p = 0.011) and a negative impact from the two-year autocorrelation term highlighted by the exploratory changepoint analysis (p = 0.033). The inclusion of an interaction term between the three changepoint analysis regimes (1982-1998, 1999-2006, & 2007-2019) and both the commercial landings index and the large zooplankton index allowed these drivers to differently impact size spectrum slope within the three regimes. Over those three periods the impact of landings during the first regime was positive (p = 0.009), the 1999-2006 regime was negative (p = 0.032), and during the third regime there was no relationship (p = 0.088). For the large zooplankton index during the first two regimes there was no significant impact (p = 0.2, p = >0.9), however during the third regime large zooplankton had a positive effect on the spectrum slope (p = <0.001). None of the terms for the regime periods themselves were significant  in the final model. Models containing only stationary effects on the predictors were not retained by model selection. This agreed with visual assessments of their predictive performance and residual trends.

```{r}
#| label: gom-standard-models

# change na. action for dredge
options(na.action = "na.fail")


# Incorporate the patterns from the changepoint analysis
# What if we enforce the break_periods
gom_df <- mutate(gom_df, 
  regime = case_when(
    year < 1999 ~ "yrs_1982-1998",
    year < 2007 ~ "yrs_1999_2006",
    year >= 2007 ~ "yrs_2007_2019"
  ))


# Add the autoregressive predictors neatly/manually
gom_df_lag <- gom_df %>% 
  mutate(
    blag1 = dplyr::lag(b, 1),
    blag2 = dplyr::lag(b, 2),
    zpslag1 = dplyr::lag(zp_small, 1),
    zpslag2 = dplyr::lag(zp_small, 2)) %>% 
  drop_na(blag2)


# kitchen sink Autoregressive model
gom_ar_models = lm(
  b ~ 
    regime * SST + 
    regime * GSI + 
    regime * landings + 
    regime * zp_small + 
    regime * zp_large + 
    regime * blag1 + 
    regime * blag2, 
  data = gom_df_lag)



# Model selection on models with lags at 1+2 years
results_gom <- dredge(gom_ar_models)


# Best autoregressive model
best_gom <- get.models(results_gom, subset = delta == 0)[[1]]

```



```{r}
#| label: gom-top-models-table

# Okay quick issues, for non-stationary effects, maybe just take the models under delta = 2
# and from them pull the summary

# Pull details for models with delta AIC < 2
top_mods_gom <- get.models(results_gom, subset = delta < 2)

# Peformance, Gets the AIC order 
top_perf <- map_dfr(top_mods_gom, glance, .id = "id") %>% arrange(AIC)
top_perf$Delta <- top_perf$AIC - min(top_perf$AIC)
top_perf <- top_perf %>% 
  arrange(Delta) %>% 
  mutate(`Model ID` = row_number(),
         `Model ID` = str_c("Model ", `Model ID`))



# Top Model Coefficients, join in the model number based on AIC order
top_coef <- map_dfr(top_mods_gom, tidy, .id = "id") %>% 
  left_join(
    select(top_perf, id, `Model ID`)) %>% 
  select(-id)

# Reshape Performance Specs to bind_row later
top_perf <- select(top_perf, `Model ID`, r2 = r.squared, Delta) %>% 
  pivot_longer(cols = -1, names_to = "term", values_to = "estimate") %>% 
  mutate(fill_color = "lightgray")






# Clean up the data and put it into a graph
gom_table_data <- top_coef %>% 
  # Tidy up the names and set the color scheme for fill
  mutate(
    term = ifelse(term == "landings", "Commercial Landings", term),
    term = ifelse(term == "zp_small", "Small Zooplankton", term),
    term = ifelse(term == "zp_large", "Large Zooplankton", term),
    term = ifelse(term == "blag2", "2-yr-AR", term),
    term = str_remove_all(term, "regime"),
    term = str_remove_all(term, "yrs_"),
    term = str_replace_all(term, ":", " ~ "),
    fill_color = ifelse(estimate<0, "#EACA00", "#00736D")) %>%
  
  # Join in the Performance Details and set Factor Orders
  bind_rows(top_perf) %>%  
  mutate(
    survey_area = "Gulf of Maine",
    `Model ID` = fct_rev(`Model ID`),
    estimate = round(estimate, 2)) %>% 
  
  # Set the different types of estimates here
  mutate(
    type = case_when(
      term ==  "1999_2006" ~ "Regime\nIntercept",
      term ==  "2007_2019" ~ "Regime\nIntercept",
      term ==  "2-yr-AR" ~ "Fixed\nEffects",
      term ==  "r2" ~ "Model\nPerformance",
      term ==  "Delta" ~ "Model\nPerformance",
      str_detect(term," ~ ") ~ "Regime\nInteraction",
      TRUE ~ "Fixed\nEffects"),
    type = factor(type, 
      levels = c("Fixed\nEffects", "Regime\nIntercept", "Regime\nInteraction", "Model\nPerformance"))) 



```

::: panel-tabset






#### Model Summary

```{r}
#| label: gom regression summary
# output summary - broom.helpers and gtsummary

# Best Model
tbl1 <- best_gom %>% 
  tbl_regression() %>% 
  add_glance_table() %>% 
  bold_labels()

tbl1

```

#### VIF

```{r}
#| label: gom vif
# make a model with all of them
all_parms <- lm(
  b ~ 
    regime * SST + 
    regime * GSI + 
    regime * landings + 
    regime * zp_small + 
    regime * zp_large, 
  data = gom_df)

# Check VIF
vif(all_parms, type = "predictor") %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "Predictor") %>% 
  gt() 

```

#### Actual vs. Fitted

```{r}
#| label: gom fit
augment(best_gom, gom_df_lag ) %>% 
# augment(best_gom, gom_df) %>% 
  ggplot(aes(year)) +
  geom_line(aes(y = b, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = .fitted, color = "Predicted"), linewidth = 1) +
  scale_color_gmri() +
  labs(color = "Color", y = "Spectrum Slope")
```


#### Residual Autocorrelation

```{r}
#| label: gom residuals
augment(best_gom, gom_df_lag ) %>% 
  ggplot(aes(year, y = .resid)) +
  geom_line(linewidth = 1, color = "gray30") +
  geom_point(size = 2, color = "gray30") +
  geom_hline(yintercept = 0, linetype = 2, color = "darkred", linewidth = 1) +
  scale_color_gmri() +
  labs(color = "Color", y = "Model Residuals")
```
:::


### Georges Bank

The best supported model for Georges Bank retained two predictors: sea surface temperature anomalies (p = 0.028) and the small zooplankton index at a 2-year lag (p = 0.006). With both the SST anomalies and the lagged small zooplankton index having negative effects on size spectrum slope. Five candidate models were within a delta-AIC range of <2. These top-models retained only bottom-up drivers as the best predictors, usually as some pair containing the small zooplankton at 0-2 year lags and either SST anomalies or the Gulf Stream Index. Suggesting that this region’s size spectrum changes are most highly correlated with environmental forces and not commercial landings.

```{r}
#| label: gb models

# Georges Bank showed
gb_df_lag <- gb_df %>% 
  mutate(
    b = b,
    blag1 = dplyr::lag(b, 1),
    blag2 = dplyr::lag(b, 2),
    zpslag1 = dplyr::lag(zp_small, 1),
    zpslag2 = dplyr::lag(zp_small, 2)) %>% 
  drop_na(blag2)



# fit model with all parameters
all_parms <- lm(b ~ 
                  SST + 
                  GSI + 
                  landings + 
                  zp_small + 
                  zp_large + 
                  zpslag1 + 
                  zpslag2, 
                data = gb_df_lag)

# Refine the dredge
results <- dredge(all_parms)

#grab best model
# subset(results, delta == 0)
bmod <- get.models(results, subset = delta == 0)[[1]]

```




```{r}
#| label: gb-top-models-table


# Pull details for models with delta AIC < 2
top_mods_gb <- get.models(results, subset = delta < 2)

# Peformance, Gets the AIC order 
top_perf <- map_dfr(top_mods_gb, glance, .id = "id") %>% arrange(AIC)
top_perf$Delta <- top_perf$AIC - min(top_perf$AIC)
top_perf <- top_perf %>% arrange(Delta) %>% 
  mutate(`Model ID` = row_number(),
         `Model ID` = str_c("Model ", `Model ID`))



# Top Model Coefficients, join in the model number based on AIC order
top_coef <- map_dfr(top_mods_gb, tidy, .id = "id") %>% 
  left_join(
    select(top_perf, id, `Model ID`)) %>% 
  select(-id)

# Reshape Performance Specs to bind_row later
top_perf <- select(top_perf, `Model ID`, r2 = r.squared, Delta) %>% 
  pivot_longer(cols = -1, names_to = "term", values_to = "estimate") %>% 
  mutate(fill_color = "lightgray")


# Clean up the data and put it into a graph
gb_table_data <- top_coef %>% 
  # Tidy up the names and set the color scheme for fill
  mutate(
    term = ifelse(term == "landings", "Commercial Landings", term),
    term = ifelse(term == "zp_small", "Small Zooplankton", term),
    term = ifelse(term == "zp_large", "Large Zooplankton", term),
    term = ifelse(term == "zpslag1", "Small Zooplankton 1-yr Lag", term),
    term = ifelse(term == "zpslag2", "Small Zooplankton 2-yr Lag", term),
    fill_color = ifelse(estimate<0, "#EACA00", "#00736D")) %>%
  
  # Join in the Performance Details and set Factor Orders
  bind_rows(top_perf) %>%  
  mutate(
    survey_area = "Georges Bank",
    `Model ID` = fct_rev(`Model ID`),
    estimate = round(estimate, 2)) %>% 
  
  # Set the different types of estimates here
  mutate(
    type = case_when(
      term ==  "r2" ~ "Model\nPerformance",
      term ==  "Delta" ~ "Model\nPerformance",
      str_detect(term,":") ~ "Regime\nInteraction",
      TRUE ~ "Fixed\nEffects"),
    type = factor(type, 
      levels = c("Fixed\nEffects", "Model\nPerformance"))) 


```

::: panel-tabset




#### Model Summary

```{r}
#| label: gb regression summary
# output summary - brrom.helpers and gtsummary

# Best Model
tbl1 <- bmod %>% 
  tbl_regression() %>% 
  add_glance_table() %>% 
  bold_labels()

tbl1

```

#### VIF

```{r}
#| label: gb vif
all_parms <- lm(b ~ 
                  SST + 
                  GSI + 
                  landings + 
                  zp_small + 
                  zp_large + 
                  zpslag1 + 
                  zpslag2, 
                data = gb_df_lag)

vif(all_parms) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "Predictor") %>% 
  rename(VIF = ".") %>% 
  arrange(VIF) %>% 
  gt() 

```

#### Actual vs. Fitted

```{r}
#| label: gb fit
augment(bmod, gb_df_lag) %>% 
  ggplot(aes(year)) +
  geom_line(aes(y = b, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = .fitted, color = "Predicted"), linewidth = 1) +
  scale_color_gmri() +
  labs(color = "Color", y = "Spectrum Slope")
```

#### Residual Autocorrelation

```{r}
#| label: gb residuals
augment(bmod, gb_df_lag) %>% 
  ggplot(aes(year, y = .resid)) +
  geom_line(linewidth = 1, color = "gray30") +
  geom_point(size = 2, color = "gray30") +
  geom_hline(yintercept = 0, linetype = 2, color = "darkred", linewidth = 1) +
  scale_color_gmri() +
  labs(color = "Color", y = "Model Residuals")
```
:::



### Southern New England

Exploratory analysis for Southern New-England was suggestive of a lack of trend in size spectrum slopes. This was further confirmed by the model selection process. The best supported model of Southern New England retained only commercial landings, however this relationship was not significant (p = 0.13) and had very low performance (r-squared = 0.06). 

```{r}
#| label: sne models

# fit model with all parameters
all_parms <- lm(b ~ SST + GSI + landings + zp_small + zp_large, data = sne_df)

# Refine the dredge
results <- dredge(all_parms)

#grab best model
# subset(results, delta == 0)
bmod <- get.models(results, subset = delta == 0)[[1]]

```



```{r}
#| label: sne-top-models-table



# Pull details for models with delta AIC < 2
top_mods_sne <- get.models(results, subset = delta < 2)

# Peformance, Gets the AIC order 
top_perf <- map_dfr(top_mods_sne, glance, .id = "id") %>% arrange(AIC)
top_perf$Delta <- top_perf$AIC - min(top_perf$AIC)
top_perf <- top_perf %>% arrange(Delta) %>% 
  mutate(`Model ID` = row_number(),
         `Model ID` = str_c("Model ", `Model ID`))



# Top Model Coefficients, join in the model number based on AIC order
top_coef <- map_dfr(top_mods_sne, tidy, .id = "id") %>% 
  left_join(
    select(top_perf, id, `Model ID`)) %>% 
  select(-id)

# Reshape Performance Specs to bind_row later
top_perf <- select(top_perf, `Model ID`, r2 = r.squared, Delta) %>% 
  pivot_longer(cols = -1, names_to = "term", values_to = "estimate") %>% 
  mutate(fill_color = "lightgray")


# Clean up the data and put it into a graph
sne_table_data <- top_coef %>% 
  # Tidy up the names and set the color scheme for fill
  mutate(
    term = ifelse(term == "landings", "Commercial Landings", term),
    term = ifelse(term == "zp_small", "Small Zooplankton", term),
    term = ifelse(term == "zp_large", "Large Zooplankton", term),
    fill_color = ifelse(estimate<0, "#EACA00", "#00736D")) %>%
  
  # Join in the Performance Details and set Factor Orders
  bind_rows(top_perf) %>%  
  mutate(
    survey_area = "Southern New England",
    `Model ID` = fct_rev(`Model ID`),
    estimate = round(estimate, 2)) %>% 
  
  # Set the different types of estimates here
  mutate(
    type = case_when(
      term ==  "r2" ~ "Model\nPerformance",
      term ==  "Delta" ~ "Model\nPerformance",
      str_detect(term,":") ~ "Regime\nInteraction",
      TRUE ~ "Fixed\nEffects"),
    type = factor(type, 
      levels = c("Fixed\nEffects", "Model\nPerformance"))) 

```

::: panel-tabset




#### Model Summary

```{r}
#| label: sne regression summary
# output summary - brrom.helpers and gtsummary

# Best Model
tbl1 <- bmod %>% 
  tbl_regression() %>% 
  add_glance_table() %>% 
  bold_labels()

tbl1

```

#### VIF

```{r}
#| label: sne vif
all_parms <- lm(b ~ 
                  SST + 
                  GSI + 
                  landings + 
                  zp_small + 
                  zp_large, 
                data = sne_df)


vif(all_parms) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "Predictor") %>% 
  rename(VIF = ".") %>% 
  arrange(VIF) %>% 
  arrange(VIF) %>% 
  gt() 

```

#### Actual vs. Fitted

```{r}
#| label: sne fit
augment(bmod, sne_df ) %>% 
  ggplot(aes(year)) +
  geom_line(aes(y = b, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = .fitted, color = "Predicted"), linewidth = 1) +
  scale_color_gmri() +
  labs(color = "Color", y = "Spectrum Slope")
```

#### Residual Autocorrelation

```{r}
#| label: sne residuals
augment(bmod, sne_df ) %>% 
  ggplot(aes(year, y = .resid)) +
  geom_line(linewidth = 1, color = "gray30") +
  geom_point(size = 2, color = "gray30") +
  geom_hline(yintercept = 0, linetype = 2, color = "darkred", linewidth = 1) +
  scale_color_gmri() +
  labs(color = "Color", y = "Model Residuals")
```
:::



### Mid-Atlantic Bight

Three top models were selected for the Mid-atlantic Bight region, all with non-stationary predictors. The best supported model showed that increases in the small zooplankton index had a negative impact on spectrum slope (p = 0.020). This result was present in the other top candidates, which each also included a negative correlation to large zooplankton or with commercial landings. Model performance was low among the top models (r-squared 0.14-0.16).


```{r}
#| label: mab models

# fit model with all parameters
all_parms <- lm(b ~ SST + GSI + landings + zp_small + zp_large, data = mab_df)

# Refine the dredge
results <- dredge(all_parms)

#grab best model
# subset(results, delta == 0)
bmod <- get.models(results, subset = delta == 0)[[1]]

```




```{r}
#| label: mab-top-models-table



# Pull details for models with delta AIC < 2
top_mods_mab <- get.models(results, subset = delta < 2)

# Peformance, Gets the AIC order 
top_perf <- map_dfr(top_mods_mab, glance, .id = "id") %>% arrange(AIC)
top_perf$Delta <- top_perf$AIC - min(top_perf$AIC)
top_perf <- top_perf %>% arrange(Delta) %>% 
  mutate(`Model ID` = row_number(),
         `Model ID` = str_c("Model ", `Model ID`))



# Top Model Coefficients, join in the model number based on AIC order
top_coef <- map_dfr(top_mods_mab, tidy, .id = "id") %>% 
  left_join(
    select(top_perf, id, `Model ID`)) %>% 
  select(-id)

# Reshape Performance Specs to bind_row later
top_perf <- select(top_perf, `Model ID`, r2 = r.squared, Delta) %>% 
  pivot_longer(cols = -1, names_to = "term", values_to = "estimate") %>% 
  mutate(fill_color = "lightgray")


# Clean up the data and put it into a graph
mab_table_data <- top_coef %>% 
  # Tidy up the names and set the color scheme for fill
  mutate(
    term = ifelse(term == "landings", "Commercial Landings", term),
    term = ifelse(term == "zp_small", "Small Zooplankton", term),
    term = ifelse(term == "zp_large", "Large Zooplankton", term),
    fill_color = ifelse(estimate<0, "#EACA00", "#00736D"),
    fill_color = ifelse(term == "(Intercept)", "lightgray", fill_color)) %>%
  
  # Join in the Performance Details and set Factor Orders
  bind_rows(top_perf) %>%  
  mutate(
    survey_area = "Mid-Atlantic Bight",
    `Model ID` = fct_rev(`Model ID`),
    estimate = round(estimate, 2)) %>% 
  
  # Set the different types of estimates here
  mutate(
    type = case_when(
      term ==  "r2" ~ "Model\nPerformance",
      term ==  "Delta" ~ "Model\nPerformance",
      str_detect(term,":") ~ "Regime\nInteraction",
      TRUE ~ "Fixed\nEffects"),
    type = factor(type, 
      levels = c("Fixed\nEffects", "Model\nPerformance"))) 


```


::: panel-tabset



#### Model Summary

```{r}
#| label: mab regression summary
# output summary - brrom.helpers and gtsummary

# Best Model
tbl1 <- bmod %>% 
  tbl_regression() %>% 
  add_glance_table() %>% 
  bold_labels()

tbl1

```

#### VIF

```{r}
#| label: mab vif
all_parms <- lm(b ~ SST + GSI + landings + zp_small + zp_large, 
                data = mab_df)

vif(all_parms) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "Predictor") %>% 
  rename(VIF = ".") %>% 
  arrange(VIF) %>% 
  gt() 

```

#### Actual vs. Fitted

```{r}
#| label: mab fit
augment(bmod, mab_df ) %>% 
  ggplot(aes(year)) +
  geom_line(aes(y = b, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = .fitted, color = "Predicted"), linewidth = 1) +
  scale_color_gmri() +
  labs(color = "Color", y = "Spectrum Slope")
```

#### Residual Autocorrelation

```{r}
#| label: mab residuals
augment(bmod, mab_df ) %>% 
  ggplot(aes(year, y = .resid)) +
  geom_line(linewidth = 1, color = "gray30") +
  geom_point(size = 2, color = "gray30") +
  geom_hline(yintercept = 0, linetype = 2, color = "darkred", linewidth = 1) +
  scale_color_gmri() +
  labs(color = "Color", y = "Model Residuals") 
```

:::



:::


### Model Selection Results

```{r}
#| label: model selection table
#| fig-height: 10
#| fig-width: 8

bind_rows(
  list(
    gom_table_data,
    sne_table_data,
    mab_table_data,
    gb_table_data)) %>% 
  mutate(
    survey_area = factor(survey_area, levels = area_levels_long),
    `Model ID` = factor(`Model ID`, levels = str_c("Model ", c(5:1))))%>% 
ggplot( aes(term, `Model ID`)) +
  geom_tile(aes(fill = I(fill_color)), color = "gray20", linewidth = 0.2, alpha = 0.75) +
  geom_text(aes(label = estimate)) +
  facet_grid(survey_area ~ type, scales = "free", space = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + 
  labs(
    x = "Model Predictor",
    y = "",
    title = "Spectrum Slope Driver Exploration",
    subtitle = "Interaction terms should be added to the corresponding fixed effect to determine net-influence."
  )
```


# Discussion



### Potential Drivers Timeseries:

The following panels show the historical changes in each of the drivers. Landings have been scaled by average total landings within each region across all years. SST and GSI have not been scaled. This is different from how they are implemented in the regression analyses, when the landings were scaled over the 1982-2019 period.

::: panel-tabset
#### Surface Temperature Anomalies

```{r}
#| label: temperature-index
#| fig.height: 8


# Plot the timelines
temp_regimes %>% 
  mutate(anom_direction = ifelse(sst_anom > 0, "Positive", "Negative")) %>% 
  ggplot( aes(yr, sst_anom)) +
  geom_col(aes(fill = anom_direction),  
           linewidth = 0.75, alpha = 0.8) +
  geom_hline(yintercept = 0, linewidth = 0.5, lty = 1) +
  scale_x_continuous(expand = expansion(add = c(0.25,0.25))) +
  scale_y_continuous(labels = number_format(suffix = " \u00b0C")) +
  scale_fill_gmri() +  
  facet_wrap(~survey_area, ncol = 1) +
  theme(
    legend.title = element_blank(),
    legend.background = element_rect(fill = "transparent"),
    legend.key = element_rect(fill = "transparent", color = "transparent")) + 
  labs(
    title = "NW-Atlantic Regional SST Anomalies",
    x = "", 
    y = "Surface Temperature Anomaly",
    caption = "Anomalies calculated using 1982-2011 reference period.") +
  guides(fill = "none",
         label = "none",
         color = guide_legend(keyheight = unit(0.5, "cm")))
```

#### GARFO Landings

```{r}
#| label: landings-index
#| fig.height: 8

####  Landings  ####
landings_summ %>% 
  mutate(anom_direction = ifelse(total_wt_z > 0, "Positive", "Negative")) %>% 
  ggplot(aes(year, total_wt_z, group = survey_area)) +
  geom_col(aes(fill = anom_direction),  
           linewidth = 0.75, alpha = 0.8) +
  geom_hline(yintercept = 0, linewidth = 0.5, lty = 1) +
  scale_fill_gmri() +  
  facet_wrap(~survey_area, ncol = 1) +
  scale_x_continuous(breaks = seq(1950, 2030, by = 10)) +
  labs(x = "Year", y = "Finfish Landings Index (z)",
       title = "GARFO Finfish Landings")  +
  guides(fill = "none")
```

#### Climate Modes

```{r}
#| label: climate-modes

####  Climate Drivers  ####
clim_drivers %>% 
  group_by(year = lubridate::year(Time), Var, EPU) %>% 
  summarise(Value = mean(Value, na.rm = T)) %>% 
  mutate(anom_direction = ifelse(Value > 0, "Positive", "Negative"),
         Var = str_to_title(Var)) %>% 
  ggplot(aes(year, Value, group = EPU)) +
  geom_col(aes(fill = anom_direction),  
           linewidth = 0.75, alpha = 0.8) +
  geom_hline(yintercept = 0, linewidth = 0.5, lty = 1) +
  scale_fill_gmri() +  
  facet_wrap(~Var, ncol = 1) +
  scale_y_continuous(labels = scales::comma_format()) +
  scale_x_continuous(limits = c(1960, 2020), 
                     breaks = seq(1950, 2030, by = 10)) +
  labs(x = "Year", y = "Index",
       title = "Environmental Drivers") +
  guides(fill = "none")
```

#### Zooplankton Indices


```{r}
#| fig.height: 8

zp_index %>% 
  pivot_longer(cols = starts_with("zp"), 
               names_to = "Var", values_to = "Value") %>% 
  group_by(year, Var, survey_area) %>% 
  summarise(Value = mean(Value, na.rm = T)) %>% 
  mutate(anom_direction = ifelse(Value > 0, "Positive", "Negative"),
         Var = ifelse(
           Var == "zp_large", 
           "Large Zooplankton", 
           "Small Zooplankton"),
         survey_area = factor(survey_area, levels = area_levels_long)) %>% 
  ggplot(aes(year, Value, group = survey_area)) +
  geom_col(aes(fill = anom_direction),  
           linewidth = 0.75, alpha = 0.8) +
  geom_hline(yintercept = 0, linewidth = 0.5, lty = 1) +
  scale_fill_gmri() +  
  facet_grid(survey_area~Var) +
  scale_y_continuous(labels = scales::comma_format()) +
  scale_x_continuous(limits = c(1960, 2020), 
                     breaks = seq(1950, 2030, by = 10)) +
  labs(x = "Year", y = "Index",
       title = "Environmental Drivers") +
  guides(fill = "none")
```


:::



# Supplemental Materials





::: panel-tabset
## Species Functional Groups

```{r}
#| label: tbl-functional-groups
#| tbl-cap: "Common and scientific names for the species that constitute each functional group used in our analyses. X markers are used to indicate which regions each species has been caught in the data."


# Tidy up a key of what is labeled what
spec_key <- catch_size_bins %>% 
  distinct(comname, scientific_name, hare_group) %>% 
  mutate(
    hare_group = ifelse(hare_group %in% c("elasmobranch", "groundfish"), hare_group, str_c(hare_group)),
    hare_group = ifelse(is.na(hare_group), "Not Classified", hare_group),
    hare_group = str_to_title(hare_group),
    hare_group = fct_drop(hare_group),
    hare_group = factor(hare_group, levels = c("Coastal", "Diadromous", "Elasmobranch", "Groundfish",  "Pelagic", "Reef", "Not Classified"))) %>% 
  group_by(hare_group) %>% 
  mutate(n = str_c("(",n(),")")) %>% 
  ungroup()


# Reformat the data to check presence absence
catch_size_bins %>% 
  distinct(comname, survey_area) %>% 
  mutate(Presence = "X") %>% 
  complete(comname, survey_area) %>% 
  left_join(spec_key, by = "comname") %>% 
  mutate(comname = str_to_title(comname),
         Presence = ifelse(is.na(Presence) == FALSE, "X", ""),
         Presence = ifelse(is.na(Presence) == TRUE, "", Presence)) %>% 
  left_join(area_df, by = "survey_area") %>% 
  select(hare_group, n, comname, scientific_name, area, Presence) %>% 
  pivot_wider(names_from = area, values_from = Presence) %>% 
  arrange(hare_group, comname) %>% 
  group_by(hare_group,n) %>% 
  gt()  %>% 
  gt::tab_header(title = md("**Functional Group Assignments and Regional Presence/Absence**")) %>% 
  tab_stubhead(label = md("**Functional Group**")) %>% 
  cols_label(
    comname = md("*Common Name*"),
    scientific_name = md("*Scientific Name*"),
    `Gulf of Maine` = md("*Gulf of Maine*"),
    `Georges Bank` = md("*Georges Bank*"),
    `Southern New England` = md("*Southern New England*"),
    `Mid-Atlantic Bight` = md("*Mid-Atlantic Bight*")
    ) %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_row_groups()) %>% 
  gt::tab_source_note(source_note = md("*Functional group assignments adapted from Hare et al. 2010*"))

```

## GARFO Landings Summary

```{r}
#| label: garfo-landings-summary

# Add the labels into the landings data and remove what we don't need there:
landings %>% 
  filter(between(year, 1960, 2019),
         !str_detect(sppname, "CONFIDENTIAL")) %>% 
  mutate(
    decade = floor_decade(year),
    sppname = str_to_title(sppname)) %>% 
  group_by(survey_area, decade,  sppname) %>% 
  summarise(avg_landings_lb = mean(`landed lbs`, na.rm = T),
            across(.cols = c(landings_lb = `landed lbs`, 
                             value = value), 
                   .fns = sum, 
                   .names = "total_{.col}"),
            .groups = "drop") %>% 
  group_by(survey_area, decade) %>% 
  slice_max(order_by = total_landings_lb, n = 3) %>% 
  gt() %>% 
  gt::tab_header(title = md("**Top Commercial Fisheries Landings of Northeastern US (by weight)**")) %>% 
  tab_stubhead(label = md("*Harvest Region*")) %>% 
  fmt_number(columns = c(avg_landings_lb, total_landings_lb, total_value),
             use_seps = T, 
             sep_mark = ",",
             suffixing = T) %>% 
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_row_groups()) %>% 
  cols_label(
    decade            = md("*Decade*"),
    sppname           = md(""),
    avg_landings_lb   = md("*Avg. Annual Landings (lb.)*"),
    total_landings_lb = md("*Total Landings (lb.)*"),
    total_value       = md("*Total Value ($)*")) %>%
  gt::tab_source_note(source_note = md("*Landings data obtained from the Greater Atlantic Regional Fishing Office (GARFO)*"))
  
```




:::



### Single Driver Correlations


```{r}
#| label: period-correlations-processing

# Run Correlations for Each period and for the entire period
driver_period_correlations <- map(driver_period_list, function(drivers){
  
  #------ debug copy-paste line
  # Step 1. Get the correlation matrix (data is pre-scaled)
  corr_matrix <- rcorr(as.matrix(drivers))
  
  # R-squares
  rho_df <- as.data.frame(corr_matrix$r) %>%
    rownames_to_column(var = "var1") %>%
    pivot_longer(cols = -var1, 
                 names_to = "var2", 
                 values_to = "r2")

  # Significance Dataframe
  sig_df <- as.data.frame(corr_matrix$P) %>%
    rownames_to_column(var = "var1") %>%
    pivot_longer(cols = -var1, 
                 names_to = "var2", 
                 values_to = "pval")
  
  # Join
  corr_both <- left_join(rho_df, sig_df, by = c("var1", "var2")) 


  # Step 2. Tidy Up for ggplot
  # Make correlation matrix & significance a dataframe for GGplot
  # Goal: Flag the different Functional Group size changes somehow
  
  
  # Mega Cleanup Code Chunk
  # This first step adds metadata, but mostly creates a way to filter out correlations to be in the direction we want. no sst as response variables for example
  # GOAL: All the drivers in one column, and all the responses in another, order doesn't matter
  corr_tidy <- corr_both %>% 
    mutate(
    
    # 1.These flag whether or not the matrix variables in each column are the independent or dependent variables
    is_var1_driver   = ifelse(str_detect(var1, "landings|index|anom|avg|zp"), T, F),
    is_var2_response = ifelse(str_detect(var2, "slope|int"), T, F),
    
    # 2. Column to Flag what type of driver it is:
    driver_type = case_when(
        str_detect(var1, "landings")       ~ "Top-Down",
        str_detect(var1, "stream|anom|zp") ~ "Bottom-Up",
        str_detect(var1, "length")         ~ "Avg. Length",
        str_detect(var1, "weight")         ~ "Avg. Weight"),
    
    # 3. Column to Flag what area the driver data is from
    driver_area = case_when(
        str_detect(var1, "All")          ~ "Northeast Shelf, Full",
        str_detect(var1, "Georges")      ~ "Georges Bank",
        str_detect(var1, "Gulf")         ~ "Gulf of Maine",
        str_detect(var1, "Southern")     ~ "Southern New England",
        str_detect(var1, "Mid-Atlantic") ~ "Mid-Atlantic Bight"),
    
    # 4. Set Factor Levels for plot
      driver_area = factor(
        driver_area, 
        levels = c(
          "All", 
          "Gulf of Maine", 
          "Georges Bank", 
          "Southern New England", 
          "Mid-Atlantic Bight")),
      
    
    # 5. Shorthand Names for the Spectra Parameters - Response
    response_short = case_when(
      str_detect(var2, "int")   ~ "Spectra Intercept",
      str_detect(var2, "isd")   ~ "ISD Exponent",
      str_detect(var2, "slope") ~ "Spectra Slope"),
    
    
    # 6. These are the areas of the drivers
    response_area = case_when(
      str_detect(var2, "All")          ~ "All",
      str_detect(var2, "Georges")      ~ "Georges Bank",
      str_detect(var2, "Gulf")         ~ "Gulf of Maine",
      str_detect(var2, "Southern")     ~ "Southern New England",
      str_detect(var2, "Mid-Atlantic") ~ "Mid-Atlantic Bight"),
    
    
    # 7. Set levels for the driver areas
    response_area = factor(
      response_area, 
      levels = c(
        "All", 
        "Gulf of Maine", 
        "Georges Bank", 
        "Southern New England", 
        "Mid-Atlantic Bight")),
    
    # 8. Flag when the areas are the same or not:
    area_match = driver_area == response_area
  )
  

# Now we filter rows out to focus in on:
# 1. one direction driver & response
# 2. response and driver in same area, unless its GSI
corr_tidy_pairs <- corr_tidy %>% 
  filter( 
      is_var2_response,
      is_var1_driver,
      area_match | str_detect(var1, "index")) 
  
  
# From here it is just text formatting for plots:
# Because we are faceting we can have repeated names to simplify things
corr_tidy_pairs <- corr_tidy_pairs %>% 
  mutate(
      var1 = ifelse(str_detect(var1, "landings"), "Commercial_Landings", var1),
      var1 = ifelse(str_detect(var1, "sst"), "SST_Anomalies", var1),
      var1 = ifelse(str_detect(var1, "gulf_stream_index"), "Gulf Stream Index", var1),
      var1 = ifelse(str_detect(var1, "zp_small"), "Small Zooplankton", var1),
      var1 = ifelse(str_detect(var1, "zp_large"), "Large Zooplankton", var1),
      var1 = str_replace_all(var1, "_", " "),
      
      # Make a functional group column
      hare_group = case_when(
        str_detect(var1, "elasmobranch") ~ "Elasmobranch",
        str_detect(var1, "pelagic")      ~ "Pelagic",
        str_detect(var1, "coastal")      ~ "Coastal",
        str_detect(var1, "diadromous")   ~ "Diadromous",
        str_detect(var1, "groundfish")   ~ "Groundfish",
        str_detect(var1, "reef")         ~ "Reef",
        TRUE ~ ""),
      
      # Add the functional group column ahead of the average length and weight columns
      var1 = ifelse(str_detect(var1, "avg"), hare_group, var1),
      
      # Set The Factor Levels groups
      var1 = factor(
        var1,
        levels = c(
          "Gulf Stream Index",
          "SST Anomalies",
          "Small Zooplankton",
          "Large Zooplankton",
          "Groundfish",
          "Elasmobranch",
          "Coastal",
          "Diadromous",
          "Pelagic",
          "Reef",
          "Commercial Landings")),
      
      # Flag Significance Levels
      signif = ifelse(pval < 0.05, "*", ""))

  return(corr_tidy_pairs)
  
  #------- 

  
  
  
})

```

::: panel-tabset
#### All Years

```{r}
#| label: correlations-all
#| fig-height: 6
#| fig-width: 8

period <- "1982-2019"
plot_period_correlations(driver_period_correlations[[period]], period)
```

#### 1982-1998

```{r}
#| label: correlations-80s
#| fig-height: 6
#| fig-width: 8

period <- "1982-1998"
plot_period_correlations(driver_period_correlations[[period]], period)
```

#### 1999-2006

```{r}
#| label: correlations-90s
#| fig-height: 6
#| fig-width: 8

period <- "1999-2006"
plot_period_correlations(driver_period_correlations[[period]], period)
```

#### 2007-2019

```{r}
#| label: correlations-20s
#| fig-height: 6
#| fig-width: 8

period <- "2007-2019"
plot_period_correlations(driver_period_correlations[[period]], period)
```

:::




```{r}
#| label: weisberg-synthesis
#| eval: false


# Stuff for Sara:


# 1. Plot of Gulf of Maine Slopes, no gray lines
sara_plot <- region_indices %>% 
  filter(survey_area == "GoM") %>% 
  ggplot(aes(yr, b, group = area_titles)) +
  # geom_line(aes(group = area_titles), linewidth = 1) +
  geom_point(size = 1) +
  geom_smooth(method = 'loess', formula = 'y ~ x', color = "orange") +
  #geom_pointrange(aes(group = area_titles, ymin = confMin, ymax = confMax), size = 0.25) +
  scale_color_gmri() +
  facet_wrap(~area_titles, ncol = 1) +
  labs(x = "Year", y = "Abundance Size Spectrum Slope (b)", title = "Exponent of Abundance Size Spectra")

# The decline in weight of groundfish to go with:
# Average individual weight
functional_group_size %>% 
  filter(survey_area == "GoM", hare_group %in% c("groundfish", "elasmobranch")) %>% 
  ggplot(aes(Year, mean_wt_kg)) +
  geom_point(size = 1) + 
    geom_smooth(color = "orange", method = "loess", linewidth = 1) +
    facet_grid(hare_group~area_titles, scales = "free_y") + 
    theme(axis.text = element_text(size = 8)) +
    labs(title = "Average Weight",
         y = "Weight (kg)",
         color = "Region")

# How much biomass is just gf and elsmo:

# Plot abundance by region and Group
fgroup_area$weight_bins %>% 
  filter(survey_area == "GoM") %>% 
  ggplot(aes(Year, strat_lwbio_mill, fill = hare_group)) +
    geom_col(color = "gray30", linewidth = 0.1, position = "fill") +
    facet_grid(.~survey_area) +
    scale_fill_gmri() +
    scale_y_continuous(labels = percent_format()) +
    labs(
      y = "Fraction of Biomass",  
      x = NULL, 
      fill = "Functional Group") +
    theme(
      legend.position = "bottom",
      axis.text.x = element_text(angle = 45, hjust = 1))


# Table of averages over different periods for all of them:
region_indices %>%
  filter(yr %in% c(1970:2000)) %>%
  mutate(year_range = "1970-2000") %>%
  group_by(year_range, area_titles) %>%
  summarise(
    avg_slope = mean(b, na.rm = T),
    sd_slope = sd(b, na.rm = T)
  )

```

# References
