---
title: "Stratification and Aggregation Validation"
author: "Adam A. Kemberling"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float:
        collapsed: FALSE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, comment = NA)
options(knitr.kable.NA = '')

###__ Packages  ####
library(here)
library(janitor)
library(gmRi)
library(patchwork)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(ggforce)



# Load the build code and stratification function
box_paths <- research_access_paths(os.use = "unix")
mills_path <- box_paths$mills
res_path <- box_paths$res

#### Set theme  ####
theme_set(theme_minimal())

```

`r use_gmri_style_rmd(css_file = "gmri_rmarkdown.css")`

# About

The goal of this markdown is to validate consistency of outcomes for various aggregation scales. The approach is to take 4 years from the survey data, 2006-2010 to cover the change in survey design. Then we will track comparisons between estimating biomasses at a station level or in aggregate.

## Starting Data

```{r}
####__  Data  ####

# 2020 Raw data
load(here("data/NEFSC/Survdat_Nye_Aug 2020.RData"))
trawl_20 <- survdat %>% clean_names()
rm(survdat)

# And filter
trawl_20 <- filter(trawl_20, est_year >=2006 & est_year <= 2010)
```

## Clean up Function

The size spectra clean up function does any data prep including:   

 * Removal of incomplete records   
 * Spatial filters   
 * Pairing species to growth coefficients   
 * Calculating stratum effort and area weighting coefficients   
 * Estimating stratified catch rates   
 

```{r}
source(here("R/01_nefsc_ss_build.R"))
weights_20 <- load_ss_data(survdat = trawl_20, survdat_source = "tester")

```

**NOTE:**

The cleanup function retains the station-level information of the raw data, so its possible to compare individual stations across the raw and prepped data:

```{r}
test_station <- weights_20$id[1]
test_species <- unique(weights_20$comname)[4]

print(paste0("The test station is: ", test_station))
print(paste0("The test species is: ", test_species))

show_comparison <- function(x, test_station, test_species, names = "original"){
  if(names == "original"){
    x %>% 
      mutate(id = as.character(id)) %>% 
      filter(id == test_station, 
             comname == toupper(test_species)) %>% 
      select(id, comname, length, numlen, biomass, abundance,)
    } else{
    x %>% 
      filter(id == test_station, 
             comname == tolower(test_species)) %>% 
      select(id, comname, length, numlen, numlen_adj, biom_adj, abundance)
  }

}
t_raw <- show_comparison(trawl_20, test_station, test_species, names = "original")
t_clean <- show_comparison(weights_20, test_station, test_species, names = "cleaned")


bind_rows(list("raw data" = t_raw, "cleaned data" = t_clean), .id = "Data Source") %>% 
  mutate_if(is.numeric, round, 1) %>% 
  mutate_if(is.character, tolower) %>% 
  kable() %>%  
  kable_styling(font_size = 10)
```

This example is good because it shows how abundance does not always equal what the sum of the column `numlen` returns. `numlen_adj` is then an adjustment to convert the numbers found in `numlen` to total the observed "abundance".


# Raw Comparisons

Looking at the weights dataset we prepared, the first obvious comparison to make is how to weights from length weight relationships compare to the bulk biomass measurements recorded on the ship. In the below plots its clear that they do not match up, and in these years the aggregate biomass measure is almost 3x the l-w counterparts, with an inconsistent relationship in the fall.

```{r}
# Annual Summary
ann_summs <- weights_20 %>% 
  group_by(est_year) %>% 
  summarise(`L-W Biomass` = sum(sum_weight_kg),
            `Bulk Biomass` = sum(biom_adj),
            `Area Stratified Total L-W` = sum(expanded_lwbio_s),
            `Area Stratified Total BIOMASS` = sum(expanded_biom_s),
            season = "SPRING + FALL")

# Seasonal Differences
seas_summs <- weights_20 %>% 
  group_by(est_year, season) %>% 
  summarise(`L-W Biomass` = sum(sum_weight_kg),
            `Bulk Biomass` = sum(biom_adj),
            `Area Stratified Total L-W` = sum(expanded_lwbio_s),
            `Area Stratified Total BIOMASS` = sum(expanded_biom_s))


# reformat for plotting
summaries_long <- bind_rows(ann_summs, seas_summs) %>% 
  pivot_longer(names_to = "Biomass Source", values_to = "Total Survey Biomass (kg)", 
               cols = c(`L-W Biomass`, `Bulk Biomass`)) %>% 
  pivot_longer(names_to = "Stratified Biomass Source", values_to = "Area Projected Biomass (kg)", 
               cols = c(`Area Stratified Total L-W`,`Area Stratified Total BIOMASS`)) %>% 
  mutate(season = factor(season, levels = c("SPRING + FALL", "SPRING", "FALL")))


# Total Biomass from survey
ggplot(summaries_long, aes(est_year, `Total Survey Biomass (kg)`, color = `Biomass Source`)) +
  geom_line() +
  facet_wrap(~season, ncol = 1) +
  scale_y_continuous(labels = comma_format())+
  scale_color_gmri() +
  labs(x = "") +
  theme(legend.position = "bottom")


# What that biomass is extended out to total areas
ggplot(summaries_long, aes(est_year, `Area Projected Biomass (kg)`, color = `Stratified Biomass Source`)) +
  geom_line() +
  facet_grid(season~`Stratified Biomass Source`) +
  scale_y_continuous(labels = comma_format())+
  scale_color_gmri() +
  labs(x = "") +
  theme(legend.position = "bottom")

```



# Aggregation Comparisons {.tabset .tabset-pills}

Another comparison of interest is whether it makes a difference to calculate catch rates for individual lengths of a species, expanding those rates out to get a stratified biomass. Or to aggregate before estimating catch rates.

The following code will look into that comparison.



## Size Specific Catch Rates {.tabset}

The way the size spectra build code is written catch rates for each length caught are estimated and weighted for every stratum in which they are caught. Weighting is adjusted by the amount of effort within that stratum in that year, and the relative size of the stratum to the overall sampling coverage that year.

In this way, using the example above there are catch per tow estimates for American Plaice of lengths 15, 16, and 30 cm.

```{r}
weights_20 %>% 
    filter(id == test_station, 
           comname == tolower(test_species)) %>% 
      select("Species"                                      = comname, 
             "Length"                                       = length, 
             "Number at Length"                             = numlen_adj, 
             "Weight (g)"                                   = sum_weight_kg, 
             "Stratum"                                      = stratum,
             "Tows in Stratum (by year)"                    = strat_ntows, 
             "Abundance per Tow"                            = abund_tow_s, 
             "Weighted Abundance per Tow (by stratum area)" = wt_abund_s,
             "Projected Abundance (stratum)"                = expanded_abund_s) %>% 
  mutate(`Weight (g)` = `Weight (g)` * 1000) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  kable() %>%  
  kable_styling(font_size = 10, )
```

The question is whether the sum of these "Projected Abundances" is the same  when these catch rates are done in aggregate using the following function:

```{r}
agg_species_metrics
```


### Aggregate After

```{r}
cod_20 <- weights_20 %>% filter(comname == "atlantic cod")

# adding up stratified weights at the end
cod_20 %>% 
  group_by(est_year, season, comname) %>% 
  summarise(stratified_abundance = sum(expanded_abund_s),
            stratified_lwbio     = sum(expanded_lwbio_s),
            stratified_biom      = sum(expanded_biom_s)) %>% 
  kable() %>% 
  kable_styling(font_size = 10)
```

### Aggregate First 

```{r}

# Aggregating then calculating the catch rates etc.
cod_20 %>% 
  agg_species_metrics(est_year, season) %>% 
  select(Year                   = est_year, 
         Season                 = season, 
         Species                = comname, 
         `Stratified Abundance` = strat_abundance_s,
         `Stratified LWbio`       = lw_strat_biomass_s,
         `Stratified Biom`        = fscs_strat_biomass_s) %>% 
  kable() %>% 
  kable_styling(font_size = 10)
```

# LW Coefficient Comparisons {.tabset .tabset-pills}

Looks like the root cause is just the difference between what the mean prediction from species-specific coefficients is when compared to the biomass column in survdat.

Going to plot out how each species compares individually and rank their fits.


```{r, fig.height=10}
# Thoughts - difference between lw biomass vs. time 
# (i.e. is their relationship changing)
# how does that relate to temperature and survey change'
# good way to flag species that are way off

# Load all years and species
weights_all <- load_ss_data(survdat_source = "2020")


# Get the total weights for each station, then do it each year for every species
mismatch_ranks <- function(x){
  species_summaries <- x %>% 
    group_by(est_year, comname, id) %>% 
    summarise(number_caught = sum(numlen_adj),
              length_weight_biomass = sum(sum_weight_kg),
              ship_measured_biomass = sum(biom_adj),
              lw_underestimate = ship_measured_biomass - length_weight_biomass) %>% 
    ungroup()
  
  
  # Format rankings to plot
  n_species <- length(unique(species_summaries$comname))
  ranks <- species_summaries %>% 
    mutate(comname = fct_reorder(.f = comname, .x = lw_underestimate, mean),
           species_rank = as.numeric(comname),
           extremes = ifelse( species_rank >= n_species - 15, "Extreme Under-Estimation", NA),
           extremes = ifelse(species_rank <= 15, "Extreme Over-Estimation", extremes)) 
  
  return(ranks)
  
}


# Plot flags top and bottom 15 species that don't match well with survdat$biomass
mismatch_plot <- function(x){
  # All ranks in one panel
  all_ranks <- x %>% 
    ggplot(aes(lw_underestimate, species_rank)) +
      geom_boxplot(aes(group = species_rank), color = "gray40") +
      geom_boxplot(data = filter(x, is.na(extremes) == FALSE),
                   aes(group = species_rank, color = extremes)) +
      geom_vline(xintercept = 0, size = 0.5, color = "black", linetype = 2) +
      scale_x_continuous(labels = comma_format()) +
      scale_color_gmri() +
      guides("color" = guide_legend(title = "")) +
      labs(x = "Underestimation of Biomass from LW Regressions\nShipboard Weight - LW Biomass (kg)",
           y = "Underestimation Rank") +
    theme(legend.position = c(0.2, 1))
  
  # Focus on top 5
  top_5 <- x %>% 
    filter(extremes == "Extreme Under-Estimation") %>% 
    ggplot(aes(lw_underestimate, comname)) +
      geom_boxplot(aes(group = species_rank), color = gmri_cols("orange")) +
      #geom_vline(xintercept = 0, size = 0.5, color = "black", linetype = 2) +
      scale_color_gmri() +
      scale_x_continuous(labels = comma_format()) +
      labs(x = "",
           y = "", 
           subtitle = "L-W mass < 'survdat$Biomass'") +
    theme(title = element_text(hjust = 0.5))
  
  # Focus on bottom 5
  bot_5 <- x %>% 
    filter(extremes == "Extreme Over-Estimation") %>% 
    ggplot(aes(lw_underestimate, comname)) +
      geom_boxplot(aes(group = species_rank), color = gmri_cols("gmri blue")) +
      #geom_vline(xintercept = 0, size = 0.5, color = "black", linetype = 2) +
      scale_color_gmri() +
      scale_x_continuous(labels = comma_format()) +
      labs(x = "Underestimation of Biomass from LW Regressions\nShipboard Weight - LW Biomass (kg)",
           y = "", 
           subtitle = "L-W mass > 'survdat$Biomass'") +
    theme(title = element_text(hjust = 0.5))
    
  
  
  # Patchwork plot
  plot_out <- all_ranks | (top_5/bot_5)
  return(plot_out)
}



```


## SVSPP Species

These species got a priority pass when matching with L-W coefficients because they are from a more recent and region specific source.

```{r, fig.height=8}
# Get ranks for everything
all_source_ranks <- weights_20 %>%
  mismatch_ranks() 

# Just Wigley
wigley_ranks <- weights_all %>% filter(source == "wigley") %>% 
  mismatch_ranks() 

wigley_ranks %>% 
  mismatch_plot()
```


## Fishbase Species

These species were matched up with coefficients from fishbase and are likely less accurate than the ones with svspp codes.

```{r, fig.height=8}
fishbase_ranks <- weights_all %>% filter(source == "fishbase") %>% 
  mismatch_ranks()

fishbase_ranks %>% 
  mismatch_plot()
```


## LW Regression Fit Diagnostics {.tabset}


### Subsample Errors

Take the species that stood out above and plot the mismatch versus the station abundances. This should highlight whether its an issue of subsampling. Might be beneficial to  split it for seasons for dogfish and some other species.


```{r, fig.height=10}


#species that are caught in large numbers
subsampling_suspects <- c("spiny dogfish", "haddock", "blueback herring", "american shad", "windowpane", "silver rag", "atlantic cod", "pollock", "smooth dogfish", "acadian redfish")



# plot
all_source_ranks %>% 
  filter(comname %in% subsampling_suspects) %>% 
  ggplot(aes(number_caught, lw_underestimate)) +
      geom_point(color = gmri_cols("gmri blue")) +
      geom_smooth(method = "lm", color = gmri_cols("orange")) +
      facet_wrap(~comname, scales = "free", ncol = 2) +
      labs(x = "Single Station Abundance", y = "survdat$BIOMASS - LW Biomass")
  
```


### Mismatch Outliers

Dig into the cases where things are really off by flagging the outliers and plotting those relationships.

```{r, fig.height=10}

# Flagging outliers using ruler
library(ruler)


#### Outlier Rules ####


#Z-score
isnt_out_z <- function(x, thres = 3, na.rm = TRUE) {
  abs(x - mean(x, na.rm = na.rm)) <= thres * sd(x, na.rm = na.rm)
}


# Median Absolute Deviation https://en.wikipedia.org/wiki/Median_absolute_deviation
isnt_out_mad <- function(x, thres = 3, na.rm = TRUE) {
  abs(x - median(x, na.rm = na.rm)) <= thres * mad(x, na.rm = na.rm)
}



# Tukeys Fences - box plot outliers https://en.wikipedia.org/wiki/Outlier#Tukey&%2339;s_fences
isnt_out_tukey <- function(x, k = 1.5, na.rm = TRUE) {
  quar <- quantile(x, probs = c(0.25, 0.75), na.rm = na.rm)
  iqr <- diff(quar)
  
  (quar[1] - k * iqr <= x) & (x <= quar[2] + k * iqr)
}


# Mahalanobis Distance https://en.wikipedia.org/wiki/Mahalanobis_distance
maha_dist <- . %>% select_if(is.numeric) %>%
    mahalanobis(center = colMeans(.), cov = cov(.))
isnt_out_maha <- function(tbl, isnt_out_f, ...) {
  tbl %>% maha_dist() %>% isnt_out_f(...)
}



# Definition of non-outlier row
isnt_out_funs <- funs(
  z = isnt_out_z,
  mad = isnt_out_mad,
  tukey = isnt_out_tukey
)


#Implementation
all_source_ranks <- all_source_ranks %>% 
  split(.$comname) %>% 
  map_dfr(~ .x %>% 
  mutate(abundance_outliers = isnt_out_tukey(number_caught),
         estimation_outliers = isnt_out_tukey(lw_underestimate))
)


# Plot High Catch Outliers
all_source_ranks %>% 
  #Filter to just outliers
  filter(abundance_outliers == F,
         comname %in% subsampling_suspects)%>% 
  ggplot(aes(number_caught, lw_underestimate)) +
      geom_point(color = gmri_cols("gmri blue")) +
      geom_smooth(formula = y ~ x, method = "lm", color = gmri_cols("orange")) +
      facet_wrap(~comname, scales = "free", ncol = 2) +
      labs(x = "Single Station Abundance", y = "survdat$BIOMASS - LW Biomass",
           caption = "Data shown are outliers in station abundance by species.")


# Plot the estimation outliers
# Plot High Catch Outliers
all_source_ranks %>% 
  #Filter to just outliers
  filter(estimation_outliers == F,
         comname %in% subsampling_suspects)%>% 
  ggplot(aes(number_caught, lw_underestimate)) +
      geom_point(color = gmri_cols("gmri blue")) +
      geom_smooth(formula = y ~ x, method = "lm", color = gmri_cols("orange")) +
      facet_wrap(~comname, scales = "free", ncol = 2) +
      labs(x = "Single Station Abundance", y = "survdat$BIOMASS - LW Biomass",
           caption = "Data shown are outliers in estimate error between BIOMASS and L-W weights.")


```


