---
title: "Size Spectrum Sensitivity"
author: "Adam A. Kemberling"
date: "9/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

####  Packages  ####
library(here)
library(gmRi)
library(sf)
library(janitor)
library(sizeSpectra)
library(patchwork)
library(tidyverse)



####  Support Functions  ####
source(here("R/support/sizeSpectra_support.R"))



#File Paths
mills_path <- shared.path(os.use = "unix", group = "Mills Lab", folder = NULL)
res_path <- shared.path(os.use = "unix", group = "RES Data", folder = NULL)

####  Data  ####
dataBin <- read_csv(here("data/NEFSC/nefsc_databin_allsizes.csv"),
                    guess_max = 1e5, col_types = cols())
```

# NEFFSC Size Spectra Sensitivity Exploration

Goal of this analysis is to explore the sensitivity of NEFSC size spectra slopes to adjustments in minimum/maximum bodymass cutoffs.

Additional exploration will be done to examine catch composition through:
 * Percent of total biomass sourced from different classes of fishes  
 * Cumulative distribution figures, examining the % of total biomass within discrete weight bins
 
 
```{r}

```
 
 