---
title: "NMFS Trawl Data Cleaning"
author: "LGCarlson"
date: "6/24/2019"
output:
  html_document: default
---

### NMFS Trawl Data Cleaning

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)

# Detect the operating system
os.use<- .Platform$OS.type

# Set path to shared folders

           #########CHANGE THIS COMPUTER NAME IF PC#######
computer.name<- "lcarlson" # Needed for PC users

shared.path<- switch(os.use, 
                     "unix" = paste("~/Box/", user.name, sep = ""),
                     "windows" = paste("C:/Users/", computer.name, "/Box/", sep = ""))

library(tidyverse)
library(zoo)
library(gridExtra)
library(RColorBrewer)
library(viridis)

check = function(x){summary(as.factor(x))}
```


```{r load from shared path, echo=FALSE}
load(paste(shared.path, "Mills Lab/Data/Survdat_Nye_allseason.Rdata", sep = ""))
theme_set(theme_light())
```

In order of operations:
* Select columns of interest 
* Select stratum for full offshore survey
* Eliminate strata not sampled for full time span
* Select fall and spring surveys only
* Select years 1970-2017
* If biomass is NA but abundance is >0, make biomass be 0.01
* If biomass is 0 and abundance is >0, make biomass be 0.01
* If abundance is NA and biomass >0, make abundance 1
* If abundance is 0 and biomass >0, make abundance 1
* Remove values where biomass AND abundance are NAs

```{r create trawldata}
trawl.data<-survdat %>%
  dplyr::select(ID, EST_YEAR,SEASON, STRATUM, DECDEG_BEGLAT,DECDEG_BEGLON,
         SVSPP, COMNAME, CATCHSEX,BIOMASS, AVGDEPTH, ABUNDANCE, LENGTH, NUMLEN) %>%
  filter(STRATUM >= 01010 & STRATUM <= 01760) %>%
  filter(STRATUM!=1310 & STRATUM!=1320 & STRATUM!=1330 & STRATUM!=1350 &
           STRATUM!=1410 & STRATUM!=1420 & STRATUM!=1490) %>%
  filter(SEASON == "SPRING" | SEASON == "FALL") %>%
  filter(EST_YEAR >= 1970 & EST_YEAR < 2018) %>%
  mutate(BIOMASS = ifelse(is.na(BIOMASS) == TRUE & ABUNDANCE > 0,0.01, BIOMASS)) %>% 
  mutate(BIOMASS = ifelse(BIOMASS == 0 & ABUNDANCE > 0,0.01, BIOMASS)) %>% 
  mutate(ABUNDANCE = ifelse(is.na(ABUNDANCE) == TRUE & BIOMASS > 0, 1, ABUNDANCE)) %>% 
  mutate(ABUNDANCE = ifelse(ABUNDANCE == 0 & BIOMASS > 0, 1, ABUNDANCE)) %>%   
  filter(!is.na(BIOMASS),
         !is.na(ABUNDANCE))
```


Generate template of all sampling occasions by selecting distinct cases. 
In order of operations:
* Select distinct tows
* Select columns of interest for tow metadata template

```{r create template}
template<-trawl.data %>%
  distinct(ID, .keep_all = TRUE) %>%
  dplyr::select(ID, EST_YEAR, SEASON, STRATUM,DECDEG_BEGLAT,DECDEG_BEGLON,AVGDEPTH)
```



Create a dataframe to add common names by SVSPP code. 
In order of operations:
* Select only SVSPP and COMNAME
* Select distinct SVSPP values


```{r create COMNAME dataframe}
COMNAMEdata<-trawl.data %>% 
  dplyr::select(SVSPP, COMNAME) %>% 
  distinct(SVSPP, .keep_all = TRUE)
```

Biomass species-1 year-1
* Sum total biomass by species in each year. 

```{r annual sum}
biomass.spp.year<-trawl.data %>% 
  distinct(ID, COMNAME, CATCHSEX, BIOMASS, ABUNDANCE, .keep_all = TRUE) %>% 
  dplyr::group_by(SVSPP, EST_YEAR) %>% 
  dplyr::summarize(ANNUAL_SPECIES_BIOMASS = sum(BIOMASS)) 
```

Biomass species-1 tow-1
* Sum total biomass by species in each tow.
* Add "template" metadata for each tow.
* Add annual species biomass column. 
* Will use this dataframe to calculate center of biomass later on

```{r species by tow sum}
biomass.spp<-trawl.data %>% 
  distinct(ID, COMNAME, CATCHSEX, BIOMASS, ABUNDANCE, .keep_all = T) %>% 
  group_by(ID, SVSPP) %>% 
  dplyr::summarize(SPECIES_BIOMASS = sum(BIOMASS), SPECIES_ABUNDANCE = sum(ABUNDANCE)) %>% 
  right_join(template, by = "ID") %>% 
  left_join(biomass.spp.year, by = c("EST_YEAR", "SVSPP"))
```


Calculate total biomass per tow, not grouped by species. Will use this later to calculate proportion of species biomass per by tow.

```{r biomass per tow sum}
biomass.tow<-trawl.data %>% 
  distinct(ID, COMNAME, CATCHSEX, BIOMASS, ABUNDANCE, .keep_all = T) %>% 
  group_by(ID) %>% 
  dplyr::summarize(TOW_BIOMASS = sum(BIOMASS), TOW_ABUNDANCE = sum(ABUNDANCE)) %>% 
  right_join(template, by = "ID")
```


Calculate the number of species in each tow. This is a little indirect, but it does the job.
The output (NSPECIES) 

```{r species per tow sum}
spp.tow<-trawl.data %>% 
  group_by(ID) %>% 
  dplyr::count(SVSPP) %>%              # number of times a species was encountered in each tow
  mutate(NSPECIES = n/n) %>%    # assigns each present species a "1"
  group_by(ID) %>% 
  summarise(NSPECIES = sum(NSPECIES))     #calculates the total number of species in each tow
```


Create conversion factor to adjust NUMLEN column according to ABUNDANCE (which has been modified for gear/etc by NOAA)
```{r, message=F, warning=F}
conversion_factor<-trawl.data %>% 
  group_by(ID, COMNAME, CATCHSEX, ABUNDANCE) %>% 
  dplyr::summarize(ABUNDANCE_RAW = sum(NUMLEN)) %>% 
  mutate(CONVERS = ABUNDANCE/ABUNDANCE_RAW)
```

Adjust NUMLEN colum and check that this worked
```{r}
adjusted_numlen<-trawl.data %>% 
  left_join(conversion_factor, by = c("ID", "COMNAME", "CATCHSEX", "ABUNDANCE")) %>% 
  mutate(NUMLEN_ADJ = NUMLEN*CONVERS) %>% 
  mutate(NUMLEN_ADJ = round(NUMLEN_ADJ,2)) 

adjusted_numlen %>% 
  group_by(ID, COMNAME, CATCHSEX, ABUNDANCE) %>% 
  dplyr::summarize(ABUNDANCE_ADJ = sum(NUMLEN_ADJ)) %>% 
  mutate(diff = ABUNDANCE_ADJ - ABUNDANCE) %>% 
  ggplot(aes(diff)) + geom_histogram()


numlen_final<-adjusted_numlen %>% 
  dplyr::select(ID, SVSPP, CATCHSEX, ABUNDANCE, NUMLEN_ADJ) %>% 
  mutate_at(5, round, 2) 


adjusted_numlen_trawl<-adjusted_numlen %>% 
  dplyr::select(-ABUNDANCE_RAW, -CONVERS)
#write_rds(adjusted_numlen_trawl, path=paste(shared.path, "Mills Lab/Projects/Pew_project/Data/NMFS_trawl_adjnumlen.rds", sep = ""))

```


* Calculate the mean length of each species weighted by the number of individuals caught of that size (ie. 10 fish of length 10 count for more than 1 fish of length 4). mean = 9.45 not mean = 7
* Use distinct to create one weighted mean for each species in each tow

length.spp:
* Nest all length measurements by species in a "data" column. 
* Join with the mean length dataframe for possible future analysis of fish lengths/sizes.

```{r fish lengths}
mean.lengths<-trawl.data %>% 
  left_join(numlen_final, by = c("ID","SVSPP", "CATCHSEX","ABUNDANCE")) %>% 
  group_by(SVSPP,ID) %>% 
  mutate(MEAN_LENGTH = weighted.mean(LENGTH, NUMLEN_ADJ,na.rm = TRUE), n_LENGTHS = sum(NUMLEN_ADJ)) %>% 
  dplyr::select(ID, SVSPP, MEAN_LENGTH, n_LENGTHS) %>% 
  distinct(ID, SVSPP, .keep_all = TRUE)

length.spp<-trawl.data %>% 
  left_join(numlen_final, by = c("ID","SVSPP", "CATCHSEX","ABUNDANCE")) %>%
  group_by(ID, SVSPP) %>% 
  nest(c(LENGTH,NUMLEN_ADJ), key = "LENGTH") %>% 
  left_join(mean.lengths, by = c("ID", "SVSPP"))
```

Save the nested length data with other relevant information into an rds file for future use.

```{r}
lengthdata<-length.spp %>% 
  left_join(biomass.spp, by = c("ID", "EST_YEAR","SVSPP","SEASON","STRATUM","DECDEG_BEGLAT","DECDEG_BEGLON","AVGDEPTH")) 

#saveRDS(lengthdata, file = "NMFS_trawl_nestedlengths.rds")
```





```{r, include=FALSE}
#Create the ultimate "longer" dataset by allowing one row per individual measured. 
#* Use the "uncount()" function to duplicate rows according to a weighting variable (NUMLEN in this case)
#* Calculate the number of measured invididuals versus the total number caught

trawl_long<-trawl.data %>% 
  filter(NUMLEN > 0) %>% 
  uncount(NUMLEN)


nmeasured<-trawl_long %>% 
  group_by(ID, COMNAME) %>% 
  count(LENGTH) %>% 
  group_by(ID, COMNAME) %>% 
  summarise(nmeasured = sum(n))

nmeasured2<- trawl.data %>% 
  group_by(ID, COMNAME) %>% 
  dplyr::summarise(nmeasured = sum(NUMLEN))

notmeasured<-biomass.spp %>% 
  left_join(COMNAMEdata, by = "SVSPP") %>% 
    dplyr::select(ID, COMNAME, SPECIES_ABUNDANCE) %>% 
  left_join(nmeasured2, by = c("ID", "COMNAME")) %>% 
  mutate(notmeasured = SPECIES_ABUNDANCE - nmeasured)

notmeasured %>% 
  group_by(COMNAME) %>% 
  dplyr::summarize(total_abundance = sum(SPECIES_ABUNDANCE, na.rm = TRUE), total_measured = sum(nmeasured, na.rm = TRUE), total_notmeasured = sum(notmeasured, na.rm = TRUE))  %>% 
  mutate(percent_measured = total_measured/total_abundance) %>% 
  filter(COMNAME %in% species58) %>% 
  mutate(COMNAME = fct_reorder(COMNAME, percent_measured)) %>% 
  ggplot(aes(x = COMNAME, y = percent_measured)) + geom_point() + coord_flip()
  
```







### Create "clean dataset" in a long form by tow ID and species

Long data with biomass by species by tow, mean length, and tow metadata (sum biomass, avg depth, lat/long, etc). 
In order of operations:
* Start with biomass by species by tow data
* Add common name data 
* Select tow#, species number, common name, and biomass
* Join biomass.tow to later calculate proportion of tow's biomass by species

```{r long cleaned data}
clean_trawl<-biomass.spp %>% 
  left_join(COMNAMEdata, by = "SVSPP") %>% 
  dplyr::select(ID, SVSPP, COMNAME, SPECIES_BIOMASS) %>% 
  left_join(mean.lengths, by = c("ID", "SVSPP")) %>% 
  left_join(biomass.tow, by = "ID") %>% 
  mutate(DECADE = 10*EST_YEAR %/% 10)
```

### Create "tidy dataset" where all species level information is nested

Tidy dataset by tow with fully nested species data. 

In order of operations:
* Start with biomass by species by tow data
* Add common name data 
* Join biomass.tow to later calculate proportion of tow's biomass by species
* Select columns of interest 
* Calculate proportion of tow's biomass by species
* Remove Tow biomass column
* Nest species-level data within each tow (not a long dataset)
* Add in tow biomass and number of species per tow data

```{r tidy cleaned data}
tidy_trawl<-mean.lengths %>% 
  left_join(biomass.spp, by = c("ID", "SVSPP")) %>% 
  left_join(COMNAMEdata, by = "SVSPP") %>% 
  left_join(biomass.tow, by = "ID") %>% 
  dplyr::select(ID, SVSPP, COMNAME, SPECIES_BIOMASS, MEAN_LENGTH, n_LENGTHS, TOW_BIOMASS) %>% 
  mutate(BIOMASS_PROP = SPECIES_BIOMASS/TOW_BIOMASS) %>% 
  dplyr::select(-TOW_BIOMASS) %>% 
  group_by(ID) %>% 
  nest() %>% 
  left_join(biomass.tow, by = "ID") %>% 
  left_join(spp.tow, by = "ID")
```


### Center of biomass 

Calculate center of species biomass for each year.

In order of operations:
* Start with biomass by species by tow data
* Remove biomass NAs
* Calculate the biomass weighted latitude and longitude for each tow
* Calculate the center of mass latitude and longitude
* Remove any lat/long NA values
* Add common name data

LGC: Corrected typo:
>summarise(CENTER_LAT = sum(weightedLAT), CENTER_LAT = sum(weightedLON)) %>% filter(!is.na(CENTER_LAT),
         !is.na(CENTER_LAT)) %>%
>  left_join(COMNAMEdata, by = "SVSPP")

```{r center of bio}
centerofbiomass<-biomass.spp %>% 
  filter(!is.na(SPECIES_BIOMASS), 
                !is.na(ANNUAL_SPECIES_BIOMASS)) %>% 
  mutate(weightedLAT = (SPECIES_BIOMASS/ANNUAL_SPECIES_BIOMASS)*DECDEG_BEGLAT) %>% 
  mutate(weightedLON = (SPECIES_BIOMASS/ANNUAL_SPECIES_BIOMASS)*DECDEG_BEGLON) %>% 
  group_by(SVSPP,EST_YEAR) %>% 
  summarise(CENTER_LAT = sum(weightedLAT, na.rm = TRUE), CENTER_LON = sum(weightedLON, na.rm = TRUE)) %>% 
  filter(!is.na(CENTER_LAT), 
         !is.na(CENTER_LON)) %>% 
  left_join(COMNAMEdata, by = "SVSPP")
```


Select only species that occur in all years of the time series.

```{r species in all years, message=FALSE, warning=FALSE}
center_fullTS<-centerofbiomass %>% 
  group_by(SVSPP, COMNAME) %>% 
  count() %>%                                        #calculates number of years each spp occurs
  filter(n == (max(trawl.data$EST_YEAR) - min(trawl.data$EST_YEAR) + 1)) %>%   #filter missing yr species
  dplyr::select(SVSPP)

center_fullTS<-centerofbiomass %>%           #center of biomass only for species present in all years
  semi_join(center_fullTS, by = "SVSPP")        
```



```{r}
center_fullTS %>% 
  distinct(COMNAME, .keep_all = TRUE)
```

```{r}
species58<-as.vector(unique(center_fullTS$COMNAME))
```


Plot center of biomass. 

Kathy: I made these plots similar to the ones you made so you can see that the calculations turned out the same.

```{r plotting loop, warning=FALSE,message=FALSE}
loopbydf<-COMNAMEdata %>% 
  semi_join(center_fullTS, by = "SVSPP") %>% 
  arrange(SVSPP) %>% 
  mutate(loopby = seq(1,62,1))

plotlist<-list()

for(i in loopbydf$loopby){
tempdf<-center_fullTS %>% 
  filter(SVSPP == loopbydf$SVSPP[loopbydf$loopby == i]) %>% 
  mutate(smoothed = c(NA,NA,rollmean(CENTER_LAT,5),NA,NA))

plotlist[[i]]<-ggplot() + geom_point(data = tempdf,aes(x = EST_YEAR, y = CENTER_LAT)) + geom_line(data = tempdf,aes(x = EST_YEAR, y = CENTER_LAT), size = .5) + geom_line(data = tempdf,aes(x = EST_YEAR, y = smoothed), size = 0.9, color = "blue") + labs(x = "Year", y = "Center of latitude", title = paste('Species:', unique(tempdf$COMNAME))) + theme(panel.grid = element_blank())
}

#just plots the first 8 for example
list1 = plotlist[c(1:8)]
do.call(grid.arrange,c(list1, ncol = 4))
```


Bring in species traits data and pull out "functional groups." 
```{r functional groups, message=FALSE, warning=FALSE}
speciestraits<-read_csv(file = paste(shared.path, "Mills Lab/Projects/Pew_project/Metadata/species names and characteristics.csv", sep = "")) %>% 
  dplyr::select(SVSPP, COMNAME, HARE_ET_AL_2016_Functional_Group) %>% 
  rename("functgroup" = "HARE_ET_AL_2016_Functional_Group") %>% 
  replace(.,is.na(.), "Unclassified")  
```

### Some length explorations

Long data with biomass by species by tow, nested lengths by species, and tow metadata (sum biomass, avg depth, lat/long, etc). 

```{r}
nestedlengths<-biomass.spp %>% 
  left_join(COMNAMEdata, by = "SVSPP") %>% 
  dplyr::select(ID, SVSPP, COMNAME, SPECIES_BIOMASS, SPECIES_ABUNDANCE) %>% 
  left_join(length.spp, by = c("ID", "SVSPP")) %>% 
  left_join(biomass.tow, by = "ID") %>% 
  mutate(DECADE = 10*EST_YEAR %/% 10)
```


```{r}
nestedlengths %>% 
  mutate_at(14:15, round,0) %>% 
  left_join(speciestraits, by = "COMNAME") %>% 
  group_by(DECADE,DECDEG_BEGLAT,functgroup) %>% 
  filter(!is.na(functgroup)) %>% 
  summarise(mLENGTH = mean(MEAN_LENGTH, na.rm = TRUE), sd = sd(MEAN_LENGTH,na.rm = TRUE)) %>% 
  ggplot(aes(DECDEG_BEGLAT,mLENGTH)) + geom_errorbar(aes(ymin=mLENGTH-sd,ymax=mLENGTH+sd),width = 0.6) +
  geom_point() + facet_grid(rows = vars(fct_reorder(functgroup,mLENGTH)), cols = vars(DECADE)) + 
  labs(x = "Latitude (N)", y = "Length (cm)") + theme(panel.grid = element_blank()) + coord_flip() + 
  theme(axis.text = element_text(size = 12),axis.title = element_text(size = 14), 
        strip.text = element_text(size = 12))
```

```{r}
nestedlengths %>% 
  mutate(DEPTH_ROUND = 50*AVGDEPTH %/% 50) %>% 
  left_join(speciestraits, by = "COMNAME") %>% 
  group_by(DECADE,DEPTH_ROUND,functgroup) %>% 
  filter(!is.na(functgroup)) %>% 
  summarise(mLENGTH = mean(MEAN_LENGTH, na.rm = TRUE), sd = sd(MEAN_LENGTH,na.rm = TRUE)) %>% 
  ggplot(aes(DEPTH_ROUND,mLENGTH)) + geom_errorbar(aes(ymin=mLENGTH-sd,ymax=mLENGTH+sd),width = 25) +
  geom_point() + coord_flip() + scale_x_continuous(breaks = c(0,150,300,450)) + 
  facet_grid(rows = vars(fct_reorder(functgroup,mLENGTH)), cols = vars(DECADE)) + 
  labs(x = "Depth (m)", y = "Length (cm)") + theme(panel.grid = element_blank()) + 
  theme(axis.text = element_text(size = 12),axis.title = element_text(size = 14), 
        strip.text = element_text(size = 12)) 
```



```{r}
commonspp<-c("SILVER HAKE", "SPINY DOGFISH", "RED HAKE", "LONGFIN SQUID", "GOOSEFISH", "AMERICAN LOBSTER")

nestedlengths %>% 
  mutate_at(14:15, round,0) %>% 
  filter(COMNAME %in% commonspp) %>% 
  group_by(DECDEG_BEGLAT, DECADE, COMNAME) %>% 
  summarise(mLENGTH = mean(MEAN_LENGTH, na.rm = TRUE), sd = sd(MEAN_LENGTH,na.rm = TRUE)) %>% 
  ggplot(aes(DECDEG_BEGLAT,mLENGTH)) + geom_errorbar(aes(ymin=mLENGTH-sd,ymax=mLENGTH+sd),width = 0.6) +
  geom_point() + facet_grid(rows = vars(fct_reorder(COMNAME,mLENGTH)), cols = vars(DECADE)) + 
  coord_flip() + labs(x = "Latitude (N)", y = "Length (cm)") + theme(panel.grid = element_blank()) +
  theme(axis.text = element_text(size = 12),axis.title = element_text(size = 14), 
        strip.text = element_text(size = 12)) + scale_x_continuous(breaks = c(0,150,300,450))
```


```{r}
clean_trawl %>% 
  group_by(EST_YEAR) %>% 
  summarize(min = min(MEAN_LENGTH, na.rm = TRUE), max = max(MEAN_LENGTH, na.rm = TRUE), q25 = quantile(MEAN_LENGTH,probs = 0.25, na.rm = TRUE), q75 = quantile(MEAN_LENGTH,probs = 0.75, na.rm = TRUE), mean = mean(MEAN_LENGTH, na.rm = TRUE), n = n())
```


```{r}
clean_trawl %>% 
  group_by(DECADE) %>% 
  ggplot(aes(MEAN_LENGTH)) + geom_histogram() + facet_wrap(~DECADE) + xlim(0,200)
```


```{r}
p <- c(0.05,0.1,0.15,0.20,0.25,0.75,0.80,0.85,0.90,0.95)

p_names <- paste0(p*100, "%")
p_funs <- map(p, ~partial(quantile, probs = .x, na.rm = TRUE)) %>% 
  set_names(nm = p_names)

quanttable<-clean_trawl %>% 
  mutate(DECADE = 10*EST_YEAR %/% 10) %>% 
  filter(!is.na(DECADE)) %>%
  group_by(DECADE) %>% 
  summarize_at(vars(MEAN_LENGTH), p_funs)
quanttable
```

```{r}
decadevector<-as.vector(unique(quanttable$DECADE))
clean_trawl<-clean_trawl %>% 
  filter(!is.na(MEAN_LENGTH))

clean_trawl$quant = 0.5

for(i in decadevector){
  
temp_quanttable<-quanttable %>% 
  filter(DECADE == i)

if(clean_trawl$DECADE == i){
clean_trawl$quant[clean_trawl$MEAN_LENGTH <= temp_quanttable$`25%`] = "0.25"
clean_trawl$quant[clean_trawl$MEAN_LENGTH <= temp_quanttable$`10%`] = "0.10"

clean_trawl$quant[clean_trawl$MEAN_LENGTH >= temp_quanttable$`75%`] = "0.75"
clean_trawl$quant[clean_trawl$MEAN_LENGTH >= temp_quanttable$`90%`] = "0.90"

}
else{clean_trawl$quant = clean_trawl$quant}
}

```



```{r}
clean_trawl %>% 
  group_by(DECADE) %>% 
  ggplot(aes(MEAN_LENGTH, fill = quant)) + geom_histogram() + facet_wrap(~DECADE, ncol = 5) + xlim(0,180) + scale_fill_manual(values = c("#252525","#737373","#bdbdbd","#737373","#252525")) + labs(x = "Length (cm)", y = "count", fill = "Quantile") + theme(panel.grid = element_blank()) + theme(strip.background = element_blank(), strip.text = element_text(size = 12, color = "black"))

#ggsave(filename = "quantilehist.pdf", plot = last_plot(), height = 3.5, width = 11)
```

```{r}
clean_trawl %>% 
  group_by(DECADE, quant) %>% 
  summarise(n = n(), mean = mean(MEAN_LENGTH, na.rm = TRUE))
```

```{r}

clean_trawl %>% 
  left_join(speciestraits, by = "COMNAME") %>% 
  filter(MEAN_LENGTH < 200, 
         !is.na(functgroup)) %>% 
  ggplot(aes(MEAN_LENGTH, fill = functgroup)) + geom_histogram(binwidth = 1) + facet_wrap(~quant, scales = "free_x", ncol = 5) + labs(x = "Length (cm)", y = "count", fill = "Functional group") + theme(panel.grid = element_blank()) + theme(strip.background = element_blank(), strip.text = element_text(size = 12, color = "black")) + scale_fill_manual(values = rev(brewer.pal(9, "PuBuGn")))

#ggsave(filename = "functgrouphist.pdf", plot = last_plot(), height = 3.5, width = 11)
```

```{r}
clean_trawl %>% 
  mutate(LATblock = 2*DECDEG_BEGLAT %/% 2) %>% 
  group_by(LATblock) %>% 
  summarize(n = n())
```


```{r, warning=FALSE, message=FALSE}
speciesvector<-as.vector(unique(center_fullTS$COMNAME))

for(s in speciesvector){
  
temp_df<-clean_trawl %>% 
  mutate(LATblock = 2*DECDEG_BEGLAT %/% 2) %>% 
  filter(COMNAME == s) %>% 
  group_by(DECADE, LATblock) %>% 
  summarise(mLENGTH = mean(MEAN_LENGTH), sd = sd(MEAN_LENGTH), n = sum(n_LENGTHS)) 

speciesplot<-ggplot(temp_df, aes(LATblock, mLENGTH, label = n)) +
  geom_errorbar(aes(ymin=mLENGTH-sd,ymax=mLENGTH+sd),width = 1) +
  geom_point() + coord_flip() + facet_wrap(~DECADE, ncol = 5) + geom_text(hjust = -0.25, vjust = -0.25) +
  theme(panel.grid = element_blank()) + xlim(33,45) +
  labs(y = "Length (cm)", x = "Latitude", title = (paste('Species:', s, sep=' '))) +
  theme(strip.background = element_blank(), strip.text = element_text(size = 12, color = "black")) + 
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 12))
print(speciesplot)

#name = paste('lengthxlat', s, sep='_')
#ggsave(speciesplot,filename = paste(shared.path, "Mills Lab/Projects/Pew_project/Leading-Trailing_Edge_Project/Temp_Figures/lengths/byspecies/", name,".pdf",sep = ""), width = 11, height = 4)
}
```


```{r}
clean_trawl %>% 
  filter(COMNAME %in% speciesvector) %>% 
  group_by(COMNAME) %>% 
  summarise(mLENGTH = mean(MEAN_LENGTH, na.rm = TRUE), sd = sd(MEAN_LENGTH, na.rm = TRUE), maxLENGTH = max(MEAN_LENGTH, na.rm = TRUE), mLAT = weighted.mean(DECDEG_BEGLAT,SPECIES_BIOMASS, na.rm = TRUE), n = n(), se = (sd/n)) %>% 
  left_join(speciestraits, by = "COMNAME") %>% 
  ggplot(aes(fct_reorder(COMNAME,mLENGTH), mLENGTH, 
            color = fct_relevel(functgroup, "Elasmobranch", "Coastal", "Unclassified",
                                "Groundfish","Pelagic", "Invertebrates"))) +
  geom_errorbar(aes(ymin=mLENGTH-sd,ymax=mLENGTH+sd),width = 1) +
  geom_point() + coord_flip() +
  theme(panel.grid = element_blank()) +
  labs(y = "Mean length (cm)", x = "", color = "Functional group") +
  theme(strip.background = element_blank(), strip.text = element_text(size = 12, color = "black")) + 
  theme(axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 7), axis.title = element_text(size = 12)) + scale_color_viridis(discrete = TRUE)

#ggsave(filename = "speciesmeanlength.pdf", plot = last_plot(), height = 8.5, width = 10)
```

















